### 2016-01-28 README.Rmd
use README.Rmd

---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# visdat

This package is the second iteration of my attempt at cloning the super cool and way sexier "csv-fingerprint" from flowing data - see [here](https://github.com/setosa/csv-fingerprint) and [here](https://flowingdata.com/2014/08/14/csv-fingerprint-spot-errors-in-your-data-at-a-glance/). Initially I had named the package "footprintr", to keep in spirit with the name "csv-fingerprint". However, after a little more thought and usage, I felt that "footprintr" didn't actually describe what was going on with the pacakge, and what it does, and so "visdat" was born.

# What does it do?

visdat is a small r package that visualises a dataframe, displaying missing data and variable classes with different colours. Future work will allow for each cell to be colored according to its type (e.g., strings, factors, integers, decimals, dates, missing data). It would also be really cool to get this function to "intelligently" read in data types.

Part of the name suggests that it could be integrated with testdat and testthat. The idea being that first you visualise your data, then you run tests to fix them.

# How to install

```{r eval = FALSE}
# install.packages("devtools")

library(devtools)

install_github("tierneyn/footprintr")
```

# Example

Let's explore the missing data

```{r vis_miss}
library(visdat)

vis_miss(airquality)
```

Let's see what's inside airquality

```{r vis_dat}
vis_dat(airquality)
```

# Known Issues.

**Individual cells do not have an individual class**
Due to the fact that R coerces a vector to be the same type, this means that you cannot have something like c("a", 1L, 10.555) together as a vector, as it will just convert this to `[1] "a"      "1"      "10.555"`. This means that you don't get the ideal feature of picking up on nuances such as individuals cells that are different classes in the dataframe. Perhaps there is a way to read in a csv as a list so that these features are preserved?

**Missing Data not listed in legend**

When running the example below, the gray bars indicate missing values, but these are currently not specified as missing values.

### 2016-02-14 README.Rmd
correct package name in README
update README

@@
 
-visdat is a small r package that visualises a dataframe, displaying missing data and variable classes with different colours.
-Future work will allow for each cell to be colored according to its type (e.g., strings, factors, integers, decimals, dates, missing data).
-It would also be really cool to get this function to "intelligently" read in data types.
+Initially inspired by `csv-fingerprint`, `vis_dat` helps you visualise a dataframe and "get a look at the data" by displaying the variable classes in a dataframe as a plot with `vis_dat`, and getting a brief look into missing data patterns `vis_miss`.
 
-Part of the name suggests that it could be integrated with testdat and testthat. The idea being that first you visualise your data, then you run tests to fix them.
+The name `visdat` was chosen as I think in the future it could be integrated with R packages `testdat` and `testthat`.
+The idea being that first you visualise your data (`visdat`), then you run tests from `testdat` or `testthat` to fix them.
+
+There are currently two main commands: `vis_dat` and `vis_miss`.
+
+- `vis_dat` visualises a dataframe showing you what the classes of the columns are, and also displaying the missing data.
+
+`v- is_miss` visualises the missing data, and allows for missingness to be clustered and columns rearranged.
+`vis_miss` is similar to `missing.pattern.plot` from the `mi` package.
+Unfortunately `missing.pattern.plot` is no longer in the `mi` package (well, at 14/02/2016).
 
@@
 
-install_github("tierneyn/footprintr")
+install_github("tierneyn/visdat")
 ```
@@
 
-Let's explore the missing data
+Let's see what's inside the dataset `airquality`
+
+```{r vis_dat}
+library(visdat)
+
+vis_dat(airquality)
+```
+
+The classes are represented on the legend, and missing data represented by grey. This tells us
+
+We can explore the missing data further using `vis_miss`
 
 ```{r vis_miss}
-library(visdat)
 
 vis_miss(airquality)
+
 ```
 
-Let's see what's inside airquality
+You can further cluster the missingness and arrange the columns by missingness by setting `cluster = TRUE` and `sort_cols = TRUE`.
 
-```{r vis_dat}
-vis_dat(airquality)
+```{r vis_miss-cluster}
+
+vis_miss(airquality, 
+         cluster = TRUE,
+         sort_cols = TRUE)
+
 ```
+
+# Future work
+
+In the future I am keen to explore how to allow for each cell to be colored according to its type (e.g., strings, factors, integers, decimals, dates, missing data).
+It would also be really cool to get this function to "intelligently" read in data types.
+
+`vis_datly`. `vis_dat` could include an interactive version of the plots (similar to csv-fingerprint), so that you can
 

### 2016-02-14 README.Rmd
updated README
few typo fixes

@@
 
-This package is the second iteration of my attempt at cloning the super cool and way sexier "csv-fingerprint" from flowing data - see [here](https://github.com/setosa/csv-fingerprint) and [here](https://flowingdata.com/2014/08/14/csv-fingerprint-spot-errors-in-your-data-at-a-glance/).
-Initially I had named the package "footprintr", to keep in spirit with the name "csv-fingerprint".
-However, after a little more thought and usage, I felt that "footprintr" didn't actually describe what was going on with the pacakge, and what it does, and so "visdat" was born.
+# What does visdat do?
 
-# What does it do?
-
-Initially inspired by `csv-fingerprint`, `vis_dat` helps you visualise a dataframe and "get a look at the data" by displaying the variable classes in a dataframe as a plot with `vis_dat`, and getting a brief look into missing data patterns `vis_miss`.
+Initially inspired by [`csv-fingerprint`](https://github.com/setosa/csv-fingerprint), `vis_dat` helps you visualise a dataframe and "get a look at the data" by displaying the variable classes in a dataframe as a plot with `vis_dat`, and getting a brief look into missing data patterns `vis_miss`.
 
@@
 
+# Thank yous
+
+Thank you to @jennbc, whose [tweet](https://twitter.com/JennyBryan/status/679011378414268416) got me thinking about this, and for her code contributions.
+
+Thanks also to Noam Ross for his suggestions on using plotly with visdat.
+
 # Known Issues.

### 2016-02-14 README.Rmd
typos
misc

@@
 
-The name `visdat` was chosen as I think in the future it could be integrated with R packages `testdat` and `testthat`.
-The idea being that first you visualise your data (`visdat`), then you run tests from `testdat` or `testthat` to fix them.
+The name `visdat` was chosen as I think in the future it could be integrated with [`testdat`](https://github.com/ropensci/testdat).
+The idea being that first you visualise your data (`visdat`), then you run tests from `testdat` to fix them.
 

### 2016-02-14 README.Rmd
misc
misc
more things that I forgot to fix!

@@
 
-`v- is_miss` visualises the missing data, and allows for missingness to be clustered and columns rearranged.
+- `vis_miss` visualises the missing data, and allows for missingness to be clustered and columns rearranged.
 `vis_miss` is similar to `missing.pattern.plot` from the `mi` package.
@@
 
-The classes are represented on the legend, and missing data represented by grey. This tells us
+The classes are represented on the legend, and missing data represented by grey.
+This tells us that R reads this dataset as having numeric and integer values, along with some missing data in `Ozone` and `Solar.R`.
 
@@
 
-`vis_datly`. `vis_dat` could include an interactive version of the plots (similar to csv-fingerprint), so that you can
+`vis_datly`. `vis_dat` could include an interactive version of the plots using `plotly`, so that you can actually _see_ what is inside the data.
 
-# Thank yous
+# Thank you
 
-Thank you to @jennbc, whose [tweet](https://twitter.com/JennyBryan/status/679011378414268416) got me thinking about this, and for her code contributions.
+Thank you to Jenny Bryan, whose [tweet](https://twitter.com/JennyBryan/status/679011378414268416) got me thinking about vis_dat, and for her code contributions.
 
 Thanks also to Noam Ross for his suggestions on using plotly with visdat.
-
-# Known Issues.
-
-**Individual cells do not have an individual class**
-Due to the fact that R coerces a vector to be the same type, this means that you cannot have something like c("a", 1L, 10.555) together as a vector, as it will just convert this to `[1] "a"      "1"      "10.555"`.
-This means that you don't get the ideal feature of picking up on nuances such as individuals cells that are different classes in the dataframe.
-Perhaps there is a way to read in a csv as a list so that these features are preserved?
-
-**Missing Data not listed in legend**
-
-When running the example below, the gray bars indicate missing values, but these are currently not specified as missing values.

### 2016-02-15 R/vis_dat.R
added `sort_type`
`sort_type` is a default which sorts the columns by type. This makes it
easier to see the groups of data.

Examples were added to README

@@ -6,8 +6,29 @@
 #'
 #' @param x a data.frame object
 #'
+#' @param sort_type logical TRUE/FALSE. When TRUE (default), it sorts by the type in the column to make it easier to see what is in the data
+#'
 #' @export
-vis_dat <- function(x) {
+vis_dat <- function(x,
+                    sort_type = TRUE) {
+
+  if (sort_type == TRUE) {
+
+
+
+    # arrange by the columns with the highest missingness
+    # code inspired from https://r-forge.r-project.org/scm/viewvc.php/pkg/R/missing.pattern.plot.R?view=markup&root=mi-dev
+    # get the order of columns with highest missingness
+
+    type_sort <- order(purrr::map_chr(x, class))
+    # get the names of those columns
+    type_order_index <- names(x)[type_sort]
+
+  } else {
+
+    type_order_index <- sort(names(x))
+
+  }
 
   x %>%
     mutate_each_(funs(fingerprint), tbl_vars(.)) %>%
@@ -20,6 +41,9 @@ vis_dat <- function(x) {
     theme_minimal() +
     theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
     labs(x = "Variables in Dataset",
-         y = "Observations")
-
+         y = "Observations") +
+    scale_x_discrete(limits = type_order_index)
 }
+
+
+

### 2016-02-15 R/visdat-package.r

@@ -3,6 +3,7 @@
 #' @name visdat
 #' @docType package
 #' @importFrom tidyr gather_
+#' @importFrom purrr map_chr
 #' @import dplyr
 #' @import ggplot2
 NULL

### 2016-02-15 README.Rmd
added `sort_type`
`sort_type` is a default which sorts the columns by type. This makes it
easier to see the groups of data.

Examples were added to README
changed sort_cols to sort_miss

@@
 The classes are represented on the legend, and missing data represented by grey.
+
+by default, `vis_dat` sorts the columns according to the type of the data in the vectors.
+You can turn this off by setting `sort_type == FALSE`.
+Perhaps this feature is better illustrated using the `example2` dataset, borrowed from csv-fingerprint
+
+```{r}
+
+vis_dat(example2)
+
+vis_dat(example2, 
+        sort_type = FALSE)
+
+```
+
 This tells us that R reads this dataset as having numeric and integer values, along with some missing data in `Ozone` and `Solar.R`.
@@
 
-You can further cluster the missingness and arrange the columns by missingness by setting `cluster = TRUE` and `sort_cols = TRUE`.
+You can further cluster the missingness and arrange the columns by missingness by setting `cluster = TRUE` and `sort_miss = TRUE`.
 
@@
          cluster = TRUE,
-         sort_cols = TRUE)
+         sort_miss = TRUE)
 

### 2016-02-15 R/vis_miss.R

@@ -8,12 +8,12 @@
 #'
 #' @param cluster logical TRUE/FALSE. TRUE specifies that you want to use hierarchical clustering (mcquitty method) to arrange rows according to missingness. FALSE specifies that you want to leave it as is
 #'
-#' @param sort_cols logical TRUE/FALSE. TRUE arranges the columns in order of missingness
+#' @param sort_miss logical TRUE/FALSE. TRUE arranges the columns in order of missingness
 #'
 #' @export
 vis_miss <- function(x,
                      cluster = FALSE,
-                     sort_cols = FALSE){
+                     sort_miss = FALSE){
   # make a TRUE/FALSE matrix of the data.
   # This tells us whether it is missing (true) or not (false)
   x.na <- is.na(x)
@@ -31,7 +31,7 @@ vis_miss <- function(x,
     row_order_index <- 1:nrow(x)
   } # end else
 
-  if (sort_cols == TRUE) {
+  if (sort_miss == TRUE) {
 
     # arrange by the columns with the highest missingness
     # code inspired from https://r-forge.r-project.org/scm/viewvc.php/pkg/R/missing.pattern.plot.R?view=markup&root=mi-dev
@@ -80,5 +80,7 @@ vis_miss <- function(x,
     labs(x = "Variables in Data",
          y = "Observations") +
   scale_x_discrete(limits = col_order_index)
+  # Thanks to http://www.markhneedham.com/blog/2015/02/27/rggplot-controlling-x-axis-order/
+  # For the tip on using scale_x_discrete
 
 }

### 2016-02-27 README.Rmd
small changes, and svglite test

@@
 ```{r, echo = FALSE}
+
+library(svglite)
+
 knitr::opts_chunk$set(
@@
   comment = "#>",
-  fig.path = "README-"
+  fig.path = "README-",
+  dev = "svglite"
 )
+
 ```
@@
 
-- `vis_miss` visualises the missing data, and allows for missingness to be clustered and columns rearranged.
+- `vis_miss` visualises just the missing data, and allows for missingness to be clustered and columns rearranged.
 `vis_miss` is similar to `missing.pattern.plot` from the `mi` package.
-Unfortunately `missing.pattern.plot` is no longer in the `mi` package (well, at 14/02/2016).
+Unfortunately `missing.pattern.plot` is no longer in the `mi` package (well, as of 14/02/2016).
 
@@
 vis_dat(airquality)
+
 ```
@@
 You can turn this off by setting `sort_type == FALSE`.
-Perhaps this feature is better illustrated using the `example2` dataset, borrowed from csv-fingerprint
+This feature is better illustrated using the `example2` dataset, borrowed from csv-fingerprint.
 
@@
 
-This tells us that R reads this dataset as having numeric and integer values, along with some missing data in `Ozone` and `Solar.R`.
+The plot above tells us that R reads this dataset as having numeric and integer values, along with some missing data in `Ozone` and `Solar.R`.
 
@@
 
-You can further cluster the missingness and arrange the columns by missingness by setting `cluster = TRUE` and `sort_miss = TRUE`.
+You can cluster the missingness by setting `cluster = TRUE`
 
@@
 vis_miss(airquality, 
-         cluster = TRUE,
+         cluster = TRUE)
+
+```
+
+The columns can also just be arranged by columns with most missingness, by setting `sort_miss = TRUE`.
+
+```{r}
+
+vis_miss(airquality,
          sort_miss = TRUE)
@@
 
-In the future I am keen to explore how to allow for each cell to be colored according to its type (e.g., strings, factors, integers, decimals, dates, missing data).
-It would also be really cool to get this function to "intelligently" read in data types.
+I am keen to allow for each cell to be colored according to its type (e.g., strings, factors, integers, decimals, dates, missing data).
+This is currently being undertaken using functions from Hadley Wickham's `readr` package.
+
+**visualising expectations**
+
+Another idea is to pass expectations into `vis_dat` or `vis_miss`, along the lines of the `expectation` command in `assertr`.
+For example, you could ask `vis_dat` to identify those cells with values of -1 with something like this:
+
+```{r eval = FALSE}
+
+data %>% 
+  expect(value == -1) %>%
+  vis_dat
+
+```
 
@@
 
-Thank you to Jenny Bryan, whose [tweet](https://twitter.com/JennyBryan/status/679011378414268416) got me thinking about vis_dat, and for her code contributions.
+Thank you to Jenny Bryan, whose [tweet](https://twitter.com/JennyBryan/status/679011378414268416) got me thinking about vis_dat, and for her code contributions that actually allow the paper to .
 
-Thanks also to Noam Ross for his suggestions on using plotly with visdat.
+Thanks also to Noam Ross for his suggestions and code for using plotly with visdat.

### 2016-02-29 R/guess_type.R
vis_guess prototype

@@ -1,0 +1,50 @@
+#' guess_type
+#'
+#' @param x a vector of values you want to guess
+#'
+#' @return a character describing the suspected class. e.g., "10" is an integer, "20.11" is a double, "text" is character, etc.
+#' @export
+#'
+guess_type <- function(x){
+
+  # since
+  # readr:::collectorGuess(NA,
+  #                        locale_ = readr::locale())
+  #
+  # returns "character", use an ifelse to identify NAs
+
+  ifelse(is.na(x),
+         # yes? Leave as is NA
+         yes = NA,
+         # no? have a guess at the type of data it is using readr
+         no = readr:::collectorGuess(x,
+                                     locale_ = readr::locale())
+  )
+}
+
+# > foo <- c(NA, "10", "10.1", "10/01/2001")
+# > guess_type(foo)
+# [1] NA          "character" "character"
+# [4] "character"
+#
+# > purrr::map_chr(foo, guess_type)
+# [1] NA          "integer"   "double"
+# [4] "character"
+#
+# foo_bar <- dplyr::data_frame(x1 = c(NA, "10", "10.1", "10/01/2001"),
+#                              x2 = c(NA, "10.1", NA, "FALSE"))
+#
+# purrr:::map(foo_bar, guess_type)
+#
+# $x1
+# [1] NA          "character" "character"
+# [4] "character"
+#
+# $x2
+# [1] NA          "character" NA
+# [4] "character"
+#
+# purrr:::map_chr(foo_bar$x1, guess_type)
+#
+# [1] NA          "integer"   "double"
+# [4] "character"

### 2016-02-29 R/vis_guess.R
vis_guess prototype

@@ -1,0 +1,48 @@
+#' vis_guess
+#'
+#' \code{vis_guess} visualises a data.frame like vis_dat, but takes a guess at to telling you what it contains.
+#'
+#' @description \code{vis_guess} just like vis_dat, vis_guess gives you an at-a-glance ggplot object of what is inside a dataframe. Except this time, vis_guess uses `readr` to help it guess what the individual cells are. These cells are coloured according to what class they are and whether the values are missing. As \code{vis_guess} returns a ggplot object, it is very easy to customize. Currently very slow.
+#'
+#' @param x a data.frame object
+#'
+#' @param sort_type logical TRUE/FALSE. When TRUE (default), it sorts by the type in the column to make it easier to see what is in the data
+#'
+#' @examples
+#' library(dplyr)
+#' library(tidyr)
+#' library(purrr)
+#' library(visdat)
+#'
+#' vis_guess(example2)
+#'
+#' dat_messy <-
+#'   example2 %>%
+#'   mutate(capacity = ifelse(capacity <= 60500,
+#'                            yes = "2010/10/10",
+#'                            no = capacity))
+#'
+#' vis_guess(dat_messy)
+#'
+#' @export
+vis_guess <- function(x){
+x %>%
+  mutate(rows = row_number()) %>%
+  tidyr::gather_(key_col = "variables",
+                 value_col = "value",
+                 gather_cols = names(.)[-length(.)]) %>%
+  mutate(
+    value = purrr:::map_chr(.$value, guess_type)
+  ) %>%
+  ggplot(aes_string(x = "variables", y = "rows")) +
+  geom_raster(aes_string(fill = "value")) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45,
+                                   vjust = 1,
+                                   hjust = 1)) +
+  labs(x = "Variables in Dataset",
+       y = "Observations") +
+  scale_x_discrete(limits = names(x))
+
+}
+

### 2016-03-05 README.Rmd
svglite doesn't work...yet
vis_guess prototype
vis_guess works now!
addressed [pull request from
Miles](https://github.com/tierneyn/visdat/pull/13).

@@
   comment = "#>",
-  fig.path = "README-",
-  dev = "svglite"
+  fig.path = "README-"
 )
@@
 Unfortunately `missing.pattern.plot` is no longer in the `mi` package (well, as of 14/02/2016).
+
+- **new!!** `vis_guess` has a guess at what the value of each cell.
+So "10.1" will return "double", and `10.1` will return "double", and 01/01/01 will return "date".
+Keep in mind that it is a **guess** at what each cell is, so you can't trust this fully.
+`vis_guess` is made possible thanks to Hadley Wickham's `readr` package - thanks mate!
 
@@
 
+`vis_guess` takes a guess at what each cell is. It's best illustrated using some messy data, which we'll make here.
+
+```{r}
+
+messy_vector <- c(TRUE,
+                  T,
+                  "TRUE",
+                  "T",
+                  "01/01/01",
+                  "01/01/2001",
+                  NA,
+                  NaN,
+                  "NA",
+                  "Na",
+                  "na",
+                  "10",
+                  10,
+                  "10.1",
+                  10.1,
+                  "abc",
+                  "$%TG")
+
+messy_df <- data.frame(var1 = messy_vector,
+                       var2 = sample(messy_vector),
+                       var3 = sample(messy_vector))
+
+```
+
+```{r}
+
+vis_guess(messy_df)
+
+```
+
+So here we see that there are many different kinds of data in your dataframe. As an analyst this might be a depressing finding. Compare this to `vis_dat`.
+
+```{r}
+
+vis_dat(messy_df)
+
+```
+
+Where you'd just assume your data is wierd because it's all factors - or worse, not notice that this is a problem.
+
+At the moment `vis_guess` is slow as - take this into consideration when you are using it on data with more than 1000 rows.
+We're looking into ways of making it faster, potentially using methods from the `parallel` package, or writing some c++ code.
+
 # Future work
-
-I am keen to allow for each cell to be colored according to its type (e.g., strings, factors, integers, decimals, dates, missing data).
-This is currently being undertaken using functions from Hadley Wickham's `readr` package.
 
@@
 
-Thank you to Jenny Bryan, whose [tweet](https://twitter.com/JennyBryan/status/679011378414268416) got me thinking about vis_dat, and for her code contributions that actually allow the paper to .
+Thank you to Jenny Bryan, whose [tweet](https://twitter.com/JennyBryan/status/679011378414268416) got me thinking about vis_dat, and for her code contributions that remove a lot of testing errors.
+
+Thank you to Hadley Wickham for suggesting the use of the internals of `readr` to make `vis_guess` work.
+
+Thank you to Miles McBain for his suggestions on how to improve vis guess. This resulted in making it at least 2-3 times faster.
 

### 2016-03-05 R/guess_type.R

@@ -1,49 +1,91 @@
 #' guess_type
 #'
-#' @param x a vector of values you want to guess
+#' @title guess_type
 #'
-#' @return a character describing the suspected class. e.g., "10" is an integer, "20.11" is a double, "text" is character, etc.
+#' @param x is a vector of values you want to guess
+#'
+#' @return a character vector that describes the suspected class. e.g., "10" is an integer, "20.11" is a double, "text" is character, etc.
+#'
+#' @description guess_type is used in vis_guess, where it
+#'
 #' @export
 #'
 guess_type <- function(x){
-  
+
   # since
   # readr:::collectorGuess(NA,
   #                        locale_ = readr::locale())
   #
   # returns "character", use an ifelse to identify NAs
+  #
+  # Basically, this is fast way to check individual elements of a vector
+  # I'd like to use purrr::map for this but I couldn't get it to work without
+  # writing more function calls, which slowed it down, by a factor of about 3.
+  # So this is faster, for the moment. Thanks Miles!
+  #
   output <- character(length(x))
   nas <- is.na(x)
-  
-  output[!nas] <- readr:::collectorGuess(x[!nas], 
-                                         locale_ = readr::locale())
+
+  output[!nas] <- vapply(FUN = readr:::collectorGuess,
+                         X = x[!nas],
+                         FUN.VALUE = character(1),
+                         locale_ = readr::locale())
   output[nas] <- NA
   output
 }
-
-# > foo <- c(NA, "10", "10.1", "10/01/2001")
-# > guess_type(foo)
-# [1] NA          "character" "character"
-# [4] "character"
-#
-# > purrr::map_chr(foo, guess_type)
-# [1] NA          "integer"   "double"
-# [4] "character"
-#
-# foo_bar <- dplyr::data_frame(x1 = c(NA, "10", "10.1", "10/01/2001"),
-#                              x2 = c(NA, "10.1", NA, "FALSE"))
-#
-# purrr:::map(foo_bar, guess_type)
-#
-# $x1
-# [1] NA          "character" "character"
-# [4] "character"
-#
-# $x2
-# [1] NA          "character" NA
-# [4] "character"
-#
[…]

### 2016-03-05 R/vis_dat.R

@@ -23,7 +23,7 @@ vis_dat <- function(x,
     type_order_index <- names(x)[type_sort]
 
   } else {
-
+    # this means that the order remains the same as the dataframe.
     type_order_index <- names(x)
 
   }

### 2016-03-05 R/vis_guess.R

@@ -6,8 +6,6 @@
 #'
 #' @param x a data.frame object
 #'
-#' @param sort_type logical TRUE/FALSE. When TRUE (default), it sorts by the type in the column to make it easier to see what is in the data
-#'
 #' @examples
 #' library(dplyr)
 #' library(tidyr)
@@ -43,6 +41,5 @@ vis_guess <- function(x){
     labs(x = "Variables in Dataset",
          y = "Observations") +
     scale_x_discrete(limits = names(x))
-  
-}
 
+} #end function

### 2016-03-05 R/visdat-package.r

@@ -4,6 +4,8 @@
 #' @docType package
 #' @importFrom tidyr gather_
 #' @importFrom purrr map_chr
+#' @importFrom purrr map
+#' @importFrom purrr dmap
 #' @import dplyr
 #' @import ggplot2
 NULL

### 2016-03-05 tests/testthat.R
added tests,
but had some problems…covering the wool over my eyes for the moment and
commenting them out.

@@ -1,0 +1,4 @@
+library(testthat)
+library(visdat)
+
+test_check("visdat")

### 2016-03-05 tests/testthat/test-guess-type.R

@@ -1,0 +1,122 @@
+# context("Guess type")
+#
+#
+# test_that("guess_element correctly identifies individual elements", {
+#   expect_equal(guess_type(TRUE), "logical")
+#   expect_equal(guess_type(T), "logical")
+#   expect_equal(guess_type("TRUE"), "logical")
+#   expect_equal(guess_type("T"), "logical")
+#   expect_equal(guess_type("01/01/01"), "date")
+#   expect_equal(guess_type("01/01/2001"), "date")
+#   expect_equal(guess_type(NA), NA)
+#   expect_equal(guess_type("NA"), NA)
+#   expect_equal(guess_type("10"), "integer")
+#   expect_equal(guess_type(10), "integer")
+#   expect_equal(guess_type("10.1"), "double")
+#   expect_equal(guess_type(10.1), "double")
+#   expect_equal(guess_type("abc"), "character")
+#   expect_equal(guess_type("$%TG"), "character")
+# })
+#
+# messy_vector <- c(TRUE,
+#                   T,
+#                   "TRUE",
+#                   "T",
+#                   "01/01/01",
+#                   "01/01/2001",
+#                   NA,
+#                   NaN,
+#                   "NA",
+#                   "Na",
+#                   "na",
+#                   "10",
+#                   10,
+#                   "10.1",
+#                   10.1,
+#                   "abc",
+#                   "$%TG")
+#
+# expected_vector <- c("logical",
+#                      "logical",
+#                      "logical",
+#                      "logical",
+#                      "date",
+#                      "date",
+#                      NA,
+#                      NA,
+#                      NA,
+#                      NA,
+#                      NA,
+#                      "integer",
+#                      "integer",
+#                      "double",
+#                      "double",
+#                      "character",
+#                      "character")
+#
+# messy_list <- list(TRUE,
+#                    T,
+#                    "TRUE",
+#                    "T",
+#                    "01/01/01",
+#                    "01/01/2001",
+#                    NA,
+#                    NaN,
+#                    "NA",
+#                    "Na",
+#                    "na",
+#                    "10",
+#                    10,
+#                    "10.1",
+#                    10.1,
+#                    "abc",
+#                    "$%TG")
+#
+# test_that("guess_vector correctly identifies a single vector", {
[…]

### 2016-05-11 README.Rmd
updated README
removed call to `svglite` and updated some of the statuses

@@
     variant: markdown_github
+always_allow_html: yes
 ---
@@
 ```{r, echo = FALSE}
-
-library(svglite)
 
@@
 ```{r vis_dat}
+
 library(visdat)
@@
 
-At the moment `vis_guess` is slow as - take this into consideration when you are using it on data with more than 1000 rows.
-We're looking into ways of making it faster, potentially using methods from the `parallel` package, or writing some c++ code.
+At the moment `vis_guess` is very slow.
+Please take this into consideration when you are using it on data with more than 1000 rows.
+We're looking into ways of making it faster, potentially using methods from the `parallel` package, or extending the c++ code from `readr:::collectorGuess`.
 
-# Future work
+# Interactivity
+
+Thanks to Carson Sievert, you can now add some really nifty interactivity into visdat by using `plotly::ggplotly`, allowing for information to be revealed upon mouseover of a cell. For example:
+
+```{r}
+
+library(plotly)
+
+vis_dat(example2) %>% ggplotly()
+
+```
+
+# Road Map
+
+**Allowing for missings to be visualised/seen when there is a small proportion**
+
+This problem will be approached by adding in a bar at the top of the visualisation that indicates whether there are any missings in a column.
+I will also add into the legend the % of overall missing/absent.
 
@@
 
-Another idea is to pass expectations into `vis_dat` or `vis_miss`, along the lines of the `expectation` command in `assertr`.
+The idea here is to pass expectations into `vis_dat` or `vis_miss`, along the lines of the `expectation` command in `assertr`.
 For example, you could ask `vis_dat` to identify those cells with values of -1 with something like this:
@@
 
-`vis_datly`. `vis_dat` could include an interactive version of the plots using `plotly`, so that you can actually _see_ what is inside the data.
-
 # Thank you
@@
 
-Thanks also to Noam Ross for his suggestions and code for using plotly with visdat.
+Thanks also to Carson Sievert for writing the code that combined plotly with visdat, and for Noam Ross for suggesting this in the first place.

### 2016-05-11 R/vis_compare.R

@@ -1,0 +1,83 @@
+#' vis_compare
+#'
+#' \code{vis_compare} compare two dataframes and see where they are different.
+#'
+#' @description \code{vis_compare}, like the other vis_* families, gives an at-a-glance ggplot of a particular feature. In this case, it is the difference between two dataframes. This function has not been implemented yet, but this serves as a note of how it might work. It would be very similar to vis_miss, where you basically colouring cells according to "match" and "non-match". The code for this would be pretty crazy simple. `x <- 1:10; y <- c(1:5, 10:14) ;x == y` returns ` [1]  TRUE  TRUE  TRUE  TRUE  TRUE FALSE FALSE FALSE FALSE FALSE` Here black could indicate a match, and white a non match. One of the challenges with this would be cases where the datasets are different dimensions. One option could be to return as a message the columns that are not in the same dataset. Matching rows could be done by row number, and just lopping off the trailing ones and spitting out a note. Then, if the user wants, it could use an ID/key to match by.
+#'
+#' @param x a data.frame
+#'
+#' @param sort_match logical TRUE/FALSE. TRUE arranges the columns in order of most matches.
+#'
+#'
+
+# NOTE: This function is not yet complete, and is unexported as it is still mostly in the idea phase
+
+vis_compare <- function(x,
+                        sort_match = FALSE){
+  # make a TRUE/FALSE matrix of the data.
+  # This tells us whether it is missing (true) or not (false)
+  x.na <- is.na(x)
+
+  # But we would want some sort of equivalent function where it
+
+  if (sort_match == TRUE) {
+
+    # arrange by the columns with the highest missingness
+    # code inspired from https://r-forge.r-project.org/scm/viewvc.php/pkg/R/missing.pattern.plot.R?view=markup&root=mi-dev
+    # get the order of columns with highest missingness
+    na_sort <- order(colSums(is.na(x)), decreasing = TRUE)
+
+    # get the names of those columns
+    col_order_index <- names(x)[na_sort]
+
+    # original code was a bit slower:
+    #
+    # col_order_index <-
+    #   x.na %>%
+    #   as.data.frame %>%
+    #   dplyr::summarise_each(funs(mean)) %>%
+    #   names
+
+  } else {
+
+    col_order_index <- names(x)
+
+  }
+
+  # Arranged data by dendrogram order index
+
+  d <- x.na %>%
+    as.data.frame %>%
+    mutate(rows = row_number()) %>%
+    # gather the variables together for plotting
[…]

### 2016-05-11 README.Rmd
update README
Don’t evaluate the plotly commands
add travis and appveyor and badges

@@
     variant: markdown_github
-always_allow_html: yes
 ---
@@
 # visdat
+
+<!-- add a TravisCI badge -->
+[![Travis-CI Build Status](https://travis-ci.org/tierneyn/visdat.svg?branch=master)](https://travis-ci.org/tierneyn/visdat)
+
+<!-- Add an appVeyor badge -->
+
+[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/tierneyn/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/tierneyn/visdat)
 
@@
 
-Thanks to Carson Sievert, you can now add some really nifty interactivity into visdat by using `plotly::ggplotly`, allowing for information to be revealed upon mouseover of a cell. For example:
+Thanks to Carson Sievert, you can now add some really nifty interactivity into visdat by using `plotly::ggplotly`, allowing for information to be revealed upon mouseover of a cell.
+The code to do this can be seen below, but is not shown as the github README doesn't support HTML interactive graphics...yet.
 
-```{r}
+```{r eval = FALSE}
 

### 2016-05-11 README.Rmd
adjust read badges
this should put the badges onto one line

--- 2016-05-11 README.Rmd
@@
 # visdat
+
+<!-- add a TravisCI badge -->
+<!-- Add an appVeyor badge -->
+[![Travis-CI Build Status](https://travis-ci.org/tierneyn/visdat.svg?branch=master)](https://travis-ci.org/tierneyn/visdat) [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/tierneyn/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/tierneyn/visdat)
 

### 2016-05-12 README.Rmd
update README to with updated username for travis

--- 2016-05-11 README.Rmd
@@
 # visdat
+
+<!-- add a TravisCI badge -->
+<!-- Add an appVeyor badge -->
+[![Travis-CI Build Status](https://travis-ci.org/njtierney/visdat.svg?branch=master)](https://travis-ci.org/njtierney/visdat) [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/njtierney/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/njtierney/visdat)
 

### 2016-05-12 R/vis_miss.R
add a default % of complete/missing to legend

@@ -10,10 +10,13 @@
 #'
 #' @param sort_miss logical TRUE/FALSE. TRUE arranges the columns in order of missingness
 #'
+#' @param show_perc logical TRUE/FALSE. TRUE now adds in the % of missing/complete data in the whole dataset into the legend. Default value is TRUE
+#'
 #' @export
 vis_miss <- function(x,
                      cluster = FALSE,
-                     sort_miss = FALSE){
+                     sort_miss = FALSE,
+                     show_perc = TRUE){
   # make a TRUE/FALSE matrix of the data.
   # This tells us whether it is missing (true) or not (false)
   x.na <- is.na(x)
@@ -70,6 +73,39 @@ vis_miss <- function(x,
 
   d$value <- tidyr::gather_(x, "variables", "value", names(x))$value
 
+  # calculate the overall % missingness to display in legend ------------
+
+  if (show_perc == TRUE){
+
+    x <- example2
+    p_miss <- (mean(is.na(x)) * 100)
+    p_miss <- round(p_miss, 1)
+
+    # calculate % present
+    p_pres <- 100-p_miss
+
+    # create the labels
+    p_miss_lab <- paste("Missing (",
+                        p_miss,
+                        "%)",
+                        sep = "")
+
+    p_pres_lab <- paste("Present (",
+                        p_pres,
+                        "%)",
+                        sep = "")
+
+  } else {
+
+    p_miss_lab <- "Missing"
+
+    p_pres_lab <- "Present"
+
+  }
+
+
+  # note: There should be some sort fo really easy test to do here
+
     # then we plot it
     ggplot(data = d,
            aes_string(x = "variables",
@@ -79,8 +115,8 @@ vis_miss <- function(x,
     geom_raster(aes_string(fill = "valueType")) +
     # change the colour, so that missing is grey, present is black
     scale_fill_grey(name = "",
-                    labels = c("Present",
-                               "Missing")) +
+                    labels = c(p_pres_lab,
+                               p_miss_lab)) +
     theme_minimal() +
     theme(axis.text.x = element_text(angle = 45,
                                      vjust = 1,

### 2016-05-12 README.Rmd
update NEWS and fix appveyor in README

--- 2016-05-11 README.Rmd
@@
 <!-- add a TravisCI badge -->
-[![Travis-CI Build Status](https://travis-ci.org/tierneyn/visdat.svg?branch=master)](https://travis-ci.org/tierneyn/visdat)
-
 <!-- Add an appVeyor badge -->
-
-[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/tierneyn/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/tierneyn/visdat)
+[![Travis-CI Build Status](https://travis-ci.org/njtierney/visdat.svg?branch=master)](https://travis-ci.org/njtierney/visdat) [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/tierneyn/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/tierneyn/visdat)
 

### 2016-05-12 README.Rmd
update badges

--- 2016-05-11 README.Rmd
@@
 <!-- Add an appVeyor badge -->
-[![Travis-CI Build Status](https://travis-ci.org/tierneyn/visdat.svg?branch=master)](https://travis-ci.org/tierneyn/visdat) [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/tierneyn/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/tierneyn/visdat)
+[![Travis-CI Build Status](https://travis-ci.org/njtierney/visdat.svg?branch=master)](https://travis-ci.org/njtierney/visdat) [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/tierneyn/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/njtierney/visdat)
 

### 2016-05-12 R/vis_guess.R

@@ -7,6 +7,7 @@
 #' @param x a data.frame object
 #'
 #' @examples
+#' \dontrun{
 #' library(dplyr)
 #' library(tidyr)
 #' library(visdat)
@@ -20,7 +21,7 @@
 #'                            no = capacity))
 #'
 #' vis_guess(dat_messy)
-#'
+#'}
 #' @export
 vis_guess <- function(x){
 

### 2016-05-29 README.Rmd
minor updates. see news.md

@@
 
-There are currently two main commands: `vis_dat` and `vis_miss`.
+There are currently three main commands: `vis_dat`, `vis_miss`, and `vis_guess`
 
@@
 
-- **new!!** `vis_guess` has a guess at what the value of each cell.
-So "10.1" will return "double", and `10.1` will return "double", and 01/01/01 will return "date".
+-  `vis_guess` has a guess at what the value of each cell.
+So "10.1" will return "double", and `10.1` will return "double", and `01/01/01` will return "date".
 Keep in mind that it is a **guess** at what each cell is, so you can't trust this fully.
@@
 
-# Example
+# Examples
+
+## Using `vis_dat`
 
@@
 
+## using `vis_miss`
+
 We can explore the missing data further using `vis_miss`
@@
 ```
+
+The percentages of missing/complete in `vis_miss` are accurate to 1 decimal place.
 
@@
 ```
+
+When there is <0.1% of missingness, `vis_miss` indicates that there is >1% missingness.
+
+```{r}
+
+test_miss_df <- data.frame(x1 = 1:10000,
+                           x2 = rep("A", 10000),
+                           x3 = c(rep(1L, 9999), NA))
+
+vis_miss(test_miss_df)
+
+```
+
+`vis_miss` will also indicate when there is no missing data at all
+
+```{r}
+
+vis_miss(mtcars)
+
+```
+
+## using `vis_guess`
 
@@
 
-**Allowing for missings to be visualised/seen when there is a small proportion**
-
-This problem will be approached by adding in a bar at the top of the visualisation that indicates whether there are any missings in a column.
-I will also add into the legend the % of overall missing/absent.
-
 **visualising expectations**
@@
 
-# Thank you
+# Thank yous
 

### 2016-05-29 R/miss_guide_label.R

@@ -1,0 +1,49 @@
+#' miss_guide_label
+#'
+#' \code{miss_guide_label} is an internal function for vis_miss to label the legend.
+#'
+#' @param x is a dataframe passed from vis_miss(x).
+#'
+#' @return a data_frame with two columns \code{p_miss_lab} and \code{p_pres_lab}, containing the labels to use for present and missing. A dataframe is returned because I think it is a good style habit compared to a list.
+#'
+miss_guide_label <- function(x) {
+
+    p_miss <- (mean(is.na(x)) * 100)
+
+    if (p_miss == 0) {
+
+      p_miss_lab <- "No Missing Values"
+
+      p_pres_lab <- "Present (100%)"
+
+    } else if (p_miss < 0.1) {
+
+      p_miss_lab <- "Missing (< 0.1%)"
+
+      p_pres_lab <- "Present (> 99.9%)"
+
+    } else {
+
+      # calculate rounded percentages
+      p_miss <- round(p_miss, 1)
+      p_pres <- 100 - p_miss
+
+    # create the labels
+    p_miss_lab <- paste("Missing (",
+                        p_miss,
+                        "%)",
+                        sep = "")
+
+    p_pres_lab <- paste("Present (",
+                        p_pres,
+                        "%)",
+                        sep = "")
+    }
+
+    label_frame <- dplyr::data_frame(p_miss_lab = paste(p_miss_lab),
+                                     p_pres_lab = paste(p_pres_lab))
+
+    return(label_frame)
+
+
+}

### 2016-05-29 R/vis_dat.R

@@ -51,5 +51,6 @@ vis_dat <- function(x,
                                      hjust = 1)) +
     labs(x = "Variables in Dataset",
          y = "Observations") +
-    scale_x_discrete(limits = type_order_index)
+    scale_x_discrete(limits = type_order_index) +
+    guides(fill = guide_legend(title = "Type"))
 }

### 2016-05-29 R/vis_guess.R

@@ -41,13 +41,14 @@ vis_guess <- function(x){
                       y = "rows",
                       # text assist with plotly mouseover
                       text = "value")) +
-    geom_raster(aes_string(fill = "guess")) +
-    theme_minimal() +
-    theme(axis.text.x = element_text(angle = 45,
-                                     vjust = 1,
-                                     hjust = 1)) +
-    labs(x = "Variables in Dataset",
-         y = "Observations") +
-    scale_x_discrete(limits = names(x))
+      geom_raster(aes_string(fill = "guess")) +
+      theme_minimal() +
+      theme(axis.text.x = element_text(angle = 45,
+                                       vjust = 1,
+                                       hjust = 1)) +
+      labs(x = "Variables in Dataset",
+           y = "Observations") +
+      scale_x_discrete(limits = names(x))  +
+      guides(fill = guide_legend(title = "Type"))
 
 } #end function

### 2016-05-29 R/vis_miss.R

@@ -76,24 +76,31 @@ vis_miss <- function(x,
 
   if (show_perc == TRUE){
 
-    x <- example2
-    p_miss <- (mean(is.na(x)) * 100)
-    p_miss <- round(p_miss, 1)
+    temp <- miss_guide_label(x)
 
-    # calculate % present
-    p_pres <- 100-p_miss
+    p_miss_lab <- temp$p_miss_lab
 
-    # create the labels
-    p_miss_lab <- paste("Missing (",
-                        p_miss,
-                        "%)",
-                        sep = "")
+    p_pres_lab <- temp$p_pres_lab
 
-    p_pres_lab <- paste("Present (",
-                        p_pres,
-                        "%)",
-                        sep = "")
+    # # x <- example2
+    # p_miss <- (mean(is.na(x)) * 100)
+    # p_miss <- round(p_miss, 1)
+    #
+    # # calculate % present
+    # p_pres <- 100-p_miss
+    #
+    # # create the labels
+    # p_miss_lab <- paste("Missing (",
+    #                     p_miss,
+    #                     "%)",
+    #                     sep = "")
+    #
+    # p_pres_lab <- paste("Present (",
+    #                     p_pres,
+    #                     "%)",
+    #                     sep = "")
 
+    # else if show_perc FALSE
   } else {
 
     p_miss_lab <- "Missing"
@@ -102,27 +109,25 @@ vis_miss <- function(x,
 
   }
 
-
-  # note: There should be some sort fo really easy test to do here
-
     # then we plot it
     ggplot(data = d,
            aes_string(x = "variables",
                       y = "rows",
                       # text assists with plotly mouseover
                       text = "value")) +
-    geom_raster(aes_string(fill = "valueType")) +
-    # change the colour, so that missing is grey, present is black
-    scale_fill_grey(name = "",
-                    labels = c(p_pres_lab,
-                               p_miss_lab)) +
-    theme_minimal() +
-    theme(axis.text.x = element_text(angle = 45,
-                                     vjust = 1,
-                                     hjust = 1)) +
-    labs(x = "Variables in Data",
-         y = "Observations") +
-  scale_x_discrete(limits = col_order_index)
+      geom_raster(aes_string(fill = "valueType")) +
+      # change the colour, so that missing is grey, present is black
+      scale_fill_grey(name = "",
+                      labels = c(p_pres_lab,
+                                 p_miss_lab)) +
+      theme_minimal() +
[…]

### 2016-05-29 tests/testthat/test-vismiss-legend.R

@@ -1,0 +1,49 @@
+## testing for miss_guide_label
+context("vis_miss Legend")
+
+
+test_that("miss_guide_label works for exactly 0.1% missing", {
+
+  # 1 missing in 1000.
+  test_miss_1 <- c(1:999, NA)
+
+  expect_equal(miss_guide_label(test_miss_1)$p_miss_lab, "Missing (0.1%)")
+  expect_equal(miss_guide_label(test_miss_1)$p_pres_lab, "Present (99.9%)")
+
+})
+
+test_that("miss_guide_label works for < 0.1% missing", {
+
+  # 1 missing in 10,000. This should produce Missing < 0.1
+  test_miss_2 <- c(1:10000, NA)
+
+  expect_equal(miss_guide_label(test_miss_2)$p_miss_lab, "Missing (< 0.1%)")
+  expect_equal(miss_guide_label(test_miss_2)$p_pres_lab, "Present (> 99.9%)")
+
+})
+
+test_that("miss_guide_label works for no missing", {
+
+  # no missings
+  test_miss_3 <- c(1:10)
+
+  expect_equal(miss_guide_label(test_miss_3)$p_miss_lab, "No Missing Values")
+  expect_equal(miss_guide_label(test_miss_3)$p_pres_lab, "Present (100%)")
+
+})
+
+test_that("miss_guide_label works for some missings", {
+
+  #
+  test_miss_4 <- c(1:10,NA)
+
+  expect_equal(
+    tidyr::extract_numeric(miss_guide_label(test_miss_4)$p_miss_lab),
+    round(mean(is.na(test_miss_4))*100, 1)
+    )
+  expect_equal(
+    tidyr::extract_numeric(miss_guide_label(test_miss_4)$p_pres_lab),
+    100 - round(mean(is.na(test_miss_4))*100, 1)
+    )
+
+})

### 2016-08-17 R/vis_dat.R
fix typo
forgot to remove testing data

@@ -22,7 +22,7 @@ vis_dat <- function(x,
 
     type_sort <- order(
       # get the class, if there are multiple classes, combine them together
-      purrr::map_chr(.x = typical_larger_data,
+      purrr::map_chr(.x = x,
                      .f = function(x) paste(class(x), collapse = "\n"))
       )
     # get the names of those columns

### 2016-08-18 README.Rmd
fix up ready
fix appveyor badge, update plots

@@
 <!-- Add an appVeyor badge -->
-[![Travis-CI Build Status](https://travis-ci.org/njtierney/visdat.svg?branch=master)](https://travis-ci.org/njtierney/visdat) [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/tierneyn/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/njtierney/visdat)
+[![Travis-CI Build Status](https://travis-ci.org/njtierney/visdat.svg?branch=master)](https://travis-ci.org/njtierney/visdat) [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/njtierney/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/njtierney/visdat)
 
@@
 You can turn this off by setting `sort_type == FALSE`.
-This feature is better illustrated using the `example2` dataset, borrowed from csv-fingerprint.
+This feature is better illustrated using the `airquality` dataset from base R.
 
@@
 
-vis_dat(example2)
+vis_dat(airquality)
 
-vis_dat(example2, 
+vis_dat(airquality, 
         sort_type = FALSE)
@@
 
-The plot above tells us that R reads this dataset as having numeric and integer values, along with some missing data in `Ozone` and `Solar.R`.
+The plot above tells us that R reads this dataset as having characters, and integer values, along with some missing data in `Ozone` and `Solar.R`.
+
+### with different data
+
+To demonstrate what visdat looks like when you have different kinds of data, we can look at the `typical_data` from within visdat, created using the excellent [`wakefield package`](https://github.com/trinker/wakefield).
+
+```{r}
+
+vis_dat(typical_data)
+
+```
+
+We can also look into using even wider data, looking at `typical_larger_data`
+
+```{r}
+
+vis_dat(typical_larger_data)
+
+```
 
@@
 vis_miss(mtcars)
+
+```
+
+## using `vis_compare`
+
+Sometimes you want to see what has changed in your data. `vis_compare` helps with that. It is currently only just barely working, so keep in mind that this is very much in its beta stages.
+
+For the sake of simplicity, lets make some changes to `iris`, and compare this new dataset
+
+```{r}
+
+iris_diff <- iris
+iris_diff[sample(1:150, 30),sample(1:4, 2)] <- NA
+
+vis_compare(iris_diff, iris)
+
+```
+
+Here the differences are marked in red.
+
+If you try and compare differences when the dimensions are different, you get an ugly error.
+
+```{r}
+
+iris_diff_2 <- iris
+iris_diff_2$new_col <- iris$Sepal.Length + iris$Sepal.Width
+
+# vis_compare(iris, iris_diff_2)
 
@@
 
-vis_dat(example2) %>% ggplotly()
+vis_dat(airquality) %>% ggplotly()
 

### 2016-08-19 R/vis_miss_ly.R
add vis_miss_ly
still under development, but a start

@@ -1,0 +1,36 @@
+#' vis_miss_ly
+#'
+#' basic implementation of vis_miss in plotly. Still need to fix the legend. this is handy when you want to interactively explore your data.
+#'
+#' @param x a dataframe
+#'
+#' @return a plotly interactive plot.
+#' @export
+#'
+#' @examples
+#'
+#' library(visdat)
+#' vis_miss_ly(airquality)
+#'
+vis_miss_ly <- function(x){
+
+  x.na <- is.na(x)
+  d <- x.na
+  n <- nrow(x)
+
+  categories <- sort(unique(c(d)))
+  m <- matrix(match(d, categories), nrow = n)
+  txt <- matrix(paste(as.matrix(x), "<br />", d), nrow = n)
+
+  plotly::plot_ly(z = m,
+          x = names(x),
+          text = txt,
+          hoverinfo = "text",
+          type = "heatmap",
+          colorbar = list(ticktext = categories,
+                          tickvals = seq_along(categories))
+  ) %>% plotly::layout(xaxis = list(title = ""))
+
+}
+
+

### 2016-11-09 README.Rmd
add vis_miss_ly
still under development, but a start
rename chunks and fix plots/figures in README

@@
 
-# What does visdat do?
-
-Initially inspired by [`csv-fingerprint`](https://github.com/setosa/csv-fingerprint), `vis_dat` helps you visualise a dataframe and "get a look at the data" by displaying the variable classes in a dataframe as a plot with `vis_dat`, and getting a brief look into missing data patterns `vis_miss`.
-
-The name `visdat` was chosen as I think in the future it could be integrated with [`testdat`](https://github.com/ropensci/testdat).
-The idea being that first you visualise your data (`visdat`), then you run tests from `testdat` to fix them.
-
-There are currently three main commands: `vis_dat`, `vis_miss`, and `vis_guess`
-
-- `vis_dat` visualises a dataframe showing you what the classes of the columns are, and also displaying the missing data.
-
-- `vis_miss` visualises just the missing data, and allows for missingness to be clustered and columns rearranged.
-`vis_miss` is similar to `missing.pattern.plot` from the `mi` package.
-Unfortunately `missing.pattern.plot` is no longer in the `mi` package (well, as of 14/02/2016).
-
--  `vis_guess` has a guess at what the value of each cell.
-So "10.1" will return "double", and `10.1` will return "double", and `01/01/01` will return "date".
-Keep in mind that it is a **guess** at what each cell is, so you can't trust this fully.
-`vis_guess` is made possible thanks to Hadley Wickham's `readr` package - thanks mate!
-
 # How to install
@@
 
+# What does visdat do?
+
+Initially inspired by [`csv-fingerprint`](https://github.com/setosa/csv-fingerprint), `vis_dat` helps you visualise a dataframe and "get a look at the data" by displaying the variable classes in a dataframe as a plot with `vis_dat`, and getting a brief look into missing data patterns `vis_miss`.
+
+The name `visdat` was chosen as I think in the future it could be integrated with [`testdat`](https://github.com/ropensci/testdat).
+The idea being that first you visualise your data (`visdat`), then you run tests from `testdat` to fix them.
+
+There are two main commands in the `visdat` package: 
+- `vis_dat()` 
+- `vis_miss()`
+
+- `vis_dat()` visualises a dataframe showing you what the classes of the columns are, and also displaying the missing data.
+
+- `vis_miss()` visualises just the missing data, and allows for missingness to be clustered and columns rearranged.
+`vis_miss()` is similar to `missing.pattern.plot` from the `mi` package.
+Unfortunately `missing.pattern.plot` is no longer in the `mi` package (well, as of 14/02/2016).
+
+There are two some experimental functions:
+
+- `vis_guess()`
+- `vis_compare()`
+
+-  `vis_guess()` has a guess at what the value of each cell, using `readr::parse_guess`.
+This means that "10.1" will return "double", and `10.1` will return "double", and `01/01/01` will return "date".
+Keep in mind that it is a **guess** at what each cell is, so you can't trust this fully.
+
+- `vis_compare()` compares two dataframes, displaying the differences.
+
 # Examples
 
-## Using `vis_dat`
+## Using `vis_dat()`
 
@@
 
-```{r vis_dat}
+```{r vis-dat-aq}
 
@@
 
-```{r}
-
-vis_dat(airquality)
+```{r vis-dat-aq-sort-type}
 
@@
 
-```{r}
+```{r vis-dat-typical-data}
 
@@
 
-```{r}
+```{r vis-dat-typical-larger-data}
 
@@
 
-## using `vis_miss`
+## using `vis_miss()`
 
-We can explore the missing data further using `vis_miss`
+We can explore the missing data further using `vis_miss()`
 
-```{r vis_miss}
+```{r vis-miss-aq}
 
@@
 
-```{r vis_miss-cluster}
+```{r vis-miss-aq-cluster}
 
@@
 
-```{r}
+```{r vis-miss-aq-sort-miss}
 
@@
 
-```{r}
+```{r vis-miss-test}
 
@@
 
-```{r}
+```{r vis-miss-mtcars}
 
@@
 
-## using `vis_compare`
+## Experimental features
 
-Sometimes you want to see what has changed in your data. `vis_compare` helps with that. It is currently only just barely working, so keep in mind that this is very much in its beta stages.
+## using `vis_compare()`
+
+Sometimes you want to see what has changed in your data. `vis_compare()` helps with that. It is currently only just barely working, so keep in mind that this is very much in its beta stages.
 
@@
 
-```{r}
+```{r vis-compare-iris}
 
@@
 
-```{r}
+```{r vis-compare-error, eval = FALSE}
 
@@
 
-# vis_compare(iris, iris_diff_2)
-
+vis_compare(iris, iris_diff_2)
+#> vis_compare is still in BETA!
+If you have suggestions or errors, post an issue at https://github.com/njtierney/visdat/issuesthe condition has length > 1 and only the first element will be usedError: `.x` (5) and `.y` (6) are different lengths
 ```
 
-## using `vis_guess`
+## using `vis_guess()`
 
-`vis_guess` takes a guess at what each cell is. It's best illustrated using some messy data, which we'll make here.
+`vis_guess()` takes a guess at what each cell is. It's best illustrated using some messy data, which we'll make here.
 
-```{r}
+```{r create-messy-vec}
 
@@
 
+set.seed(1114)
 messy_df <- data.frame(var1 = messy_vector,
@@
 
-```{r}
+```{r vis-guess-messy-df}
 
@@
 
-```{r}
+```{r visdat-messy-df}
 
@@
 vis_dat(airquality) %>% ggplotly()
+
+```
+
+## vis_dat_ly
+
+This is still under development, but it is basically a faster version of doing a ggplot and then calling ggplotly.
+
+## vis_miss_ly
+
+This is also just brand new under development, but should prove interesting once I fix up the legend, etc.
+
+```{r eval = FALSE, echo = TRUE}
+
+vis_miss_ly(airquality)
 

### 2016-11-09 R/vis_dat.R
Add examples to doco

@@ -8,6 +8,16 @@
 #'
 #' @param palette character "default", "qual" or "cb_safe". "default" (the default) provides the stock ggplot scale for separating the colours. "qual" uses an experimental qualitative colour scheme for providing distinct colours for each Type. "cb_safe" is a set of colours that are appropriate for those with colourblindness. "qual" and "cb_safe" are drawn from http://colorbrewer2.org/.
 #'
+#'
+#' @examples
+#'
+#' library(visdat)
+#'
+#' vis_dat(airquality)
+#'
+#' # experimental colourblind safe pallete
+#' vis_dat(airquality, palette = "cb_safe")
+#'
 #' @export
 vis_dat <- function(x,
                     sort_type = TRUE,

### 2016-11-09 R/vis_miss.R

@@ -10,6 +10,15 @@
 #'
 #' @param show_perc logical TRUE/FALSE. TRUE now adds in the \% of missing/complete data in the whole dataset into the legend. Default value is TRUE
 #'
+#' @examples
+#'
+#' library(visdat)
+#'
+#' vis_miss(airquality)
+#'
+#' vis_miss(airquality, cluster = TRUE)
+#'
+#' vis_miss(airquality, sort_miss = TRUE)
 #'
 #' @export
 vis_miss <- function(x,

### 2016-11-09 R/vis_compare.R
Add examples; make it clear that these are in beta phase

@@ -16,7 +16,6 @@
 #' library(visdat)
 #'
 #' vis_compare(iris, iris_diff)
-#'
 #' @export
 
 vis_compare <- function(df1,

### 2016-11-09 R/vis_guess.R

@@ -5,24 +5,39 @@
 #' @param x a data.frame object
 #'
 #' @examples
-#' \dontrun{
+#'
 #' library(tidyr)
 #' library(readr) # to load the `readr:::collectorGuess` function
 #' library(dplyr)
 #' library(visdat)
 #'
-#' vis_guess(example2)
+#' messy_vector <- c(TRUE,
+#'                  "TRUE",
+#'                  "T",
+#'                  "01/01/01",
+#'                  "01/01/2001",
+#'                  NA,
+#'                  NaN,
+#'                  "NA",
+#'                  "Na",
+#'                  "na",
+#'                  "10",
+#'                  10,
+#'                  "10.1",
+#'                  10.1,
+#'                  "abc",
+#'                  "$%TG")
+#' set.seed(1114)
+#' messy_df <- data.frame(var1 = messy_vector,
+#'                        var2 = sample(messy_vector),
+#'                        var3 = sample(messy_vector))
+#' vis_guess(messy_df)
 #'
-#' dat_messy <-
-#'   example2 %>%
-#'   mutate(capacity = ifelse(capacity <= 60500,
-#'                            yes = "2010/10/10",
-#'                            no = capacity))
-#'
-#' vis_guess(dat_messy)
-#'}
 #' @export
 vis_guess <- function(x){
+
+  warning("vis_guess is still in BETA! If you have suggestions or errors, post an issue at https://github.com/njtierney/visdat/issues")
+
 
   d <- x %>%
     mutate(rows = row_number()) %>%

### 2016-12-15 R/vis_miss.R
change colour so missing is black, and present is white
This is to draw attention to the missings, they are the focus.

@@ -27,6 +27,7 @@ vis_miss <- function(x,
                      show_perc = TRUE){
   # make a TRUE/FALSE matrix of the data.
   # This tells us whether it is missing (true) or not (false)
+  x = airquality
   x.na <- is.na(x)
 
   # switch for creating the missing clustering
@@ -92,24 +93,6 @@ vis_miss <- function(x,
 
     p_pres_lab <- temp$p_pres_lab
 
-    # # x <- example2
-    # p_miss <- (mean(is.na(x)) * 100)
-    # p_miss <- round(p_miss, 1)
-    #
-    # # calculate % present
-    # p_pres <- 100-p_miss
-    #
-    # # create the labels
-    # p_miss_lab <- paste("Missing (",
-    #                     p_miss,
-    #                     "%)",
-    #                     sep = "")
-    #
-    # p_pres_lab <- paste("Present (",
-    #                     p_pres,
-    #                     "%)",
-    #                     sep = "")
-
     # else if show_perc FALSE
   } else {
 
@@ -127,16 +110,24 @@ vis_miss <- function(x,
                       text = "value")) +
       geom_raster(aes_string(fill = "valueType")) +
       # change the colour, so that missing is grey, present is black
-      scale_fill_grey(name = "",
-                      labels = c(p_pres_lab,
-                                 p_miss_lab)) +
+      # 2016/12/15: Change the colour AGAIN, so that missing is black
+      #
+      scale_fill_manual(name = "",
+                        values = c("grey80",
+                                   "grey20"),
+                        labels = c(p_pres_lab,
+                                     p_miss_lab)) +
+      # scale_fill_grey(name = "",
+      #                 labels = c(p_pres_lab,
+      #                            p_miss_lab)) +
       theme_minimal() +
       theme(axis.text.x = element_text(angle = 45,
                                        vjust = 1,
                                        hjust = 1)) +
       labs(x = "Variables in Data",
            y = "Observations") +
-      scale_x_discrete(limits = col_order_index)
+      scale_x_discrete(limits = col_order_index) +
+      guides(fill = guide_legend(reverse = TRUE))
       # guides(fill = guide_legend(title = "Type"))
   # Thanks to http://www.markhneedham.com/blog/2015/02/27/rggplot-controlling-x-axis-order/
   # For the tip on using scale_x_discrete

### 2016-12-28 tests/testthat/test_variable_spaces.R
minor update to broken tests...

@@ -1,11 +1,13 @@
-#
-# messy_names <- data_frame(`Sepal Width` = iris$Sepal.Width,
-#                           `Sepal Length` = iris$Sepal.Length,
-#                           `Petal Length` = iris$Petal.Length,
-#                           `Species Names` = iris$Species)
-#
 # context("Test names with spaces")
 #
+# library(visdat)
+#
+# messy_names <- dplyr::data_frame(`Sepal Width` = iris$Sepal.Width,
+#                                  `Sepal Length` = iris$Sepal.Length,
+#                                  `Petal Length` = iris$Petal.Length,
+#                                  `Species Names` = iris$Species)
+#
+#
 # test_that("vis_dat works on dataframes with irregular variable names", {
-#   expect_ (vis_dat(messy_names))
+#   expect_success(vis_dat(messy_names))
 # })

### 2016-12-29 README.Rmd
update install instructions
thank Ivan Hanigan for his introduction to csv-fingerprint

@@
 
-library(devtools)
-
-install_github("tierneyn/visdat")
+devtools::install_github("njtierney/visdat")
 ```
@@
 
-## Experimental features
+# Experimental features
 
@@
 
-Thank you to Jenny Bryan, whose [tweet](https://twitter.com/JennyBryan/status/679011378414268416) got me thinking about vis_dat, and for her code contributions that remove a lot of testing errors.
+Thank you to Ivan Hanigan who [first commented](http://www.njtierney.com/r/missing%20data/rbloggers/2015/12/01/ggplot-missing-data/#comment-2388263747) this suggestion after I made a blog post about an initial prototype `ggplot_missing`, and Jenny Bryan, whose [tweet](https://twitter.com/JennyBryan/status/679011378414268416) got me thinking about vis_dat, and for her code contributions that removed a lot of errors.
 

### 2016-12-29 tests/testthat/test_variable_spaces.R
update still broken test

@@ -9,5 +9,5 @@
 #
 #
 # test_that("vis_dat works on dataframes with irregular variable names", {
-#   expect_success(vis_dat(messy_names))
+#   expect_error(vis_dat(messy_names))
 # })

### 2017-01-04 README.Rmd
add coverage badge to README

@@
 <!-- Add an appVeyor badge -->
-[![Travis-CI Build Status](https://travis-ci.org/njtierney/visdat.svg?branch=master)](https://travis-ci.org/njtierney/visdat) [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/njtierney/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/njtierney/visdat)
+[![Travis-CI Build Status](https://travis-ci.org/njtierney/visdat.svg?branch=master)](https://travis-ci.org/njtierney/visdat) [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/njtierney/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/njtierney/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/njtierney/visdat/master.svg)](https://codecov.io/github/njtierney/visdat?branch=master)
 

### 2017-01-07 paper/paper.Rmd
add paper folder for JOSS

---
title: 'visdat: Visualising whole data frames'
tags:
  - visualisation
  - R
  - exporatory data analysis
authors:
 - name: Nicholas Tierney
   orcid: 0000-0003-1460-8722
   affiliation: 1
affiliations:
 - name: Queensland University of Technology
   index: 1
date: 06 January 2017
bibliography: paper.bib
---

# Summary

In his [R for Data Science book](http://r4ds.had.co.nz/), Wickham describes 6 phases of data science:

![Process of Data Science](http://r4ds.had.co.nz/diagrams/data-science.png)

You can get insight into your data by modelling, visualising, and transforming, which Wickham describes as "Understanding" or "knowledge generation". There is some overlap here, as in this process it wouldn't be surprising that you might uncover some feature of your dataset that you would need to clean up - you might discover that some strings were indeed factors, or that gender was considered numeric in your regression, warranting you to re-visit the Tidying phase.

In the same way, I think that this process gets applied in the Tidying step of data science. You read in your data, but you then need to look at it to understand what you need to do to "fix it". And that's a key phrase here "looking at the data" - what does that mean?

On one hand, you can look at the head of the data:

```{r head-iris}

head(iris)

```

Or you can have a `glimpse` at it through `dplyr::glimpse`

```{r glimpse}

dplyr::glimpse(iris)

```

Here we see that we have doubles, and a factor. We get some insight into the data.

But we don't always have data like the canonical iris dataset. let's take a look at some data that might be a bit more typical of "messy" data.

```{r visdat-glimpse}

library(visdat)
dplyr::glimpse(typical_data)

```

Looking at this, you might then ask:

> Isn't it odd that Income is a factor? And Age is a character?

And you might start to wonder what else is different, what else changed?

And it might be a bit unclear where to go from there. Do you plot the data? Why does my plot look wierd? What are these other strange features in the data? The `visdat` package provides visualisations of an entire dataframe at once. Initially inspired by [`csv-fingerprint`](https://github.com/setosa/csv-fingerprint), `visdat` provides tools to create heatmap-like visualisations of an entire dataframe. `visdat` provides 2 main functions `vis_dat` and `vis_miss`.

`vis_dat()` helps explore the data class structure and missingness:

```{r load-data}

vis_dat(typical_data)

```

And the `vis_miss` function provides a custom plot for missing data.

```{r}

vis_miss(typical_data)

```

The name `visdat` was chosen as it borrows from the idea of [`testdat`](https://github.com/ropensci/testdat), which provides unit testing for your data.  In a similar way, `visdat` provides visual tests, the idea being that first you visualise your data (`visdat`), then you run tests from `testdat`, or a package like [`assertr`](https://cran.r-project.org/web/packages/assertr/index.html), to fix these errors.

# References

### 2017-01-07 README.Rmd
Clean up typos in README.Rmd.

@@
 There are two main commands in the `visdat` package:
-- `vis_dat()` 
-- `vis_miss()`
 
@@
 
-There are two some experimental functions:
-
-- `vis_guess()`
-- `vis_compare()`
+There are two experimental functions:
 
@@
 by default, `vis_dat` sorts the columns according to the type of the data in the vectors.
-You can turn this off by setting `sort_type == FALSE`.
+You can turn this off by setting `sort_type = FALSE`.
 This feature is better illustrated using the `airquality` dataset from base R.
@@
 
-The plot above tells us that R reads this dataset as having characters, and integer values, along with some missing data in `Ozone` and `Solar.R`.
+The plot above tells us that R reads this dataset as having characters and integer values, along with some missing data in `Ozone` and `Solar.R`.
 
-### with different data
+### With many kinds of data
 
-To demonstrate what visdat looks like when you have different kinds of data, we can look at the `typical_data` from within visdat, created using the excellent [`wakefield package`](https://github.com/trinker/wakefield).
+To demonstrate what visdat looks like when you have different kinds of data, we can use `visdat` to look at the dataset `typical_data`, which is from the excellent [`wakefield package`](https://github.com/trinker/wakefield).
 
@@
 
-## using `vis_miss()`
+## Using `vis_miss()`
 
@@
 
-## using `vis_compare()`
+## Using `vis_compare()`
 
@@
 
-Here the differences are marked in red.
+Here the differences are marked in blue.
 
@@
 
-## using `vis_guess()`
+## Using `vis_guess()`
 
@@
 
-Where you'd just assume your data is wierd because it's all factors - or worse, not notice that this is a problem.
+Where you'd just assume your data is weird because it's all factors - or worse, not notice that this is a problem.
 
@@
 
-**visualising expectations**
+**Visualising expectations**
 
@@
 
-Thank you to Ivan Hanigan who [first commented](http://www.njtierney.com/r/missing%20data/rbloggers/2015/12/01/ggplot-missing-data/#comment-2388263747) this suggestion after I made a blog post about an initial prototype `ggplot_missing`, and Jenny Bryan, whose [tweet](https://twitter.com/JennyBryan/status/679011378414268416) got me thinking about vis_dat, and for her code contributions that removed a lot of errors.
+Thank you to Ivan Hanigan who [first commented](http://www.njtierney.com/r/missing%20data/rbloggers/2015/12/01/ggplot-missing-data/#comment-2388263747) this suggestion after I made a blog post about an initial prototype `ggplot_missing`, and Jenny Bryan, whose [tweet](https://twitter.com/JennyBryan/status/679011378414268416) got me thinking about `vis_dat`, and for her code contributions that removed a lot of errors.
 
@@
 
-Thank you to Miles McBain for his suggestions on how to improve vis guess. This resulted in making it at least 2-3 times faster.
+Thank you to Miles McBain for his suggestions on how to improve `vis_guess`. This resulted in making it at least 2-3 times faster.
 
-Thanks also to Carson Sievert for writing the code that combined plotly with visdat, and for Noam Ross for suggesting this in the first place.
+Thanks also to Carson Sievert for writing the code that combined `plotly` with `visdat`, and for Noam Ross for suggesting this in the first place.

### 2017-01-08 README.Rmd
update README

--- 2017-01-04 README.Rmd
@@
 There are two main commands in the `visdat` package:
+
 - `vis_dat()` 
@@
 Please take this into consideration when you are using it on data with more than 1000 rows.
-We're looking into ways of making it faster, potentially using methods from the `parallel` package, or extending the c++ code from `readr:::collectorGuess`.
+We're looking into ways of making it faster, potentially using methods from the `parallel` package, or extending the c++ code from `readr::parse_guess`.
 
@@
 
-This is also just brand new under development, but should prove interesting once I fix up the legend, etc.
+This is also under development, and still needs some more work on the legend, etc.
 
-```{r eval = FALSE, echo = TRUE}
+```{r vis-miss-ly}
 

### 2017-01-08 paper/paper.Rmd
update the paper, reduce excess content

@@
    index: 1
-date: 06 January 2017
-bibliography: paper.bib
+date: 09 January 2017
+# bibliography: paper.bib
+output: 
+  html_document: 
+    keep_md: yes
 ---
@@
 
-In his [R for Data Science book](http://r4ds.had.co.nz/), Wickham describes 6 phases of data science:
+As described by Hadley Wickham, there are 6 phases of data science:
 
@@
 You can get insight into your data by modelling, visualising, and transforming, which Wickham describes as "Understanding" or "knowledge generation".
-There is some overlap here, as in this process it wouldn't be surprising that you might uncover some feature of your dataset that you would need to clean up - you might discover that some strings were indeed factors, or that gender was considered numeric in your regression, warranting you to re-visit the Tidying phase.
+There is some overlap here, as in this process it wouldn't be surprising that you might uncover some feature of your dataset that you would need to clean up or change.
+For example, in analysing some data, you might discover that some strings were indeed factors, or that gender was considered numeric in your regression, warranting you to re-visit the Tidying phase.
 
-In the same way, I think that this process gets applied in the Tidying step of data science.
-You read in your data, but you then need to look at it to understand what you need to do to "fix it".
-And that's a key phrase here "looking at the data" - what does that mean?
-
-On one hand, you can look at the head of the data:
+In the same way, this process gets applied in the Tidying step of data science.
+You read in your data, but you then need to look at it to understand what you need to do to make it ready for analysis.
+A phrase that often arises in this process is "looking at the data".
+It is helpful to look more closely at what this means.
+You can look at the first six rows of data, the head of the data:
 
@@
 
-Or you can have a `glimpse` at it through `dplyr::glimpse`
+Another alterantive is `dplyr::glimpse`
 
@@
 
-Here we see that we have doubles, and a factor. We get some insight into the data.
-
-But we don't always have data like the canonical iris dataset. let's take a look at some data that might be a bit more typical of "messy" data.
+This has a better printing method that gives us more information about the data, displaying doubles and a factor.
+However, we don't always have data like the canonical iris dataset.
+Let's take a look at some data that might be a bit more typical of "messy" data.
 
@@
 Initially inspired by [`csv-fingerprint`](https://github.com/setosa/csv-fingerprint), `visdat` provides tools to create heatmap-like visualisations of an entire dataframe.
-`visdat` provides 2 main functions `vis_dat` and `vis_miss`.
+`visdat` [@Tierney2017] is an R [@Rcore] package provides 2 main functions `vis_dat` and `vis_miss`.
 
@@
 
-The name `visdat` was chosen as it borrows from the idea of [`testdat`](https://github.com/ropensci/testdat), which provides unit testing for your data.
-In a similar way, `visdat` provides visual tests, the idea being that first you visualise your data (`visdat`), then you run tests from `testdat`, or a package like [`assertr`](https://cran.r-project.org/web/packages/assertr/index.html), to fix these errors.
+These functions provide useful tools to help "get a look at the data", using principled visualisation techniques.
+The plots are built using ggplot2 [@ggplot2], which provides a consistent and powerful framework for visualisations.
+Which means that users can customise and extend graphics from visdat very easily.
 

### 2017-01-08 paper/paper.Rmd
first submission of paper.md

@@
 date: 09 January 2017
-# bibliography: paper.bib
+bibliography: paper.bib
 output: 
@@
 
-As described by Hadley Wickham, there are 6 phases of data science:
+As described by Hadley Wickham [@wickham2016], there are 6 phases of data science:
 
@@
 
-Another alterantive is `dplyr::glimpse`
+Another alterantive is `glimpse`, from the `dplyr` package [@dplyr]
 
@@
 Initially inspired by [`csv-fingerprint`](https://github.com/setosa/csv-fingerprint), `visdat` provides tools to create heatmap-like visualisations of an entire dataframe.
-`visdat` [@Tierney2017] is an R [@Rcore] package provides 2 main functions `vis_dat` and `vis_miss`.
+`visdat` is an R [@Rcore] package provides 2 main functions `vis_dat` and `vis_miss`.
 
@@
 The plots are built using ggplot2 [@ggplot2], which provides a consistent and powerful framework for visualisations.
-Which means that users can customise and extend graphics from visdat very easily.
+This means that users can customise and extend graphics from visdat very easily.
 

### 2017-01-11 README.Rmd
create README-figs dir for README figs

--- 2017-01-07 README.Rmd
@@
   comment = "#>",
-  fig.path = "README-"
+  fig.path = "README-figs/README-"
 )
@@
 There are two main commands in the `visdat` package: 
+- `vis_dat()` 
+- `vis_miss()`
 

### 2017-01-15 README.Rmd

@@
 
-To demonstrate what visdat looks like when you have different kinds of data, we can use `visdat` to look at the dataset `typical_data`, which is from the excellent [`wakefield package`](https://github.com/trinker/wakefield).
+To demonstrate what visdat looks like when you have different kinds of data, we can look at the dataset `typical_data`, provided within `visdat`, and created with the excellent [`wakefield` package](https://github.com/trinker/wakefield).
 

### 2017-01-30 paper/response.Rmd
drafting the response to the review comments

# Response to reviewers

I would like to thank the reviewers for taking the time to review visdat so thoroughly. It is fantastic to hear comments such as:

>

,

>

And

>

...

Consider the above my response to the positive feedback in this review. I will now respond to suggsted changes and other constructive criticisms.

# @seaaan's Review

1. Documentation: README

> The README is very good! It gives a helpful overview of everything in the package and is clearly written, with good examples. All of the code examples worked on my computer. You might consider breaking the "experimental" parts of the package into a separate README like you do with the vignettes.

-

> Link to the vignettes from the README.

> The title of the github project is "A package to assist in visually testing a dataset", which doesn't describe the project that well. I suggest changing it to be the same title as on the DESCRIPTION file, which is effective in quickly summarizing what the package does.

-

> You don't introduce the airquality data set at first usage, but you do introduce it after the second plot. Move the introduction to first use.

> I don't understand this sentence: "When there is <0.1% of missingness, vis_miss indicates that there is >1% missingness."

- Agreed, this

2. Vignettes:

> addressing comments that relate to the vignette and README

> Make sure you have permission to reproduce the image from R4DS. The first few paragraphs of the "Using visdat" vignette are a bit unnecessary in my opinion. You could start much more simply, with something like: "When you get a new data set, you need to get a sense of what it contains and potential problems with it." and then continue with the discussion of different ways to approach this challenge (e.g. head, glimpse, ...).

- I agree with you that this is a clearer way to address this comment, and have addressed this in this commit ...

> In the "Using visdat" vignette, it says "missing data represented by black", but it shows up as gray on my computer.

- Interesting, it should be showing up as black, as this is what the [code and README reflect]() (well, grey50). Perhaps th

> I don't know why, but in the plotly section, a vis_dat() (non-interactive) plot of df_test appears between the first two interactive plots. I can't explain it, hopefully it's just a weird quirk on my computer.

- I can't seem to replicate this

> There is a "future work and experimental features" section in the "Using visdat" vignette -- I suggest transferring that to the experimental vignette and just linking to it from the "Using visdat" vignette. Perhaps move the vis_guess() and the interactive plots to the experimental features vignette as well, since they seem to be evolving.

-

> The plot in the experimental features vignette shows up small, I think you need to add knitr::opts_chunk$set(fig.width = 5, fig.height = 4) to the first chunk like you have in the other vignette.

- Agreed, this has been completed in this commit ...

# Function documentation

> Did you mean to export guess_type()?

- Yes, as I thought that it might be a useful function for users, but I can appreciate that perhaps it shoudl stay internal

> Formatting: links, code, and other formatting need to be done with .Rd syntax. For example, for code, use \code{}, not backticks. For bold, use \strong{} instead of asterisks. See guidelines here.

> Return values: The documentation doesn't include the "Value" section which is typically used to say what a function returns. I would add this. E.g. @return A \code{ggplot} object displaying the type of values in the data frame and the position of any missing values.

> ?vis_compare: The documentation for this function needs to be updated. It is more a list of ideas for how to implement the function than a description of what it does.

> ?vis_guess: First sentence of description seems to have an extra word or phrase, not sure exactly what was intended.

> ?vis_miss_ly: The reference to vis_miss in the "See Also" section should be a link. Like this: \code{\link[visdat]{vis_miss}}

> ?visdat: It's good that this page exists. However, it doesn't have any content -- add a brief overview of the package with links to the documentation page for the main functions.

# Examples

> Don't need to call library for either visdat or packages you use internally (only if you actually call a function from another package in the example code itself).

> example(vis_compare) is a good example but gives a bunch of warning messages.

> example(vis_dat): "palette" is misspelled.

> example(vis_guess): gives warning message

# Community guidelines

> There are no guidelines for how to contribute, either in the README or in a CONTRIBUTING file.

# JOSS

> The paper seems more detailed than most other JOSS papers and goes into specific functions. Additionally, does JOSS allow figures? I didn't see any in the papers I looked at or see mentions of them in the author guidelines. But I'm not an expert on JOSS, so maybe someone else can weigh in on that.

> The paper doesn't have any references in the "References" section, but does have inline references.

> If you keep it, make sure you have permission to reproduce the image from R4DS.

# Functionality

## Common issues

> There are a number of issues that apply to most or all of the functions in the package. Rather than repeat them under each function, I'll just put those comments together here.

> It is unintuitive to me to have the rows in reverse order (i.e. row 1 at the bottom) and for the columns to be clustered by type, rather than appear in the order they appear in the data frame. I think the default behavior should be for rows and columns to appear in the same order as in the input. Additionally, always putting the titles at the top of the columns makes sense to me. Including sort_type as an option might be useful, but by default it should be off. I'm not quite sure of the use case for the flip argument, but maybe there is one!

> The flip argument should be available for all of the functions if you provide it for any, for the sake of consistency. Possibly also the palette argument.

> A number of the functions emit warning messages ("Warning message: attributes are not identical across measure variables; they will be dropped"), specifically when they are called with a data frame that has >1 factor column with different levels. This arises from the call tidyr::gather_(x, "variables", "value", names(x))$value. This message is not relevant to the end user and should be suppressed.

> There is some code duplication across some of the functions, where you could replace it with a helper function. The code creating the plots is generally similar across functions. In addition, the following code blocks appear in essentially identical form in >1 function:

```{r eval = FALSE}

  dplyr::mutate(rows = seq_len(nrow(.))) %>%
  tidyr::gather_(key_col = "variables",
                 value_col = "valueType",
                 gather_cols = names(.)[-length(.)])

```

> Using a helper function for this next one in particular would allow you to suppress the warning messages just once:

```
  d$value <- tidyr::gather_(x, "variables", "value", names(x))$value
```

> The code for flipping the rows is also duplicated:

```
 if (flip == TRUE){
    suppressMessages({
      vis_dat_plot +
        ggplot2::scale_y_reverse() +
        ggplot2::scale_x_discrete(position = "top",
                                  limits = type_order_index) +
        ggplot2::theme(axis.text.x = ggplot2::element_text(hjust = 0.5)) +
        ggplot2::labs(x = "")
    })
  } else if (flip == FALSE){
    vis_dat_plot
  }

```

> vis_compare and vis_guess are indicated as being in beta, which seems also to apply to the plotly versions of the functions. The message that they emit is a helpful indication that they may change in the future. Before submitting to CRAN, however, you might consider moving those functions to the development version of the package and only uploading the functions with a stable API, then adding the beta version later once they stabilize. This is a judgment call; I think I tend towards being conservative on this issue personally.

> This is a minor issue, but was so thoroughly drummed into me by CS professors that I have to mention it. In calls like if (sort_type == TRUE), the == TRUE is redundant, so it's equivalent to just write if (sort_type). Similarly, if (x == FALSE) is equivalent to if (!x). There are a number of these in the code for various functions.

# vis_dat()

> Maximum data frame size: the heatmaps stop working on my computer with very large data sets. E.g. library(nycflights13); vis_dat(flights) hangs for a minute or so and then displays an empty grid. It works with 10000 rows, but not 100000. Similar issue with many columns. Especially because this package is designed for people to get a sense of new data sets, where users may not have realized how big the data are, I think it makes sense to prevent this issue. Possibilities: (1) max_rows = 10000 parameter, which stops if nrow(data) > max_rows and gives a descriptive error message. Then the user can increase the maximum or subset their data frame. Might also need to consider max_cols because 10000 rows and 2 cols is very different from 10000 rows and 500 cols. (2) Downsample the non-missing rows somehow and indicate to the user where omissions were made. (3) ???

> The palette argument doesn't cause color changes, as far as I can tell. I think you need to do vis_dat_plot <- vis_dat_plot + <palette>, whereas currently the value of vis_dat_plot isn't being updated in the blocks related to color palettes. Additionally, you could do ggplot2::scale_color_brewer(type = "qual", palette = "Set1") rather than explicitly writing out the color names, although the colors would be in a slightly different order than you have now.

> If the user provides a palette name, you should check that it's valid. If not, throw an error with a message. Currently, if you misspell a palette name, it just goes ahead with the default palette.

# Minor comments

> There are some lines of commented-out code: # x = airquality and # mutate_each_(funs(fingerprint), tbl_vars(.)) %>%

> Inside of the first if-block, there are some comments about arranging columns in order of missingness, but as far as I can tell, that is not done.

## vis_miss()

> It would be cool to indicate the % missing for each column individually (maybe in column labels?) like how you indicate the % missing for the data overall.

## vis_compare()

> This is a neat function! It would be cool if you could generalize it to also handle data frames of unequal dimensions, perhaps by showing which rows/columns are only in one data frame or the other. Just an idea I had though, not a requirement at all!

> Warning message: "1: In if (dim(df1) != dim(df2)) { : the condition has length > 1 and only the first element will be used". This is a consequence of calling dim(df1) != dim(df2). The dim function returns a two element vector in this case, but the != only compares the first element of each vector. (I.e. it's only comparing the number of rows). Instead, use !identical(dim(df1), dim(df2)).

> In the error message about data frames of unequal dimensions, you mean to say "vis_compare only handles dataframes of identical dimensions", not "does not handle".

## vis_guess

> vis_guess lacks some of the parameters that vis_dat has, but should have the same options (except maybe not sort_type?)

> Performance: Setting the locale once and then calling collectorGuess instead of guess_parser is about 33% faster on my computer. I think the savings is by avoiding repeatedly calling the locale function and avoiding repeatedly calling stopifnot (inside the readr code).

```
l <- readr::locale()

output[!nas] <- vapply(FUN = function(y) readr:::collectorGuess(y, locale_ = l),
                       X = x[!nas], FUN.VALUE = character(1))
If you're interested in converting to C++ for speed, I played around a little bit with implementing a vectorized version for guess_parser in readr. It's about 20X faster. Basically, I wrote a wrapper around collectorGuess that operates on a vector instead of single elements. This is faster because it does the looping in C++ rather than repeatedly transitioning from R to C++ and because it does some initial code once per vector instead of once per element. It is currently written as a modification of readr because that was more convenient, but I imagine it is possible to implement in your package as well. Let me know if you want to discuss that, but I'm definitely not an expert. To get it up and running, clone readr and then put this in readr/src/collectorGuess.cpp:
std::string collectorGuess2(const std::string& input, LocaleInfo locale) {

  // Work from strictest to most flexible
  if (canParse(input, isLogical, &locale))
    return "logical";
  if (canParse(input, isInteger, &locale))
    return "integer";
  if (canParse(input, isDouble, &locale))
    return "double";
  if (canParse(input, isNumber, &locale))
    return "number";
  if (canParse(input, isTime, &locale))
    return "time";
  if (canParse(input, isDate, &locale))
    return "date";
  if (canParse(input, isDateTime, &locale))
    return "datetime";

  // Otherwise can always parse as a character
  return "character";
}

// [[Rcpp::export]]
CharacterVector collectorGuessVector(CharacterVector input, List locale_) {
  LocaleInfo locale(locale_);
  CharacterVector output(input.size());

  if (input.size() == 0 || allMissing(input)) {
    CharacterVector result(1);
    result[0] = "character";
    return result;
  }

  for (int i = 0; i < input.size(); i++) {
    output[i] = collectorGuess2(std::string(input[i]), locale);
  }

  return output;
}

```

> Then put this in readr/R/collectors.R:

```
#' @rdname parse_guess
#' @export
guess_parser_vector <- function(x, locale = default_locale()) {
  stopifnot(is.locale(locale))
  collectorGuessVector(x, locale)
}
```

> Then rebuild readr. Back in visdat, call guess_parser_vector once with the whole vector of unknowns instead of calling guess_parser repeatedly.

## vis_miss_ly

> The pop-up window is great! Can you make it the same as for ggplotly(vis_dat())? It's especially helpful to show the row number.

> Why not have the same arguments for this function as for vis_miss?

## vis_dat_ly()

> With vis_dat(airquality) %>% ggplotly(), the window that appears on mouseover should say "type" instead of "valueType", "variable" instead of "variables" and "row" instead of "rows"

> If you provide vis_miss_ly, it makes sense to me to also provide vis_dat_ly, and possibly also vis_compare_ly and vis_guess_ly. An initial implementation could just be vis_dat(x) %>% ggplotly(), but then you have the API in place at least.

# Tests

> Cool use of vdiffr and generally good coverage. The tests don't all pass on my machine (they are mostly skipped), which I think is because I don't have the .svg reference files on my computer.

> I suggest adding a test for each function using typical_data, because that currently emits a warning message. That way once you stop the error message from occurring, you'll have a regression test for the future.

> Maybe I don't understand how vdiffr works, but I'm confused about test-vis-dat.R where in the calls to vdiffr you always provide vis_dat_plot, never any of the other plots you created. Additionally, both of the plots you create using non-default palettes are given the same name, so you're overwriting one of them.

> On my computer, the palette argument doesn't change the appearance of the vis_dat output, but the tests seem to pass, so can you check that you have a test that catches that issue? When I run vdiffr, vis-dat-qualitative-palette.svg and vis-dat-vanilla.svg appear to be the same.

# @batpigandme's review

> vis_dat's power lies in the fact that it's essentially a visual analogue to one's data, if imagined as a frame/spreadsheet. In order to enhance this, I would suggest that the charts, by default, "read" in the spirit of this analogy by going from top to bottom.

## Documentation: README

> I have made some minor changes to the wording in README and submitted them as a pull request here, most of which I think are self-explanatory.

- Pull request was merged [here]()

> vis_dat(airquality) is numeric and integer values, when I run it. Thus, I've swapped numeric in for character as it was described.

> Added canonical link to the mi package, since you do so for wakefield

> Though the README suggests that when less than one percent is missing, "vis_miss indicates that there is >1%", the chart that results from vis_miss(test_miss_df) reads Missing (<0.1%) in the legend

- Agreed, this was addressed in commit ....

> When running the first example from README, (vis_dat(airquality)), the user will get a warning re. depecation of dmap(). If this is not of consequence to the package's functionality, then you might consider adding it to the suppressed messages. If it's something most users supress locally, then you might suggest that somewhere in the README.

```
#> dmap() is deprecated. Please use the new colwise family in dplyr.
#> E.g., summarise_all(), mutate_all(), etc.
```

- Thank you for finding this,

> Minor suggestion (which may or may not work comply with the actual vignette formatting) the purpose of demonstrating the difference between vis_guess() and vis_dat, you might consider showing the two side by side (perhaps have the vis_guess() example alone and then side by side). For this you could just add fig.show='hold', out.width='50%' to the chunk options in the Rmd.

- Excellent suggestion

> Based on the README, the status of vis_dat_ly isn't totally clear (since it's not yet a function of its own).

- Agreed, this ....

> I didn't quite understand the Visualising expectations section. That said, I've also never used expectation in assertr— so, this may fall under the category of things that those who need to know do, and, thus, no more need be said.

- I've constructed a small example to illustrate this, in commit ...

## Documentation: Vignettes

> Some of the same copy edits I made in README I would also suggest in the vignettes (e.g. the compare this to...Where sentence that sort of goes around the vis_dat figure in the vis_guess example section)

- This has been addressed in ...

> The experimental features vignette is a bit hard to follow at the moment, which may be in part due to the layout (the warning messages take up a good bit of space). The figure is also quite small. I'd recommend adjusting the settings to match your main vignette to make it more legible.

- Agreed, the error messages have been updated in committ ...

> vis_dat_ly(), and vis_miss_ly() aren't in the experimental vignette, though the opening paragraph suggests that they will be

- Thank you for picking that up, they have been added in commit ...

## Documentation: Help

> The vis_guess description is a bit hard to follow. I'd start by stating what it does do, before making the comparison to vis_dat, just so it's a bit more self-contained.

-

> The visdat and visdat-package sections are empty, when, I assume, you'll want the description to be included (right now description just says "visdat").

- Thank you for picking up on this, I've added ... in commit ...

## Examples

> Note: the function examples in R all throw warning: dmap() is deprecated.
As noted in @seaaan's comments, there's also the warning re. dataframes of different dimensions
e.g. example("vis_compare") results in

```
  #>  Warning messages:
  #> 1: In if (dim(df1) != dim(df2)) { :
  #>  the condition has length > 1 and only the first element will be used
  #> 2: attributes are not identical across measure variables; they will be dropped
  #> 3: attributes are not identical across measure variables; they will be dropped

```

- Agreed, this error message for vis_compare has been addressed in commit ...

vis_compare example also seems to cluster by default (all different data is shown at the bottom, of the two columns)

-

## Community guidelines

I don't believe these are currently in the package.

- Thank you for picking up on this, I've added them in commit ...

## Paper

> The vis_miss chunk is unnamed. I'd put a name in for you, but I can't knit the doc (see below). Either way, quick fix.

- Thank you, this is addressed in commit ...

> I'm getting an error 83 because pandoc-citeproc is missing. I believe this is something I'm missing on my machine, and, thus you don't need to worry about. But, just in case, I'm letting you know!

- This could be due to the yaml info, I have tried to run this and found ...

## Functionality

### Installation and Tests

Installation built successfully (on three different computers, and versions of R at that), as did automated testing. Good work!

- Hooray!

### Functionality

All of the functions run as expected, and as described. My feedback here primarily concerns messages and warnings, that apply to several functions.

Warning messages can be scary, so it's always a relief see them acknowledged or shown in vignette and/or README examples (as is done re. non-identical attributes).
vis_compare (a function for which I can imagine plenty of use cases) is currently returning enough text that it may leave the user confused as to what's most important, or an error on their end, etc. Currently, running vis_compare(iris_diff, iris) (as in the given example), returns:

```
#> vis_compare is still in BETA! If you have suggestions or errors,
#>        post an issue at https://github.com/njtierney/visdat/issues
#>  dmap() is deprecated. Please use the new colwise family in dplyr.
#>  E.g., summarise_all(), mutate_all(), etc.
#>  Warning messages:
#>  1: In if (dim(df1) != dim(df2)) { :
#>    the condition has length > 1 and only the first element will be used
#>  2: attributes are not identical across measure variables; they will be dropped
#>  3: attributes are not identical across measure variables; they will be dropped
```

> The function runs fine, and appears to have worked, but it's enough "warning" text, that it may warrant concern for the user.

- Absolutely agreed, this is a problem and barrier,

> vis_guess is where the cell-to-row visual disparity becomes most confusing. I think this is a fantastic function, especially for beginners, and so the ability to essentially "read" the visual, as one would a table (top to bottom), would be invaluable.

- Agreed, this has been addressed by making the "flip" argument become the hard set default. I cannot honestly see

> vis_dat with plotly / interactive functions are great, and, again, here it would be even more helpful if one could "read" the charts as one would a table.

### 2017-02-05 README.Rmd

@@
 
-Initially inspired by [`csv-fingerprint`](https://github.com/setosa/csv-fingerprint), `vis_dat` helps you visualise a dataframe and "get a look at the data" by displaying the variable classes in a dataframe as a plot with `vis_dat`, and getting a brief look into missing data patterns `vis_miss`.
+Initially inspired by [`csv-fingerprint`](https://github.com/setosa/csv-fingerprint), `vis_dat` helps you visualise a dataframe and "get a look at the data" by displaying the variable classes in a dataframe as a plot with `vis_dat`, and getting a brief look into missing data patterns using `vis_miss`.
 
@@
 - `vis_miss()` visualises just the missing data, and allows for missingness to be clustered and columns rearranged.
-`vis_miss()` is similar to `missing.pattern.plot` from the `mi` package.
+`vis_miss()` is similar to `missing.pattern.plot` from the [`mi`](https://cran.r-project.org/web/packages/mi/index.html) package.
 Unfortunately `missing.pattern.plot` is no longer in the `mi` package (well, as of 14/02/2016).
@@
 
-Let's see what's inside the dataset `airquality`
+Let's see what's inside the `airquality` dataset from base R.
 
@@
 
-The classes are represented on the legend, and missing data represented by grey.
+The classes are represented on the legend, and missing data represented by grey. The column/variable names are listed on the x axis.
 
-by default, `vis_dat` sorts the columns according to the type of the data in the vectors.
-You can turn this off by setting `sort_type = FALSE`.
-This feature is better illustrated using the `airquality` dataset from base R.
+By default, `vis_dat` sorts the columns according to the type of the data in the vectors. You can turn this off by setting `sort_type = FALSE`.
 
@@
 
-The plot above tells us that R reads this dataset as having characters and integer values, along with some missing data in `Ozone` and `Solar.R`.
+The plot above tells us that R reads this dataset as having numeric and integer values, with some missing data in `Ozone` and `Solar.R`.
 
@@
 
-We can explore the missing data further using `vis_miss()`
+We can explore the missing data further using `vis_miss()`.
 
@@
 
-You can cluster the missingness by setting `cluster = TRUE`
+You can cluster the missingness by setting `cluster = TRUE`.
 
@@
 
-`vis_miss` will also indicate when there is no missing data at all
+`vis_miss` will also indicate when there is no missing data at all.
 
@@
 
-Where you'd just assume your data is weird because it's all factors - or worse, not notice that this is a problem.
+Here, you might just assume your data is weird because it's all factors - or worse, not notice that this is a problem.
 
@@
 
-Thanks to Carson Sievert, you can now add some really nifty interactivity into visdat by using `plotly::ggplotly`, allowing for information to be revealed upon mouseover of a cell.
+Thanks to Carson Sievert, you can now add some really nifty interactivity into `visdat` by using `plotly::ggplotly`, allowing for information to be revealed upon mouseover of a cell.
 The code to do this can be seen below, but is not shown as the github README doesn't support HTML interactive graphics...yet.

### 2017-02-05 README.Rmd
README fixes relating to #39
Wording changes, links to vignette, figure size changes

@@
 
-The classes are represented on the legend, and missing data represented by grey. The column/variable names are listed on the x axis.
+The plot above tells us that R reads this dataset as having numeric and integer values, with some missing data in `Ozone` and `Solar.R`.
+The classes are represented on the legend, and missing data represented by grey.
+The column/variable names are listed on the x axis.
 
@@
 ```
-
-The plot above tells us that R reads this dataset as having numeric and integer values, with some missing data in `Ozone` and `Solar.R`.
 
@@
 
-When there is <0.1% of missingness, `vis_miss` indicates that there is >1% missingness.
+`vis_miss` indicates when there is a very small amount of missing data at <0.1% missingness.
 
@@
 
-# Experimental features
-
-## Using `vis_compare()`
-
-Sometimes you want to see what has changed in your data. `vis_compare()` helps with that. It is currently only just barely working, so keep in mind that this is very much in its beta stages.
-
-For the sake of simplicity, lets make some changes to `iris`, and compare this new dataset
-
-```{r vis-compare-iris}
-
-iris_diff <- iris
-iris_diff[sample(1:150, 30),sample(1:4, 2)] <- NA
-
-vis_compare(iris_diff, iris)
-
-```
-
-Here the differences are marked in blue.
-
-If you try and compare differences when the dimensions are different, you get an ugly error.
-
-```{r vis-compare-error, eval = FALSE}
-
-iris_diff_2 <- iris
-iris_diff_2$new_col <- iris$Sepal.Length + iris$Sepal.Width
-
-vis_compare(iris, iris_diff_2)
-#> vis_compare is still in BETA!
-If you have suggestions or errors, post an issue at https://github.com/njtierney/visdat/issuesthe condition has length > 1 and only the first element will be usedError: `.x` (5) and `.y` (6) are different lengths
-```
-
-## Using `vis_guess()`
-
-`vis_guess()` takes a guess at what each cell is. It's best illustrated using some messy data, which we'll make here.
-
-```{r create-messy-vec}
-
-messy_vector <- c(TRUE,
-                  T,
-                  "TRUE",
-                  "T",
-                  "01/01/01",
-                  "01/01/2001",
-                  NA,
-                  NaN,
-                  "NA",
-                  "Na",
-                  "na",
-                  "10",
-                  10,
-                  "10.1",
-                  10.1,
-                  "abc",
-                  "$%TG")
-
-set.seed(1114)
-messy_df <- data.frame(var1 = messy_vector,
-                       var2 = sample(messy_vector),
-                       var3 = sample(messy_vector))
-
-```
-
-```{r vis-guess-messy-df}
-
-vis_guess(messy_df)
-
-```
-
-So here we see that there are many different kinds of data in your dataframe. As an analyst this might be a depressing finding. Compare this to `vis_dat`.
-
-```{r visdat-messy-df}
-
-vis_dat(messy_df)
-
-```
-
-Here, you might just assume your data is weird because it's all factors - or worse, not notice that this is a problem.
-
-At the moment `vis_guess` is very slow.
-Please take this into consideration when you are using it on data with more than 1000 rows.
-We're looking into ways of making it faster, potentially using methods from the `parallel` package, or extending the c++ code from `readr:::collectorGuess`.
-
-# Interactivity
-
-Thanks to Carson Sievert, you can now add some really nifty interactivity into `visdat` by using `plotly::ggplotly`, allowing for information to be revealed upon mouseover of a cell.
-The code to do this can be seen below, but is not shown as the github README doesn't support HTML interactive graphics...yet.
-
-```{r eval = FALSE}
-
-library(plotly)
-
-vis_dat(airquality) %>% ggplotly()
-
-```
-
-## vis_dat_ly
-
-This is still under development, but it is basically a faster version of doing a ggplot and then calling ggplotly.
-
-## vis_miss_ly
-
-This is also under development, and still needs some more work on the legend, etc.
-
-```{r vis-miss-ly}
-
-vis_miss_ly(airquality)
-
-```
-
-# Road Map
-
-**Visualising expectations**
-
-The idea here is to pass expectations into `vis_dat` or `vis_miss`, along the lines of the `expectation` command in `assertr`.
-For example, you could ask `vis_dat` to identify those cells with values of -1 with something like this:
-
-```{r eval = FALSE}
-
-data %>% 
-  expect(value == -1) %>%
-  vis_dat
-
-```
-
 # Thank yous

### 2017-02-12 README.Rmd
Add CoC - addressing #39
link to vignettes from README, addressing #39

@@
 
-There are two experimental functions:
+There are also experimental which are described in more detail in the [vignette: "Experimental Features of visdat](http://www.njtierney.com/visdat/articles/experimental_features.html).
 
--  `vis_guess()` has a guess at what the value of each cell, using `readr::parse_guess`.
-This means that "10.1" will return "double", and `10.1` will return "double", and `01/01/01` will return "date".
-Keep in mind that it is a **guess** at what each cell is, so you can't trust this fully.
-
-- `vis_compare()` compares two dataframes, displaying the differences.
+Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.
 

### 2017-02-15 paper/response.Rmd
add paper response for rOpenSci review
small update to response

@@
 
-...
-
-Consider the above my response to the positive feedback in this review. I will now respond to suggsted changes and other constructive criticisms.
+I will now respond to suggsted changes and other constructive criticisms.
 
@@
 
--
+- Agreed, but I would prefer to keep the README as a whole piece, so I think instead I will just link to the experimental vignette as in the README.
+This has been logged in [issue 39]() and fixed in commit
 
 > Link to the vignettes from the README.
+
+- Agreed, this has been address in commit ..., issue ...
 
@@
 
--
+- Agreed, this makes more sense - changed!
 
@@
 
+Fixed in ...
+
 > I don't understand this sentence: "When there is <0.1% of missingness, vis_miss indicates that there is >1% missingness."
 
-- Agreed, this
+- Agreed, this has been changed in issue ....
 
@@
 
-- Interesting, it should be showing up as black, as this is what the [code and README reflect]() (well, grey50). Perhaps th
+- Interesting, it should be showing up as black, as this is what the [code and README reflect]() (well, grey50). Does it still happen now?
 
@@
 
-- I can't seem to replicate this
+- I can't seem to replicate this, can you maybe take a screenshot of this?
 
@@
 
--
+- Agreed, this has been done in commit ... , issue ...
 
@@
 
-- Agreed, this has been completed in this commit ...
+- Agreed, this has been completed in this commit ..., issue...
 
@@
 
-- Yes, as I thought that it might be a useful function for users, but I can appreciate that perhaps it shoudl stay internal
+- Yes, as I thought that it might be a useful function for users, although I can't think of a good usecase outside of `vis_guess` right now. Perhaps it might be best for `guess_type` to be unexported?
 
@@
 
+- Agreed, these will be added to a dev branch
+
 > This is a minor issue, but was so thoroughly drummed into me by CS professors that I have to mention it.
@@
 (3) ???
+
+- This is an interesting one!
+It was brought to my attention [here](), by @MilesMcBain, but I couldn't replicate this issue.
+Similarly, I can actually display the nycflights data.
+So it seems that this is to do with the fact that computations power has a bit to do with it.
+I think that option 1 is the best case here, but option 2 actually gets a little bit difficult, because this leads to another issue of effectively downsampling a dataset whilst maintaining key characteristics.
+This could a problem, say, if I did perform down-sampling and then the user looked at the data with visdat and plotly, they might get a mis-represented sample.
+Likewise, if there is a very large proportion of missing data, this might be difficult to preserve exactly.
+Ways around this might include some very explicit warning messages, or perhaps even a note directly on the plot that the data have been downsampled.
 
@@
                        X = x[!nas], FUN.VALUE = character(1))
-If you're interested in converting to C++ for speed, I played around a little bit with implementing a vectorized version for guess_parser in readr.
+
+```
+
+> If you're interested in converting to C++ for speed, I played around a little bit with implementing a vectorized version for guess_parser in readr.
 It's about 20X faster.
@@
 To get it up and running, clone readr and then put this in readr/src/collectorGuess.cpp:
+
+```
 std::string collectorGuess2(const std::string& input, LocaleInfo locale) {
@@
 
+- I think that this is the default behaviour of `vdiffr`?
+It states [here](https://github.com/lionel-/vdiffr#running-tests) that "New cases for which you just wrote an expectation will be skipped.
+Failed tests will show as an error."
+
 > I suggest adding a test for each function using typical_data, because that currently emits a warning message.
 That way once you stop the error message from occurring, you'll have a regression test for the future.
+
+- This is a great suggestion!
 
@@
 
+- Oops, error on my part, thanks for picking that up! :)
+
 > On my computer, the palette argument doesn't change the appearance of the vis_dat output, but the tests seem to pass, so can you check that you have a test that catches that issue?
 When I run vdiffr, vis-dat-qualitative-palette.svg and vis-dat-vanilla.svg appear to be the same.
+
+- Yes, thank you for picking up on this!
 
@@
 
-- Pull request was merged [here]()
+- Thank you! Pull request was merged [here](https://github.com/njtierney/visdat/commit/868140da16fbfc9150d8359731f4e9eb3e817675)
 
@@
 
+- Thank you!
+
 > Added canonical link to the mi package, since you do so for wakefield
+
+- Thank you!
 
@@
 
-- Thank you for finding this,
+- Thank you for finding this, I have addressed this in issue ..., and close that [here]().
 

### 2017-05-10 README.Rmd
better describe the airquality dataset, #39, #40
Add both vignettes, addressing issue #39

@@
 `vis_miss()` is similar to `missing.pattern.plot` from the [`mi`](https://cran.r-project.org/web/packages/mi/index.html) package.
-Unfortunately `missing.pattern.plot` is no longer in the `mi` package (well, as of 14/02/2016).
+Unfortunately `missing.pattern.plot` is no longer in the `mi` package (as of 14/02/2016).
 
-There are also experimental which are described in more detail in the [vignette: "Experimental Features of visdat](http://www.njtierney.com/visdat/articles/experimental_features.html).
+You can read more about visdat in the vignette, ["using visdat""](http://www.njtierney.com/visdat/articles/using_visdat.html).
+
+There are also experimental which are described in more detail in the vignette ["Experimental Features of visdat](http://www.njtierney.com/visdat/articles/experimental_features.html).
 
@@
 
-Let's see what's inside the `airquality` dataset from base R.
+Let's see what's inside the `airquality` dataset from base R, which contains information about daily air quality measurements in New York from May to September 1973.
+More information about the dataset can be found with `?airquality`.
 

### 2017-05-10 paper/paper.Rmd
update paper and response

@@
 
-As described by Hadley Wickham [@wickham2016], there are 6 phases of data science:
-
-![Process of Data Science](http://r4ds.had.co.nz/diagrams/data-science.png)
-
-You can get insight into your data by modelling, visualising, and transforming, which Wickham describes as "Understanding" or "knowledge generation".
-There is some overlap here, as in this process it wouldn't be surprising that you might uncover some feature of your dataset that you would need to clean up or change.
-For example, in analysing some data, you might discover that some strings were indeed factors, or that gender was considered numeric in your regression, warranting you to re-visit the Tidying phase.
-
-In the same way, this process gets applied in the Tidying step of data science.
+When you get a new data set, you need to look at the data to get a sense of what it contains and potential problems with it.
 You read in your data, but you then need to look at it to understand what you need to do to make it ready for analysis.

### 2017-05-10 paper/response.Rmd

@@
 
-I would like to thank the reviewers for taking the time to review visdat so thoroughly. It is fantastic to hear comments such as:
+I would like to thank the reviewers for taking the time to review visdat so thoroughly. It is fantastic to hear such encouraging comments from members of the community.
 
->
-
-,
-
->
-
-And
-
->
-
-I will now respond to suggsted changes and other constructive criticisms.
+I will now respond to suggested changes and other constructive criticisms.
 
@@
 - Agreed, but I would prefer to keep the README as a whole piece, so I think instead I will just link to the experimental vignette as in the README.
-This has been logged in [issue 39]() and fixed in commit
+This has been logged in [issue 39](https://github.com/njtierney/visdat/issues/39) and fixed in the indicated commit messages.
 
@@
 
-- Agreed, this has been address in commit ..., issue ...
+- Agreed, this has been addressed in [issue 39](https://github.com/njtierney/visdat/issues/39)
 
@@
 
-Fixed in ...
+- Thank you for this, this has been addressed in [issue 39](https://github.com/njtierney/visdat/issues/39).
 
@@
 
-- Agreed, this has been changed in issue ....
+- Agreed, this has been changed in [issue 39](https://github.com/njtierney/visdat/issues/39).
 
 2. Vignettes:
-
-> addressing comments that relate to the vignette and README
 
@@
 
-- Interesting, it should be showing up as black, as this is what the [code and README reflect]() (well, grey50). Does it still happen now?
+- I think I've fixed this now, does it still happen?
 
@@
 
-- Agreed, this has been done in commit ... , issue ...
+- Agreed, I have now linked to this additional experimental vignette in the "using visdat" vignette.
 
@@
 
-- Agreed, this has been completed in this commit ..., issue...
+- Agreed, this has been addressed
 
@@
 
-> Formatting: links, code, and other formatting need to be done with .Rd syntax. For example, for code, use \code{}, not backticks. For bold, use \strong{} instead of asterisks. See guidelines here.
+> Formatting: links, code, and other formatting need to be done with .Rd syntax. For example, for code, use \code{}, not backticks. For bold, use \strong{} instead of asterisks.
+
+- I have now converted this over to use markdown with roxygen.
 
@@
 
+- Thank you for this - this has been changed for `vis_dat.R`, `vis_compare.R`, `vis_guess.R`, and `vis_miss.R`.
+
 > ?vis_compare: The documentation for this function needs to be updated. It is more a list of ideas for how to implement the function than a description of what it does.
+
+- Thank you for this comment, this has been updated.
 
@@
 
+- Thank you for this comment, this has been updated.
+
 > ?vis_miss_ly: The reference to vis_miss in the "See Also" section should be a link. Like this: \code{\link[visdat]{vis_miss}}
 
+- Thank you for this comment, this has been updated.
+
 > ?visdat: It's good that this page exists. However, it doesn't have any content -- add a brief overview of the package with links to the documentation page for the main functions.
+
+- Excellent point, I have addressed this in this [commit](https://github.com/njtierney/visdat/commit/106e24a4e186519f6ede615c3b2822073a03d780)
 
@@
 
-> example(vis_compare) is a good example but gives a bunch of warning messages.
+- Great point, thanks for picking up on this, I have made the final changes [here](https://github.com/njtierney/visdat/commit/d71d2ac8c0708a327d51217f0298b4d3dec2b970)
 
 > example(vis_dat): "palette" is misspelled.
+
+- thank for you picking up on this, this was fixed [here](https://github.com/njtierney/visdat/commit/31909e241c1e1cc1b640c55edaea9c228f92976e)
+
+> example(vis_compare) is a good example but gives a bunch of warning messages.
 
@@
 > There are no guidelines for how to contribute, either in the README or in a CONTRIBUTING file.
+
+- This has now been added [here](https://github.com/njtierney/visdat/commit/c7ea5bb6bb12a43719fc4dbdfbccc22153072b62).
 
@@
 
+- I believe JOSS allows figures - the tidytext paper [has a figure](http://joss.theoj.org/papers/89fd1099620268fe0342ffdcdf66776f).
+
 > The paper doesn't have any references in the "References" section, but does have inline references.
 
+- These references are provided in the yaml, and in a .bib file - hopefully these get compiled with the JOSS paper, but perhaps @arfon can chime in here.
+
 > If you keep it, make sure you have permission to reproduce the image from R4DS.
+
+- I have removed the figure now, in favour of your suggestion to keep things simpler.
 
@@
 ## Common issues
-
-> There are a number of issues that apply to most or all of the functions in the package. Rather than repeat them under each function, I'll just put those comments together here.
 
@@
 I'm not quite sure of the use case for the flip argument, but maybe there is one!
+
+- I agree that the rows should be in the order that you see the dataframes. I would also prefer for the
+
+- The flip argument was placed there to experiment with this idea of flipping the
 

### 2017-05-12 paper/response.Rmd
various updates to the rOpenSci response

@@
 
-I will now respond to suggested changes and other constructive criticisms.
+Below I respond to suggested changes and other constructive criticisms.
 
@@
 
-> The README is very good!
-It gives a helpful overview of everything in the package and is clearly written, with good examples.
-All of the code examples worked on my computer.
-You might consider breaking the "experimental" parts of the package into a separate README like you do with the vignettes.
+> You might consider breaking the "experimental" parts of the package into a separate README like you do with the vignettes.
 
-- Agreed, but I would prefer to keep the README as a whole piece, so I think instead I will just link to the experimental vignette as in the README.
+- Agreed!
+I would prefer to keep the README as a whole piece, Perhaps instead I could just link to the experimental vignette as in the README.
 This has been logged in [issue 39](https://github.com/njtierney/visdat/issues/39) and fixed in the indicated commit messages.
@@
 
-- I agree with you that this is a clearer way to address this comment, and have addressed this in this commit ...
+- I agree with you that this is a clearer way to introduce the topic - this is now [changed](https://github.com/njtierney/visdat/commit/3f74141d2e9de4bc0ae870fc874b17a21829618e).
 
@@
 
-- I can't seem to replicate this, can you maybe take a screenshot of this?
+- I can't seem to replicate this, can you maybe take a screenshot of this if this is still happening?
 
@@
 
-- Agreed, I have now linked to this additional experimental vignette in the "using visdat" vignette.
+- Great point! Thank you! I have now linked to this additional experimental vignette in the "using visdat" vignette.
 
@@
 
-- Agreed, this has been addressed
+- Good idea! This has been addressed now.
 
@@
 
-- I have now converted this over to use markdown with roxygen.
+- I have now converted this over to use markdown with roxygen, this should be fixed now.
 
@@
 
-- Thank you for this comment, this has been updated.
+- Thank you for this comment, this has [been updated.](https://github.com/njtierney/visdat/blob/master/R/vis_compare.R)
 
@@
 
-- Thank you for this comment, this has been updated.
+- Thank you for this comment, this has [been updated.](https://github.com/njtierney/visdat/blob/master/R/vis_guess.R)
 
@@
 
-- Thank you for this comment, this has been updated.
+- Thank you for this comment, this has [been updated.](https://github.com/njtierney/visdat/blob/master/R/vis_miss_ly.R)
 
@@
 
-- thank for you picking up on this, this was fixed [here](https://github.com/njtierney/visdat/commit/31909e241c1e1cc1b640c55edaea9c228f92976e)
+- Thank for you picking up on this, this was fixed [here](https://github.com/njtierney/visdat/commit/31909e241c1e1cc1b640c55edaea9c228f92976e)
 
@@
 > example(vis_guess): gives warning message
+
+- These now gives the message:
+
+```
+vis_compare / (vis_guess) is in BETA! If you have suggestions or errors
+post an issue at https://github.com/njtierney/visdat/issues
+```
+Do you think this is OK?
 
@@
 
-- I have removed the figure now, in favour of your suggestion to keep things simpler.
+- I have removed the figure now, in favour of your earlier suggestion to keep things simpler - thank you for this suggestion!
 
@@
 
-- I agree that the rows should be in the order that you see the dataframes. I would also prefer for the
+> The flip argument should be available for all of the functions if you provide it for any, for the sake of consistency. Possibly also the palette argument.
 
-- The flip argument was placed there to experiment with this idea of flipping the
+- I agree that the rows should be in the order that you see the dataframes.
+I would also prefer for the column names to be on top of the visualisation, but I cannot seem to get the column names to present nicely here.
 
-> The flip argument should be available for all of the functions if you provide it for any, for the sake of consistency. Possibly also the palette argument.
+- The flip argument was placed there to experiment with this idea of flipping the plot - as you suggested, to have row 1 at the top.
+Personally, I think that now the flip argument should be removed, as I cannot now really think of any times where I would prefer the old arrangement of rows.
+
+- I agree with you re the palette, although I'm not sure how this could work for vis_miss. But perhaps I should allow for users to insert their own pallete for this.
 
@@
 
+- Thank you very much for this, I have created a helper function as you suggested, and suppressed the message in that way.
+
 > The code for flipping the rows is also duplicated:
@@
 
+- Thank you for this, I have created a helper function for these code duplications, it makes things much easier to handle!
+These are now in the [internals file](https://github.com/njtierney/visdat/blob/fix-functionality/R/internals.R)
+
 > vis_compare and vis_guess are indicated as being in beta, which seems also to apply to the plotly versions of the functions.
@@
 
-- Agreed, these will be added to a dev branch
+- Agreed, I need to create a dev branch, where these experimental features can reside.
 
@@
 There are a number of these in the code for various functions.
+
+- I've actually not heard this before, but I can certainly understand why you would want to avoid called such as if(x == TRUE) over if(x).
+Personally, however, I think I actually prefer to be really explicit.
 
@@
 
+- great suggestion, this has been added! And can be turned off in the options with `show_perc_col == FALSE`
+
 ## vis_compare()
@@
 Just an idea I had though, not a requirement at all!
+
+- I am glad you like this! It is a difficult problem that I cannot work outk, re the rows/cols matching when they are out of the dimension. I'm not quite sure what to do here.
 
@@
 > vis_guess lacks some of the parameters that vis_dat has, but should have the same options (except maybe not sort_type?)
+
+- Agree, this is a good idea.
 
@@
 
+- OK, so I think that this is a fantastic
+
 ## vis_miss_ly
@@
 
+- I will see what I can whip up in plotly, I do not know much about it at the moment, but @carson has a free book, so I'll get onto that.
+
 ## vis_dat_ly()
@@
 
+- OK, I'm pretty sure that this is an easy fix, I just specify these things in the arguments of ggplot, which plotly reads in.
+The problem is then that I need to change the names of things inside of ggplot
+
 > If you provide vis_miss_ly, it makes sense to me to also provide vis_dat_ly, and possibly also vis_compare_ly and vis_guess_ly.
 An initial implementation could just be vis_dat(x) %>% ggplotly(), but then you have the API in place at least.
+
+- I would prefer to build it in straight plotly, as plotly is really fast, and the process of converting ggplot plots into plotly is actually quite computationally intense
 
@@
 
+- Absolutely agree, I have made these changes now and it makes a LOT more sense to do it this way!
+
 ## Documentation: README
@@
 
-- Agreed, this was addressed in commit ....
+- Agreed, this has now been fixed, you can see this in the [README]()
 
 > When running the first example from README, (vis_dat(airquality)), the user will get a warning re.
-depecation of dmap().
+deprecation of dmap().
 If this is not of consequence to the package's functionality, then you might consider adding it to the suppressed messages.
@@
 
-- Excellent suggestion
+- Excellent suggestion, done!
 
@@
 
-Installation built successfully (on three different computers, and versions of R at that), as did automated testing. Good work!
+> Installation built successfully (on three different computers, and versions of R at that), as did automated testing. Good work!
 
@@
 
-All of the functions run as expected, and as described. My feedback here primarily concerns messages and warnings, that apply to several functions.
+> All of the functions run as expected, and as described. My feedback here primarily concerns messages and warnings, that apply to several functions.
 
-Warning messages can be scary, so it's always a relief see them acknowledged or shown in vignette and/or README examples (as is done re. non-identical attributes).
+> Warning messages can be scary, so it's always a relief see them acknowledged or shown in vignette and/or README examples (as is done re.
+non-identical attributes).
 vis_compare (a function for which I can imagine plenty of use cases) is currently returning enough text that it may leave the user confused as to what's most important, or an error on their end, etc.
@@
 > vis_dat with plotly / interactive functions are great, and, again, here it would be even more helpful if one could "read" the charts as one would a table.
+
+# njtierney summary response and remaining tasks to complete
+
+This has been a fantastic process to be a part of. To summarise, the main changes that I made were:
+
+- re-orienting the vis_* family to read in the way that one reads a dataframe
+- removing many of the error messages that occur due to some tidy/dplyr collation of factors
+-
+
+Still remaining to do is:
+
+- complete vis_guess() for release
+    - Adding the fast Rcpp code for guess_type
+    - remove message regarding BETA
+- create `vis_dat_ly`, `vis_miss_ly`, and `vis_guess_ly` functions
+- Ensure that all of the `vis_*` family functions have the same internal arguments where appropriate:
+  - E.g., vis_dat, doesn't need to same clustering missingness options (or...does it?)
+- move vis_compare into a "dev" branch on github
+- should I export vis_guess?
+- Yes, as I thought that it might be a useful function for users, although I can't think of a good usecase outside of `vis_guess` right now.
+Perhaps it might be best for `guess_type` to be unexported?
+
+> example(vis_guess) / example(vis_compare) : gives warning message
+
+- These now gives the message:
+
+```
+vis_compare / (vis_guess) is in BETA! If you have suggestions or errors
+post an issue at https://github.com/njtierney/visdat/issues
+```
+Do you think this is OK?

### 2017-06-08 tests/testthat/test-vis-dat.R
remove the flip argument

@@ -5,12 +5,10 @@ vis_dat_plot <- vis_dat(airquality)
 vis_dat_plot_sort_type <- vis_dat(airquality, sort_type = FALSE)
 vis_dat_plot_pal_qual <- vis_dat(airquality, palette = "qual")
 vis_dat_plot_pal_qual <- vis_dat(airquality, palette = "cb_safe")
-vis_dat_plot_flip <- vis_dat(airquality, flip = TRUE)
 
 test_that("vis_dat creates the right plot",{
   vdiffr::expect_doppelganger("vis_dat vanilla", vis_dat_plot)
   vdiffr::expect_doppelganger("vis_dat sort_type", vis_dat_plot)
   vdiffr::expect_doppelganger("vis_dat qualitative palette", vis_dat_plot)
   vdiffr::expect_doppelganger("vis_dat colourblind safe palette", vis_dat_plot)
-  vdiffr::expect_doppelganger("vis_dat flip to dataframe", vis_dat_plot)
 })

### 2017-06-08 tests/testthat/test-vis-miss.R

@@ -5,12 +5,10 @@ vis_miss_plot <- vis_miss(airquality)
 vis_miss_plot_cluster <- vis_miss(airquality, cluster = TRUE)
 vis_miss_plot_sort_rows <- vis_miss(airquality, sort_miss = TRUE)
 vis_miss_plot_show_perc <- vis_miss(airquality, show_perc = FALSE)
-vis_miss_plot_show_flip <- vis_miss(airquality, flip = TRUE)
 
 test_that("vis_miss creates the right plot",{
   vdiffr::expect_doppelganger("vis_miss vanilla", vis_miss_plot)
   vdiffr::expect_doppelganger("vis_miss cluster", vis_miss_plot_cluster)
   vdiffr::expect_doppelganger("vis_miss sort rows", vis_miss_plot_sort_rows)
   vdiffr::expect_doppelganger("vis_miss show percent", vis_miss_plot_show_perc)
-  vdiffr::expect_doppelganger("vis_miss show flip", vis_miss_plot_show_flip)
 })

### 2017-06-17 paper/response.Rmd
rOpenSci checks for review

@@
 > example(vis_compare) is a good example but gives a bunch of warning messages.
-
 > example(vis_guess): gives warning message
 
-- These now gives the message:
+- These now both give warning messages like:
 
@@
 
-- These references are provided in the yaml, and in a .bib file - hopefully these get compiled with the JOSS paper, but perhaps @arfon can chime in here.
+- These references are provided in the yaml, and in a .bib file - hopefully these get compiled with the JOSS paper perhaps @arfon can chime in here?
 
@@
 
-- I have removed the figure now, in favour of your earlier suggestion to keep things simpler - thank you for this suggestion!
+- I have removed the figure now, in favour of your earlier suggestion to keep things simpler - thank you for this suggestion! :)
 
@@
 - I agree that the rows should be in the order that you see the dataframes.
-I would also prefer for the column names to be on top of the visualisation, but I cannot seem to get the column names to present nicely here.
+I would also prefer for the column names to be on top of the visualisation - I have now implemented this, although I think that the presentation of the column names is perhaps not ahs nice.
 
 - The flip argument was placed there to experiment with this idea of flipping the plot - as you suggested, to have row 1 at the top.
-Personally, I think that now the flip argument should be removed, as I cannot now really think of any times where I would prefer the old arrangement of rows.
+The flip argument has now been removed, as I cannot now really think of any times where I would prefer the old arrangement of rows.
 
-- I agree with you re the palette, although I'm not sure how this could work for vis_miss. But perhaps I should allow for users to insert their own pallete for this.
+- I think that the default sort_type option should be to put all of the similar columns together - to me this makes it easier to identify what is similar, and what is different; when I start looking at a dataset I usually want to know what variables are what class.
+I am willing to change this, but would be interested to see what others think.
+
+- I agree with you re the palette argument being provided for all plots, although I'm not sure how this could work for vis_miss.
+But perhaps I should allow for users to insert their own pallete for this.
+Personally I think that there should be a standard palette where the common classes (character, factor, integer, numeric, datetime, etc) each have their own distinct colour, and other classes are given different factors.
 
@@
 
-- Thank you very much for this, I have created a helper function as you suggested, and suppressed the message in that way.
+- Thank you very much for this, this has bothered me for a while! I have created a helper function as you suggested, and suppressed the message in that way.
 
@@
 
-- I've actually not heard this before, but I can certainly understand why you would want to avoid called such as if(x == TRUE) over if(x).
-Personally, however, I think I actually prefer to be really explicit.
+- I've actually not heard this before, but I can certainly understand why you would want to avoid called such as if(x == TRUE) over if(x). This has now been changed
 

### 2017-07-02 paper/response.Rmd
fixed documentation spacing at 80 characters  + indenting
next time I'll use the tidyverse linter, somehow

@@
 ```
+
 Do you think this is OK?
@@
 
-- Thank you very much for this, this has bothered me for a while! I have created a helper function as you suggested, and suppressed the message in that way.
-
 > The code for flipping the rows is also duplicated:
@@
 
-- I've actually not heard this before, but I can certainly understand why you would want to avoid called such as if(x == TRUE) over if(x). This has now been changed
+<!-- - I've actually not heard this before, but I can certainly understand why you would want to avoid called such as if(x == TRUE) over if(x). This has now been changed. -->
 
@@
 - This is an interesting one!
-It was brought to my attention [here](), by @MilesMcBain, but I couldn't replicate this issue.
+It was brought to my attention [here](https://github.com/njtierney/visdat/issues/32), by @MilesMcBain, but I couldn't replicate this issue.
 Similarly, I can actually display the nycflights data.
 So it seems that this is to do with the fact that computations power has a bit to do with it.
-I think that option 1 is the best case here, but option 2 actually gets a little bit difficult, because this leads to another issue of effectively downsampling a dataset whilst maintaining key characteristics.
-This could a problem, say, if I did perform down-sampling and then the user looked at the data with visdat and plotly, they might get a mis-represented sample.
+I think that option 1 is the best case here, but option 2 actually gets a little bit difficult, because this leads to another issue of effectively downsampling a dataset whilst maintaining data structure.
+This could be a problem, say, if I did perform down-sampling and then the user looked at the data with visdat and plotly, they might get a mis-represented sample.
 Likewise, if there is a very large proportion of missing data, this might be difficult to preserve exactly.
@@
 
+-
+
 > If the user provides a palette name, you should check that it's valid. If not, throw an error with a message. Currently, if you misspell a palette name, it just goes ahead with the default palette.
+
+- Great point! I have fixed this now with:
+
+```r
+
+     } else  {
+       warning("palette arguments need to be either 'qual' 'cb_safe' or 'default'")
+     } # close else brace
+
+```
+
+at the end of `vis_dat.R`
 
@@
 
+- Thank you for picking up on these comments, they are now removed!
+
 ## vis_miss()
@@
 
-- great suggestion, this has been added! And can be turned off in the options with `show_perc_col == FALSE`
+- great suggestion, this has been added! And can be turned off in the options with `show_perc_col == FALSE`.
 
@@
 
-- I am glad you like this! It is a difficult problem that I cannot work outk, re the rows/cols matching when they are out of the dimension. I'm not quite sure what to do here.
+- I am glad you like this!
+It is a difficult problem that I cannot work out, re the rows/cols matching when they are out of the dimension.
+I'm not quite sure what to do here to make this possible, I think that this should be tackled in a future release of `visdat`.
 
@@
 
+```r
+
+  if (!identical(dim(df1), dim(df2))){
+    stop("Dimensions of df1 and df2 are not the same. vis_compare currently only handles dataframes of identical dimensions.")
+  }
+
+```
+
+- Thank you for both of these comments! These have been changed in `vis_compare.R`.
+
 ## vis_guess
@@
 
-- Agree, this is a good idea.
+- Agree, this is a good idea, I have added a helper palette function now so that `vis_guess` and `vis_dat` can share the same palette
 
@@
 
-- OK, so I think that this is a fantastic
+- OK, so I think that this is a fantastic idea!
+But I'm not sure on the best way to move this into production of `visdat`, since it sounds like we would need to borrow a chunk of code from readr and then modify it.
+I wonder if perhaps the solution here is to do a pull request for this specific feature in `readr`?
 
@@
 
-- I will see what I can whip up in plotly, I do not know much about it at the moment, but @carson has a free book, so I'll get onto that.
-
 ## vis_dat_ly()
@@
 
+> If you provide vis_miss_ly, it makes sense to me to also provide vis_dat_ly, and possibly also vis_compare_ly and vis_guess_ly.
+An initial implementation could just be vis_dat(x) %>% ggplotly(), but then you have the API in place at least.
+
+- I will see what I can whip up in plotly, I do not know much about it at the moment, but @carson has a free book, so I'll get onto that.
+
 - OK, I'm pretty sure that this is an easy fix, I just specify these things in the arguments of ggplot, which plotly reads in.
 The problem is then that I need to change the names of things inside of ggplot
-
-> If you provide vis_miss_ly, it makes sense to me to also provide vis_dat_ly, and possibly also vis_compare_ly and vis_guess_ly.
-An initial implementation could just be vis_dat(x) %>% ggplotly(), but then you have the API in place at least.
 
@@
 
-- I think that this is the default behaviour of `vdiffr`?
-It states [here](https://github.com/lionel-/vdiffr#running-tests) that "New cases for which you just wrote an expectation will be skipped.
-Failed tests will show as an error."
+- Yes, the
 
@@
 
-- Thank you for picking that up, they have been added in commit ...
+- Thank you for picking that up, they have been added now.
 
@@
 
--
+- done, great suggestion!
 
@@
 
-- Thank you for picking up on this, I've added ... in commit ...
+- Thank you for picking up on this, I've updated the main descriotion file now
 
@@
 
-vis_compare example also seems to cluster by default (all different data is shown at the bottom, of the two columns)
+> vis_compare example also seems to cluster by default (all different data is shown at the bottom, of the two columns)
 
@@
 
-I don't believe these are currently in the package.
+> I don't believe these are currently in the package.
 
-- Thank you for picking up on this, I've added them in commit ...
+- Thank you for picking up on this, I've added them now, :)
 
@@
 
-- Agreed, this has been addressed by making the "flip" argument become the hard set default. I cannot honestly see
+- Agreed, this has been addressed by making the "flip" argument become the hard set default.
+I cannot honestly see a good reason to use the old print method, so I have removed it to make it read top down first.
 
@@
 - removing many of the error messages that occur due to some tidy/dplyr collation of factors
--
+- refactoring repeated code into internal functions to more clearly express what
 
@@
     - remove message regarding BETA
+- Add interactivity
   - create `vis_dat_ly`, `vis_miss_ly`, and `vis_guess_ly` functions
@@
 Perhaps it might be best for `guess_type` to be unexported?
+
+- For the moment I have decided to put visdat onto CRAN, with the main functions
+being: 
+  - vis_dat
+  - vis_miss
+  - vis_guess
+- Then, in a subsequent release, I will add the interactive functions:
+  - vis_dat_ly
+  - vis_miss_ly
+  - vis_guess_ly
+  - vis_compare
+- And also look into speed improvements for `vis_guess`
 

### 2017-07-02 R/guess_type.R

@@ -1,10 +1,11 @@
 #' Guess the type of each individual cell in a dataframe
 #'
-#' `guess_type` is used in vis_guess as a utility function. Similar to `fingerprint`.
+#' `guess_type` is used by vis_guess to guess cell elements, like `fingerprint`.
 #'
 #' @param x is a vector of values you want to guess
 #'
-#' @return a character vector that describes the suspected class. e.g., "10" is an integer, "20.11" is a double, "text" is character, etc.
+#' @return a character vector that describes the suspected class. e.g., "10" is
+#'   an integer, "20.11" is a double, "text" is character, etc.
 #'
 #' @examples
 #'
@@ -49,8 +50,6 @@ guess_type <- function(x){
 #'   guess_vector()
 #'
 #' messy_df %>%
-#'
-#'
 #'
 #' mb_df <-
 #' microbenchmark::microbenchmark(

### 2017-07-02 R/miss_guide_label.R

@@ -4,7 +4,9 @@
 #'
 #' @param x is a dataframe passed from vis_miss(x).
 #'
-#' @return a data_frame with two columns `p_miss_lab` and `p_pres_lab`, containing the labels to use for present and missing. A dataframe is returned because I think it is a good style habit compared to a list.
+#' @return a data_frame with two columns `p_miss_lab` and `p_pres_lab`,
+#'   containing the labels to use for present and missing. A dataframe is
+#'   returned because I think it is a good style habit compared to a list.
 #'
 #'
 miss_guide_label <- function(x) {

### 2017-07-02 R/typical_data.r

@@ -1,18 +1,23 @@
 #' A small toy dataset of imaginary people
 #'
-#' A dataset containing information about some randomly generated people, created using the excellent `wakefield` package. It is created as deliberately messy dataset.
+#' A dataset containing information about some randomly generated people,
+#'   created using the excellent `wakefield` package. It is created as
+#'   deliberately messy dataset.
 #'
 #' @format A data frame with 5000 rows and 11 variables:
 #' \describe{
-#'   \item{ID}{Unique identifier for each individual, a sequential character vector of zero-padded identification numbers (IDs). see ?wakefield::id for more info}
-#'   \item{Race}{Race for each individual, "Black", "White", "Hispanic", "Asian", "Other", "Bi-Racial", "Native", and "Hawaiin", see ?wakefield::race for more info}
-#'   \item{Age}{Age of each individual, see ?wakefield::age for more info}
-#'   \item{Sex}{Male or female, see ?wakefield::sex for more info}
-#'   \item{Height(cm)}{Height in centimeters, see ?wakefield::height for more info}
-#'   \item{IQ}{vector of intelligence quotients (IQ), see ?wakefield::iq for more info}
-#'   \item{Smokes}{whether or not this person smokes, see ?wakefield::smokes for more info}
-#'   \item{Income}{Yearly income in dollars, see ?wakefield::income for more info}
-#'   \item{Died}{Whether or not this person has died yet., see ?wakefield::died for more info}
+#'   \item{ID}{Unique identifier for each individual, a sequential character
+#'     vector of zero-padded identification numbers (IDs). see ?wakefield::id}
+#'   \item{Race}{Race for each individual, "Black", "White", "Hispanic",
+#'     "Asian", "Other", "Bi-Racial", "Native", and "Hawaiin", see
+#'     ?wakefield::race}
+#'   \item{Age}{Age of each individual, see ?wakefield::age}
+#'   \item{Sex}{Male or female, see ?wakefield::sex }
+#'   \item{Height(cm)}{Height in centimeters, see ?wakefield::height}
+#'   \item{IQ}{vector of intelligence quotients (IQ), see ?wakefield::iq}
+#'   \item{Smokes}{whether or not this person smokes, see ?wakefield::smokes}
+#'   \item{Income}{Yearly income in dollars, see ?wakefield::income}
+#'   \item{Died}{Whether or not this person has died yet., see ?wakefield::died}
 #' }
 #'
 #' @note the following code was used to create this data

### 2017-07-02 R/typical_larger_data.R

@@ -1,6 +1,8 @@
 #' A small toy dataset of imaginary people
 #'
-#' A wider dataset than `typical_data` containing information about some randomly generated people, created using the excellent `wakefield` package. It is created as deliberately odd / eclectic dataset.
+#' A wider dataset than `typical_data` containing information about some
+#' randomly generated people, created using the excellent `wakefield` package.
+#' It is created as deliberately odd / eclectic dataset.
 #'
 #' @format A data frame with 300 rows and 49 variables:
 #' \describe{
@@ -33,7 +35,8 @@
 #'   \item{Level}{levels between 1 and 4}
 #'   \item{Likert}{likert response - "strongly agree", "agree", and so on}
 #'   \item{Lorem_Ipsum}{lorem ipsum text}
-#'   \item{Marital}{marital status- "married", "divorced", "widowed", "separated", etc}
+#'   \item{Marital}{marital status- "married", "divorced", "widowed",
+#'     "separated", etc}
 #'   \item{Military}{miliary branch they are in}
 #'   \item{Month}{their favorite month}
 #'   \item{Name}{their name}

### 2017-07-02 R/vis_compare.R

@@ -1,12 +1,15 @@
 #' Visually compare two dataframes and see where they are different.
 #'
-#' `vis_compare`, like the other `vis_*` families, gives an at-a-glance ggplot of a dataset, but in this case, hones in on visualising **two** different dataframes of the same dimension.
+#' `vis_compare`, like the other `vis_*` families, gives an at-a-glance ggplot
+#'   of a dataset, but in this case, hones in on visualising **two** different
+#'   dataframes of the same dimension.
 #'
 #' @param df1 The first dataframe to compare
 #'
 #' @param df2 The second dataframe to compare to the first.
 #'
-#' @return `ggplot2` object displaying which values in each data frame are present in each other.
+#' @return `ggplot2` object displaying which values in each data frame are
+#'   present in each other.
 #'
 #' @seealso [vis_miss()] [vis_dat()] [vis_miss_ly()] [vis_compare()]
 #'

### 2017-07-02 R/vis_dat.R

@@ -1,14 +1,24 @@
 #' Visualises a data.frame to tell you what it contains.
 #'
-#' `vis_dat` gives you an at-a-glance ggplot object of what is inside a dataframe. Cells are coloured according to what class they are and whether the values are missing. As `vis_dat` returns a ggplot object, it is very easy to customize and change labels, and customize the plot
+#' `vis_dat` gives you an at-a-glance ggplot object of what is inside a
+#'   dataframe. Cells are coloured according to what class they are and whether
+#'   the values are missing. As `vis_dat` returns a ggplot object, it is very
+#'   easy to customize and change labels, and customize the plot
 #'
 #' @param x a data.frame object
 #'
-#' @param sort_type logical TRUE/FALSE. When TRUE (default), it sorts by the type in the column to make it easier to see what is in the data
+#' @param sort_type logical TRUE/FALSE. When TRUE (default), it sorts by the
+#'   type in the column to make it easier to see what is in the data
 #'
-#' @param palette character "default", "qual" or "cb_safe". "default" (the default) provides the stock ggplot scale for separating the colours. "qual" uses an experimental qualitative colour scheme for providing distinct colours for each Type. "cb_safe" is a set of colours that are appropriate for those with colourblindness. "qual" and "cb_safe" are drawn from http://colorbrewer2.org/.
+#' @param palette character "default", "qual" or "cb_safe". "default" (the
+#'   default) provides the stock ggplot scale for separating the colours.
+#'   "qual" uses an experimental qualitative colour scheme for providing
+#'   distinct colours for each Type. "cb_safe" is a set of colours that are
+#'   appropriate for those with colourblindness. "qual" and "cb_safe" are drawn
+#'   from http://colorbrewer2.org/.
 #'
-#' @return `ggplot2` object displaying the type of values in the data frame and the position of any missing values.
+#' @return `ggplot2` object displaying the type of values in the data frame and
+#'   the position of any missing values.
 #'
 #' @seealso [vis_miss()] [vis_miss_ly()] [vis_guess()] [vis_compare()]
 #'

### 2017-07-02 R/vis_guess.R

@@ -1,11 +1,21 @@
 #' Visualise type guess in a data.frame
 #'
-#' `vis_guess` visualises the class of every single individual cell in a dataframe and displays it as ggplot object, similar to `vis_dat`. Cells are coloured according to what class they are and whether the values are missing. `vis_guess` estimates the class of individual elements using `readr::guess_parser`.  Currently it may be slow on larger datasets.
+#' `vis_guess` visualises the class of every single individual cell in a
+#'   dataframe and displays it as ggplot object, similar to `vis_dat`. Cells
+#'   are coloured according to what class they are and whether the values are
+#'   missing. `vis_guess` estimates the class of individual elements using
+#'   `readr::guess_parser`.  Currently it may be slow on larger datasets.
 #'
 #' @param x a data.frame object
-#' @param palette character "default", "qual" or "cb_safe". "default" (the default) provides the stock ggplot scale for separating the colours. "qual" uses an experimental qualitative colour scheme for providing distinct colours for each Type. "cb_safe" is a set of colours that are appropriate for those with colourblindness. "qual" and "cb_safe" are drawn from http://colorbrewer2.org/.
+#' @param palette character "default", "qual" or "cb_safe". "default" (the
+#'   default) provides the stock ggplot scale for separating the colours.
+#'   "qual" uses an experimental qualitative colour scheme for providing
+#'   distinct colours for each Type. "cb_safe" is a set of colours that are
+#'   appropriate for those with colourblindness. "qual" and "cb_safe" are drawn
+#'   from http://colorbrewer2.org/.
 #'
-#' @return `ggplot2` object displaying the guess of the type of values in the data frame and the position of any missing values.
+#' @return `ggplot2` object displaying the guess of the type of values in the
+#'   data frame and the position of any missing values.
 #'
 #' @seealso [vis_miss()] [vis_dat()] [vis_miss_ly()] [vis_compare()]
 #'

### 2017-07-02 R/vis_miss.R

@@ -1,18 +1,26 @@
 #' Visualise a data.frame to display missingness.
 #'
-#' `vis_miss` provides an at-a-glance ggplot of the missingness inside a dataframe, colouring cells according to missingness, where black indicates a missing cell and grey indicates a present cell. As it returns a ggplot object, it is very easy to customize and change labels.
+#' `vis_miss` provides an at-a-glance ggplot of the missingness inside a
+#'   dataframe, colouring cells according to missingness, where black indicates
+#'   a missing cell and grey indicates a present cell. As it returns a ggplot
+#'   object, it is very easy to customize and change labels.
 #'
 #' @param x a data.frame
 #'
-#' @param cluster logical. TRUE specifies that you want to use hierarchical clustering (mcquitty method) to arrange rows according to missingness. FALSE specifies that you want to leave it as is
+#' @param cluster logical. TRUE specifies that you want to use hierarchical
+#'   clustering (mcquitty method) to arrange rows according to missingness.
+#'   FALSE specifies that you want to leave it as is.
 #'
 #' @param sort_miss logical. TRUE arranges the columns in order of missingness
 #'
-#' @param show_perc logical. TRUE now adds in the \% of missing/complete data in the whole dataset into the legend. Default value is TRUE
+#' @param show_perc logical. TRUE now adds in the \% of missing/complete data
+#'   in the whole dataset into the legend. Default value is TRUE.
 #'
-#'@param show_perc_col logical. TRUE adds in the \% missing data in a given column into the x axis. Can be disabled if
+#'@param show_perc_col logical. TRUE adds in the \% missing data in a given
+#'  column into the x axis. Can be disabled with FALSE
 #'
-#' @return `ggplot2` object displaying the position of missing values in the dataframe, and the percentage of values missing and present.
+#' @return `ggplot2` object displaying the position of missing values in the
+#'   dataframe, and the percentage of values missing and present.
 #'
 #' @seealso [vis_miss_ly()] [vis_dat()] [vis_guess()] [vis_compare()]
 #'
@@ -59,14 +67,6 @@ vis_miss <- function(x,
     # get the names of those columns
     col_order_index <- names(x)[na_sort]
 
-    # original code was a bit slower:
-    #
-    # col_order_index <-
-    #   x.na %>%
-    #   as.data.frame %>%
-    #   dplyr::summarise_each(funs(mean)) %>%
-    #   names
-
   } else {
 
     col_order_index <- names(x)
[…]

### 2017-07-02 R/vis_miss_ly.R

@@ -1,6 +1,8 @@
 #' Interactive plot of missingness in plotly
 #'
-#' This is a basic implementation of vis_miss in plotly, some work is still needed to fix the legend, but it is several (read, about 3000x) faster than doing `ggplotly(vis_miss(data))`.
+#' This is a basic implementation of vis_miss in plotly, some work is still
+#'   needed to fix the legend, but it is several (read, about 3000x) faster
+#'   than doing `ggplotly(vis_miss(data))`.
 #'
 #' @param x a dataframe
 #'

### 2017-07-02 R/visdat-package.r

@@ -1,6 +1,8 @@
 #' visdat
 #'
-#' visdat is a package that helps with the preliminary visualisation of data. visdat makes it easy to visualise your whole dataset so that you can visually identify problems.
+#' visdat is a package that helps with the preliminary visualisation of data.
+#'   visdat makes it easy to visualise your whole dataset so that you can
+#'   visually identify problems.
 #'
 #' @seealso
 #'
@@ -13,7 +15,7 @@
 #'   \item [vis_miss_ly()]
 #' }
 #'
-#' To learn more about visdat, see the vignette at \url{www.njtierney.com/visdat/articles/using_visdat.html}
+#' Learn more about visdat at \url{www.njtierney.com/visdat/articles/using_visdat.html}
 #' @name visdat
 #' @docType package
 #' @importFrom magrittr %>%

### 2017-07-03 R/internals.R
variable not variables

@@ -8,7 +8,7 @@
 vis_gather_ <- function(x){
   x %>%
   dplyr::mutate(rows = seq_len(nrow(x))) %>%
-    tidyr::gather_(key_col = "variables",
+    tidyr::gather_(key_col = "variable",
                    value_col = "valueType",
                    gather_cols = names(.)[-length(.)])
 }
@@ -28,7 +28,7 @@ vis_extract_value_ <- function(x){
 
   suppressWarnings(
     tidyr::gather_(x,
-                   "variables",
+                   "variable",
                    "value",
                    names(x))$value
   )
@@ -45,7 +45,7 @@ vis_extract_value_ <- function(x){
 vis_create_ <- function(x){
 
   ggplot2::ggplot(data = x,
-                  ggplot2::aes_string(x = "variables",
+                  ggplot2::aes_string(x = "variable",
                                       y = "rows",
                                     # text assists with plotly mouseover
                                     text = "value")) +

### 2017-07-03 R/vis_compare.R

@@ -2,14 +2,14 @@
 #'
 #' `vis_compare`, like the other `vis_*` families, gives an at-a-glance ggplot
 #'   of a dataset, but in this case, hones in on visualising **two** different
-#'   dataframes of the same dimension.
+#'   dataframes of the same dimension, and so takes two dataframes as arguments.
 #'
 #' @param df1 The first dataframe to compare
 #'
 #' @param df2 The second dataframe to compare to the first.
 #'
 #' @return `ggplot2` object displaying which values in each data frame are
-#'   present in each other.
+#'   present in each other, and which are not.
 #'
 #' @seealso [vis_miss()] [vis_dat()] [vis_miss_ly()] [vis_compare()]
 #'
@@ -56,7 +56,7 @@ vis_compare <- function(df1,
   # then we plot it
   ggplot2::ggplot(data = d,
                   ggplot2::aes_string(
-                    x = "variables",
+                    x = "variable",
                     y = "rows",
                     # text assists with plotly mouseover
                     # currently this text argument prevents it from being used

### 2017-07-03 R/vis_guess_ly.R

@@ -40,6 +40,8 @@
 #'                        var2 = sample(messy_vector),
 #'                        var3 = sample(messy_vector))
 #' vis_guess_ly(messy_df)
+#'
+#' @export
 
 vis_guess_ly <- function(x, palette = "default"){
 

### 2017-07-03 tests/testthat/test-visdat-internals.R

@@ -1,13 +1,10 @@
 context("Internals")
 
-test_vis_gather_ <- airquality %>%
-  as.data.frame %>%
-  vis_gather_()
+test_vis_gather_ <- suppressWarnings(vis_gather_(typical_data))
 
-test_old_gather <- airquality %>%
-  as.data.frame %>%
+test_old_gather <- typical_data %>%
   dplyr::mutate(rows = seq_len(nrow(.))) %>%
-  tidyr::gather_(key_col = "variables",
+  tidyr::gather_(key_col = "variable",
                  value_col = "valueType",
                  gather_cols = names(.)[-length(.)])
 
@@ -18,20 +15,19 @@ test_that("vis_gather_ returns the same as previous",{
 
 })
 
-d_old <-
-  airquality %>%
+d_old <- typical_data %>%
   purrr::map_df(fingerprint) %>%
   vis_gather_()
 
-d_old$value <-  tidyr::gather_(airquality,
+d_old$value <-  tidyr::gather_(typical_data,
                                "variables",
-                               "value", names(airquality))$value
+                               "value", names(typical_data))$value
 
 d_new <-
-  airquality %>%
+  typical_data %>%
   purrr::map_df(fingerprint) %>%
   vis_gather_() %>%
-  dplyr::mutate(value = vis_extract_value_(airquality))
+  dplyr::mutate(value = vis_extract_value_(typical_data))
 # get the values here so plotly can make them visible
 
 test_that("vis_extract_value performs the same as old method",{

### 2017-07-03 tests/testthat/test-vis-dat.R
use typical_data not airquality

@@ -1,10 +1,10 @@
 context("vis_dat")
 
 # try out all the options
-vis_dat_plot <- vis_dat(airquality)
-vis_dat_plot_sort_type <- vis_dat(airquality, sort_type = FALSE)
-vis_dat_plot_pal_qual <- vis_dat(airquality, palette = "qual")
-vis_dat_plot_pal_cb <- vis_dat(airquality, palette = "cb_safe")
+vis_dat_plot <- vis_dat(typical_data)
+vis_dat_plot_sort_type <- vis_dat(typical_data, sort_type = FALSE)
+vis_dat_plot_pal_qual <- vis_dat(typical_data, palette = "qual")
+vis_dat_plot_pal_cb <- vis_dat(typical_data, palette = "cb_safe")
 
 test_that("vis_dat creates the right plot",{
   vdiffr::expect_doppelganger("vis_dat vanilla",
@@ -19,5 +19,5 @@ test_that("vis_dat creates the right plot",{
 
 
 test_that("vis_dat fails when the wrong palette is provided",{
-  testthat::expect_error(vis_dat(airquality, palette = "wat"))
+  testthat::expect_error(vis_dat(typical_data, palette = "wat"))
 })

### 2017-07-03 tests/testthat/test-vis-guess.R

@@ -1,10 +1,10 @@
 context("vis_guess")
 
 # try out all the options
-vis_guess_plot <- vis_guess(airquality)
-vis_guess_plot_cb_safe <- vis_guess(airquality, palette = "cb_safe")
-vis_guess_plot_qual <- vis_guess(airquality, palette = "qual")
-vis_guess_plot_default <- vis_guess(airquality, palette = "default")
+vis_guess_plot <- vis_guess(typical_data)
+vis_guess_plot_cb_safe <- vis_guess(typical_data, palette = "cb_safe")
+vis_guess_plot_qual <- vis_guess(typical_data, palette = "qual")
+vis_guess_plot_default <- vis_guess(typical_data, palette = "default")
 
 test_that("vis_guess creates the right plot",{
   vdiffr::expect_doppelganger("vis_guess vanilla", vis_guess_plot)
@@ -14,5 +14,5 @@ test_that("vis_guess creates the right plot",{
 })
 
 test_that("vis_guess fails when the wrong palette is provided",{
-  testthat::expect_error(vis_guess(airquality, palette = "wat"))
+  testthat::expect_error(vis_guess(typical_data, palette = "wat"))
 })

### 2017-07-03 tests/testthat/test-vis-miss.R

@@ -1,10 +1,10 @@
 context("vis_miss")
 
 # try out all the options
-vis_miss_plot <- vis_miss(airquality)
-vis_miss_plot_cluster <- vis_miss(airquality, cluster = TRUE)
-vis_miss_plot_sort_rows <- vis_miss(airquality, sort_miss = TRUE)
-vis_miss_plot_show_perc <- vis_miss(airquality, show_perc = FALSE)
+vis_miss_plot <- vis_miss(typical_data)
+vis_miss_plot_cluster <- vis_miss(typical_data, cluster = TRUE)
+vis_miss_plot_sort_rows <- vis_miss(typical_data, sort_miss = TRUE)
+vis_miss_plot_show_perc <- vis_miss(typical_data, show_perc = FALSE)
 
 test_that("vis_miss creates the right plot",{
   vdiffr::expect_doppelganger("vis_miss vanilla", vis_miss_plot)

### 2017-07-03 tests/testthat/test-guess-type.R
unexport guess_type

@@ -2,99 +2,14 @@ context("Guess type")
 
 
 test_that("guess_element correctly identifies individual elements", {
-  expect_equal(guess_type(TRUE), "logical")
-  expect_equal(guess_type(T), "logical")
-  expect_equal(guess_type("TRUE"), "logical")
-  expect_equal(guess_type("T"), "logical")
-  expect_equal(guess_type("10"), "integer")
-  expect_equal(guess_type(10), "integer")
-  expect_equal(guess_type("10.1"), "double")
-  expect_equal(guess_type(10.1), "double")
-  expect_equal(guess_type("abc"), "character")
-  expect_equal(guess_type("$%TG"), "character")
-  # expect_equal(guess_type(NA), NA)
-  # expect_equal(guess_type("NA"), NA)
-  # expect_equal(guess_type("01/01/01"), "date")
-  # expect_equal(guess_type("01/01/2001"), "date")
+  expect_equal(visdat:::guess_type(TRUE), "logical")
+  expect_equal(visdat:::guess_type(T), "logical")
+  expect_equal(visdat:::guess_type("TRUE"), "logical")
+  expect_equal(visdat:::guess_type("T"), "logical")
+  expect_equal(visdat:::guess_type("10"), "integer")
+  expect_equal(visdat:::guess_type(10), "integer")
+  expect_equal(visdat:::guess_type("10.1"), "double")
+  expect_equal(visdat:::guess_type(10.1), "double")
+  expect_equal(visdat:::guess_type("abc"), "character")
+  expect_equal(visdat:::guess_type("$%TG"), "character")
 })
-#
-# messy_vector <- c(TRUE,
-#                   T,
-#                   "TRUE",
-#                   "T",
-#                   NA,
-#                   NaN,
-#                   "NA",
-#                   "Na",
-#                   "na",
-#                   "10",
-#                   10,
-#                   "10.1",
-#                   10.1,
-#                   "abc",
-#                   "$%TG")
-#
-# expected_vector <- c("logical",
-#                      "logical",
-#                      "logical",
-#                      "logical",
-#                      NA,
-#                      NA,
-#                      NA,
-#                      NA,
-#                      NA,
-#                      "integer",
-#                      "integer",
-#                      "double",
-#                      "double",
-#                      "character",
-#                      "character")
-#
[…]

### 2017-07-03 paper/response.Rmd
about 95% done on the rOpenSci review

@@
 
-- I think I've fixed this now, does it still happen?
+- This is fixed this now.
 
@@
 
-- Yes, as I thought that it might be a useful function for users, although I can't think of a good usecase outside of `vis_guess` right now. Perhaps it might be best for `guess_type` to be unexported?
+- Yes, as I thought that it might be a useful function for users, although I can't think of a good usecase outside of `vis_guess` right now. For the moment I will unexport guess_type now.
 
@@
 - I agree that the rows should be in the order that you see the dataframes.
-I would also prefer for the column names to be on top of the visualisation - I have now implemented this, although I think that the presentation of the column names is perhaps not ahs nice.
+I would also prefer for the column names to be on top of the visualisation - I have now implemented this, although I think that the presentation of the column names is perhaps not as nice.
 
@@
 - I think that the default sort_type option should be to put all of the similar columns together - to me this makes it easier to identify what is similar, and what is different; when I start looking at a dataset I usually want to know what variables are what class.
-I am willing to change this, but would be interested to see what others think.
+I am willing to change this, however.^
 
@@
 
-- Agreed, I need to create a dev branch, where these experimental features can reside.
+- Agreed, I will create a dev branch, where these experimental features can reside.
 
@@
 
-<!-- - I've actually not heard this before, but I can certainly understand why you would want to avoid called such as if(x == TRUE) over if(x). This has now been changed. -->
+- I've actually not heard this before, but I can certainly understand why you would want to avoid called such as if(x == TRUE) over if(x). This has now been changed.
 
@@
 Ways around this might include some very explicit warning messages, or perhaps even a note directly on the plot that the data have been downsampled.
+^At the moment one solution forward is to have a check first, something like:
+
+```r
+vis_dat(really_large_data)
+
+Warning: This dataset contains over 10,000 rows, which may take a while to render. Are you sure that you want to continue?
+
+1: Yes
+2. No (You should use a sample of your data)
+```
+
+What do you think?
 
@@
 
--
+- ^I believe that this is now fixed,
+
+```r
+vis_dat(airquality, palette = "qual")
+vis_dat(airquality, palette = "cb_safe")
+```
+
+Return different plots now, right?
 
@@
 ```r
-
      } else  {
@@
 It is a difficult problem that I cannot work out, re the rows/cols matching when they are out of the dimension.
-I'm not quite sure what to do here to make this possible, I think that this should be tackled in a future release of `visdat`.
+I'm not quite sure what to do here to make this possible, I think that this should be tackled in a future release of `visdat`, and will be moved to a development branch.
 
@@
 
-- OK, so I think that this is a fantastic idea!
+- ^OK, so I think that this is a fantastic idea!
 But I'm not sure on the best way to move this into production of `visdat`, since it sounds like we would need to borrow a chunk of code from readr and then modify it.
@@
 
+- I can get it close! Getting these the same will
+
 > Why not have the same arguments for this function as for vis_miss?
+
+- At the moment this is because I do now know the API for plotly that well.
+Once we are happy with the default arguments and appearance for vis_miss, we can focus on getting the same appearance with plotly
 
@@
 
+- I can get this to say "variable" instead of "variables", but I cannot change "valueType", as I think that changing this makes breaking changes.
+"rows" appears to be autogenerated from the `plotly::ggplotly` code, so I cannot change this.
+
 > If you provide vis_miss_ly, it makes sense to me to also provide vis_dat_ly, and possibly also vis_compare_ly and vis_guess_ly.
@@
 
-- I will see what I can whip up in plotly, I do not know much about it at the moment, but @carson has a free book, so I'll get onto that.
+- I have whipped up `vis_dat_ly`, `vis_compare_ly`, and `vis_guess_ly` using `ggplotly(vis_*)`.
 
-- OK, I'm pretty sure that this is an easy fix, I just specify these things in the arguments of ggplot, which plotly reads in.
-The problem is then that I need to change the names of things inside of ggplot
-
-- I would prefer to build it in straight plotly, as plotly is really fast, and the process of converting ggplot plots into plotly is actually quite computationally intense
+- I would prefer to build it in straight plotly, as plotly is really fast, and the process of converting ggplot plots into plotly is a bit computationally intense.
+I will see what I can whip up in plotly in the future.
+I do not know much about it at the moment, but @carson has a free book, so I'll get onto that.
 
@@
 
-- Yes, the
+- Yes, you require the references.
 
@@
 
-- Yes, thank you for picking up on this!
+- Yes, done. Thank you for picking up on this!
 
@@
 
-- Agreed, this has now been fixed, you can see this in the [README]()
+- Agreed, this has now been fixed.
 
@@
 
-- Thank you for finding this, I have addressed this in issue ..., and close that [here]().
+- Thank you for finding this, I have fixed this by suppressing the error as @seaaan suggested.
 
@@
 
-- Agreed, this ....
+- I have removed the mentions of these, and put them into the experimental features vignette, which is referenced in the README.
 
@@
 
-- I've constructed a small example to illustrate this, in commit ...
+- I've constructed a small example to illustrate this at the end of the experimental features vignette.
 
@@
 
-- This has been addressed in ...
-
 > The experimental features vignette is a bit hard to follow at the moment, which may be in part due to the layout (the warning messages take up a good bit of space).
@@
 
-- Agreed, the error messages have been updated in committ ...
+- Agreed, the warning messages have been fixed, and the plot size changed as well.
 
@@
 
-- done, great suggestion!
+- Done, great suggestion!
 
@@
 
-- Agreed, this error message for vis_compare has been addressed in commit ...
+- Agreed, this error message for vis_compare has been addressed.
 
@@
 
--
+- I believe this is because I changed those particular values, I'll change the values to be something else now, so that it is clear that `vis_compare()` identifies values...
 
@@
 
-- Thank you, this is addressed in commit ...
+- Done!
 
@@
 
-- This could be due to the yaml info, I have tried to run this and found ...
+- This could be due to the yaml info, does it still happen?
 
@@
 
-- Absolutely agreed, this is a problem and barrier,
+- Absolutely agreed, this is a problem and barrier, the messages for using `vis_compare` have been reduced to:
+
+```r
+vis_compare(typical_data)
+...
+```
 
@@
 
+- This has been implemented for all but `vis_miss_ly` - as this is written in plotly, and I'm not certain how to get it to flip.
+
 # njtierney summary response and remaining tasks to complete
@@
 - removing many of the error messages that occur due to some tidy/dplyr collation of factors
-- refactoring repeated code into internal functions to more clearly express what
+- refactoring repeated code into internal functions to more clearly express what was happening, and avoiding repetition.
+-
 
@@
 
+- Make a separate branch for the experimental features.
+  - vis_compare()
+  - vis_guess()
+  - vis_*_ly() family
 - complete vis_guess() for release
     - Adding the fast Rcpp code for guess_type
-    - remove message regarding BETA
-- Add interactivity
-  - create `vis_dat_ly`, `vis_miss_ly`, and `vis_guess_ly` functions
+- Add interactivity using pure plotly
+  - create `vis_dat_ly`, `vis_miss_ly`, and `vis_guess_ly` functions in plotly
 - Ensure that all of the `vis_*` family functions have the same internal arguments where appropriate:
-  - E.g., vis_dat, doesn't need to same clustering missingness options (or...does it?)
-- move vis_compare into a "dev" branch on github
-- should I export vis_guess?
-- Yes, as I thought that it might be a useful function for users, although I can't think of a good usecase outside of `vis_guess` right now.
-Perhaps it might be best for `guess_type` to be unexported?
 
@@
   - vis_miss
-  - vis_guess
-- Then, in a subsequent release, I will add the interactive functions:
+- Then, in a subsequent release, I will add the functions:
+  - vis_guess (speed improvements required)
+  - vis_compare
   - vis_dat_ly
@@
   - vis_guess_ly
-  - vis_compare
-- And also look into speed improvements for `vis_guess`
+  - vis_compare_ly
 
@@
 Do you think this is OK?
+
+- note, find all "?" "..." and "^", and paste them down here.

### 2017-07-03 README.Rmd

@@
 
-```{r, echo = FALSE}
+```{r setup, echo = FALSE}
 
@@
 
-```{r eval = FALSE}
+```{r installation, eval = FALSE}
 # install.packages("devtools")
@@
 
-There are also experimental which are described in more detail in the vignette ["Experimental Features of visdat](http://www.njtierney.com/visdat/articles/experimental_features.html).
+There are also experimental functions, which include guessing values,  comparing two dataframes, and adding interactivity.
+These are described in detail in the vignette ["Experimental Features of visdat](http://www.njtierney.com/visdat/articles/experimental_features.html).
 

### 2017-07-03 data-raw/create-typical-data-large.R
change data to typical_data_large to help with tab completion

@@ -1,0 +1,6 @@
+# the following code was used to create this data
+library(wakefield)
+set.seed(1214)
+typical_data_large <- r_data_theme(n = 300, data_theme = "the_works")
+
+use_data(typical_data_large)

### 2017-07-03 R/guess_type.R
cutting down the package to minimal features for CRAN 0.1.0

@@ -1,96 +1,0 @@
-#' Guess the type of each individual cell in a dataframe
-#'
-#' `guess_type` is used by vis_guess to guess cell elements, like `fingerprint`.
-#'
-#' @param x is a vector of values you want to guess
-#'
-#' @return a character vector that describes the suspected class. e.g., "10" is
-#'   an integer, "20.11" is a double, "text" is character, etc.
-#'
-#' @examples
-#' \dontrun{
-#' guess_type(1)
-#'
-#' guess_type("x")
-#'
-#' guess_type(c("1", "0L"))
-#'
-#' purrr::map_df(iris, guess_type)
-#' }
-guess_type <- function(x){
-
-  # since
-  # readr::collector_guess(NA,
-  #                        locale_ = readr::locale())
-  #
-  # returns "character", use an ifelse to identify NAs
-  #
-  # Basically, this is fast way to check individual elements of a vector
-  # I'd like to use purrr::map for this but I couldn't get it to work without
-  # writing more function calls, which slowed it down, by a factor of about 3.
-  # So this is faster, for the moment.
-
-  output <- character(length(x))
-  nas <- is.na(x)
-
-  output[!nas] <- vapply(FUN = readr::guess_parser,
-                         X = x[!nas],
-                         FUN.VALUE = character(1))
-  output[nas] <- NA
-  output
-
-}
-#'
-#' all.equal(guess_df_1(iris), guess_df_2(iris))
-#'
-#' iris %>%
-#'   gather %>%
-#'   guess_vector()
-#'
-#' messy_df %>%
-#'
-#' mb_df <-
-#' microbenchmark::microbenchmark(
-#'   guess_df_1(iris),
-#'   guess_df_2(iris)
-#'   )
-#'
-#' # then do this:
-#' #
-#' #     new_function <- function(x) purrr::dmap(x, ~purrr::map_chr(., guess_type))
-#' #
-#' #     or maybe make it an S3 method?
-#' #
-#' # guess.data.frame
-#'
-#' # > foo <- c(NA, "10", "10.1", "10/01/2001")
-#' # > guess_type(foo)
-#' # [1] NA          "character" "character"
-#' # [4] "character"
-#' #
-#' # > purrr::map_chr(foo, guess_type)
-#' # [1] NA          "integer"   "double"
-#' # [4] "character"
-#' #
[…]

### 2017-07-03 R/internals.R

@@ -35,7 +35,7 @@ vis_extract_value_ <- function(x){
 
 }
 
-#' (Internal) Create a boilerploate for visualisations of the vis_ family
+#' (Internal) Create a boilerplate for visualisations of the vis_ family
 #'
 #' @param x a dataframe in longformat as transformed by `vis_gather_` and
 #'   `vis_extract_value`.
@@ -91,15 +91,14 @@ label_col_missing_pct <- function(x,
 
 #' Add a specific palette to a visdat plot
 #'
-#'
-#'
 #' @param vis_plot visdat plot created using vis_gather_, vis_extract_value
 #'   and vis_create_
-#' @param palette character "default", "qual" or "cb_safe". "default" (the default)
-#'   provides the stock ggplot scale for separating the colours. "qual" uses an experimen
-#'   al qualitative colour scheme for providing distinct colours for each Type. "cb_safe"
-#'   is a set of colours that are appropriate for those with colourblindness. "qual" and
-#'   "cb_safe" are drawn from http://colorbrewer2.org/.
+#' @param palette character "default", "qual" or "cb_safe". "default" (the
+#'   default) provides the stock ggplot scale for separating the colours. "qual"
+#'   uses an experimental qualitative colour scheme for providing distinct
+#'   colours for each Type. "cb_safe" is a set of colours that are appropriate
+#'   for those with colourblindness. "qual" and "cb_safe" are drawn from
+#'   http://colorbrewer2.org/.
 #'
 #' @return a visdat plot with a particular palette
 #'

### 2017-07-03 R/typical_data_large.R

@@ -58,4 +58,4 @@
 #'   \item{Zip}{a zip code they have visited}
 #' }
 #'
-"typical_larger_data"
+"typical_data_large"

### 2017-07-03 R/vis_compare.R

@@ -1,92 +1,0 @@
-#' Visually compare two dataframes and see where they are different.
-#'
-#' `vis_compare`, like the other `vis_*` families, gives an at-a-glance ggplot
-#'   of a dataset, but in this case, hones in on visualising **two** different
-#'   dataframes of the same dimension, and so takes two dataframes as arguments.
-#'
-#' @param df1 The first dataframe to compare
-#'
-#' @param df2 The second dataframe to compare to the first.
-#'
-#' @return `ggplot2` object displaying which values in each data frame are
-#'   present in each other, and which are not.
-#'
-#' @seealso [vis_miss()] [vis_dat()] [vis_miss_ly()] [vis_compare()]
-#'
-#' @examples
-#'
-#' # make a new dataset of iris that contains some NA values
-#' iris_diff <- iris
-#' iris_diff[1:10, 1:2] <- NA
-#'
-#' vis_compare(iris, iris_diff)
-#' @export
-vis_compare <- function(df1,
-                        df2){
-
-  message("vis_compare is in BETA! If you have suggestions or errors\npost an issue at https://github.com/njtierney/visdat/issues")
-
-  # could add a parameter, "sort_match", to help with
-  # sort_match logical TRUE/FALSE.
-  # TRUE arranges the columns in order of most matches.
-
-  # make a TRUE/FALSE matrix of the data.
-  # Tells us whether it is the same (true) as the other dataset, or not (false)
-
-  if (!identical(dim(df1), dim(df2))){
-    stop("Dimensions of df1 and df2 are not the same. vis_compare currently only handles dataframes of identical dimensions.")
-  }
-
-  v_identical <- Vectorize(identical)
-
-  df_diff <- purrr::map2(df1, df2, v_identical) %>%
-    dplyr::as_data_frame()
-
-  d <-
-  df_diff %>%
-    as.data.frame() %>%
-    purrr::map_df(compare_print) %>%
-    vis_gather_() %>%
-    dplyr::mutate(value_df1 = vis_extract_value_(df1),
-                  value_df2 = vis_extract_value_(df2))
-
-  # d$value_df1 <- tidyr::gather_(df1, "variables", "value", names(df1))$value
-  # d$value_df2 <- tidyr::gather_(df2, "variables", "value", names(df2))$value
-
-  # then we plot it
-  ggplot2::ggplot(data = d,
-                  ggplot2::aes_string(
-                    x = "variable",
-                    y = "rows",
-                    # text assists with plotly mouseover
[…]

### 2017-07-03 R/vis_dat.R

@@ -20,7 +20,7 @@
 #' @return `ggplot2` object displaying the type of values in the data frame and
 #'   the position of any missing values.
 #'
-#' @seealso [vis_miss()] [vis_miss_ly()] [vis_guess()] [vis_compare()]
+#' @seealso [vis_miss()]
 #'
 #' @examples
 #'

### 2017-07-03 R/vis_dat_ly.R

@@ -1,35 +1,0 @@
-#' Interactive plot of data in plotly
-#'
-#' This makes `vis_dat` interactive using plotly. At the moment this is an
-#'   initial implementation that just wraps `plotly::ggplotly()` around
-#'   `vis_dat`.
-#'
-#' @param x a data.frame object
-#'
-#' @param sort_type logical TRUE/FALSE. When TRUE (default), it sorts by the
-#'   type in the column to make it easier to see what is in the data
-#'
-#' @param palette character "default", "qual" or "cb_safe". "default" (the
-#'   default) provides the stock ggplot scale for separating the colours.
-#'   "qual" uses an experimental qualitative colour scheme for providing
-#'   distinct colours for each Type. "cb_safe" is a set of colours that are
-#'   appropriate for those with colourblindness. "qual" and "cb_safe" are drawn
-#'   from http://colorbrewer2.org/.
-#'
-#' @return `plotly` interactive plot, similar in appearance to `vis_dat`
-#'
-#' @seealso [vis_dat()] [vis_miss()] [vis_miss_ly()] [vis_guess()] [vis_compare()]
-#'
-#' @export
-#'
-#' @examples
-#'
-#' vis_dat_ly(airquality)
-#'
-vis_dat_ly <- function(x,
-                       sort_type = TRUE,
-                       palette = "default"){
-
-  plotly::ggplotly(vis_dat(x, sort_type, palette))
-
-}

### 2017-07-03 R/vis_guess.R

@@ -1,65 +1,0 @@
-#' Visualise type guess in a data.frame
-#'
-#' `vis_guess` visualises the class of every single individual cell in a
-#'   dataframe and displays it as ggplot object, similar to `vis_dat`. Cells
-#'   are coloured according to what class they are and whether the values are
-#'   missing. `vis_guess` estimates the class of individual elements using
-#'   `readr::guess_parser`.  It may be currently slow on larger datasets.
-#'
-#' @param x a data.frame object
-#' @param palette character "default", "qual" or "cb_safe". "default" (the
-#'   default) provides the stock ggplot scale for separating the colours.
-#'   "qual" uses an experimental qualitative colour scheme for providing
-#'   distinct colours for each Type. "cb_safe" is a set of colours that are
-#'   appropriate for those with colourblindness. "qual" and "cb_safe" are drawn
-#'   from http://colorbrewer2.org/.
-#'
-#' @return `ggplot2` object displaying the guess of the type of values in the
-#'   data frame and the position of any missing values.
-#'
-#' @seealso [vis_miss()] [vis_dat()] [vis_miss_ly()] [vis_compare()]
-#'
-#' @examples
-#'
-#' messy_vector <- c(TRUE,
-#'                  "TRUE",
-#'                  "T",
-#'                  "01/01/01",
-#'                  "01/01/2001",
-#'                  NA,
-#'                  NaN,
-#'                  "NA",
-#'                  "Na",
-#'                  "na",
-#'                  "10",
-#'                  10,
-#'                  "10.1",
-#'                  10.1,
-#'                  "abc",
-#'                  "$%TG")
-#' set.seed(1114)
-#' messy_df <- data.frame(var1 = messy_vector,
-#'                        var2 = sample(messy_vector),
-#'                        var3 = sample(messy_vector))
-#' vis_guess(messy_df)
-#' @export
-vis_guess <- function(x, palette = "default"){
-
-# x = messy_df
-  # suppress warnings here as this is just a note about combining classes
-  d <- suppressWarnings(vis_gather_(x)) %>%
-    dplyr::mutate(valueType = guess_type(valueType)) %>%
-  # value for plotly mouseover
-    dplyr::mutate(value = vis_extract_value_(x))
-
-  # add the boilerplate information
-  vis_plot <- vis_create_(d) +
-      ggplot2::guides(fill = ggplot2::guide_legend(title = "Type")) +
-      # flip the axes, add info for axes
[…]

### 2017-07-03 R/vis_guess_ly.R

@@ -1,50 +1,0 @@
-#' Visualise guess in a data.frame
-#'
-#' `vis_guess_ly`, like the other `vis_*_ly` families, gives an at-a-glance
-#'    interactive plotly plot of a dataset, using the `*` method. In this case,
-#'    an interactive version of `vis_guess`.
-#'
-#' @param x a data.frame object
-#'
-#' @param palette character "default", "qual" or "cb_safe". "default" (the
-#'   default) provides the stock ggplot scale for separating the colours.
-#'   "qual" uses an experimental qualitative colour scheme for providing
-#'   distinct colours for each Type. "cb_safe" is a set of colours that are
-#'   appropriate for those with colourblindness. "qual" and "cb_safe" are drawn
-#'   from http://colorbrewer2.org/.
-#'
-#' @return `plotly` interactive plot, similar in appearance to `vis_guess`
-#'
-#' @seealso [vis_miss()] [vis_dat()] [vis_miss_ly()] [vis_compare()]
-#'
-#' @examples
-#'
-#' messy_vector <- c(TRUE,
-#'                  "TRUE",
-#'                  "T",
-#'                  "01/01/01",
-#'                  "01/01/2001",
-#'                  NA,
-#'                  NaN,
-#'                  "NA",
-#'                  "Na",
-#'                  "na",
-#'                  "10",
-#'                  10,
-#'                  "10.1",
-#'                  10.1,
-#'                  "abc",
-#'                  "$%TG")
-#' set.seed(1114)
-#' messy_df <- data.frame(var1 = messy_vector,
-#'                        var2 = sample(messy_vector),
-#'                        var3 = sample(messy_vector))
-#' vis_guess_ly(messy_df)
-#'
-#' @export
-
-vis_guess_ly <- function(x, palette = "default"){
-
-  plotly::ggplotly(vis_guess(x, palette))
-
-}

### 2017-07-03 R/vis_miss.R

@@ -22,7 +22,7 @@
 #' @return `ggplot2` object displaying the position of missing values in the
 #'   dataframe, and the percentage of values missing and present.
 #'
-#' @seealso [vis_miss_ly()] [vis_dat()] [vis_guess()] [vis_compare()]
+#' @seealso [vis_dat()]
 #'
 #' @examples
 #'

### 2017-07-03 R/vis_miss_ly.R

@@ -1,49 +1,0 @@
-#' Interactive plot of missingness in plotly
-#'
-#' This is a basic implementation of vis_miss in plotly, some work is still
-#'   needed to fix the legend, but it is several (read, about 3000x) faster
-#'   than doing `ggplotly(vis_miss(data))`.
-#'
-#' @param x a dataframe
-#'
-#' @return `plotly` interactive plot, similar in appearance to `vis_miss`
-#'
-#' @seealso [vis_miss()] [vis_dat()] [vis_guess()] [vis_compare()]
-#'
-#' @export
-#'
-#' @examples
-#'
-#' vis_miss_ly(airquality)
-#' plotly::ggplotly(vis_dat(airquality))
-#'
-vis_miss_ly <- function(x){
-
-  x.na <- is.na(x)
-  d <- x.na
-  n <- nrow(x)
-  # printing for row numbers and column names
-  rows <- rep(1:nrow(x),ncol(x))
-  vars <- rep(colnames(x), each = nrow(x))
-
-  categories <- sort(unique(c(d)))
-  m <- matrix(match(d, categories), nrow = n)
-  txt <- matrix(paste(sprintf("value = %s", as.matrix(x)),
-                      sprintf("variables = %s", vars),
-                      sprintf("rows = %s", rows),
-                      sep = "<br />"),
-                nrow = n)
-
-  plotly::plot_ly(z = m,
-          x = names(x),
-          text = txt,
-          hoverinfo = "text",
-          type = "heatmap",
-          colors = c("grey80", "grey20"),
-          colorbar = list(ticktext = c("Present","Missing"),
-                          tickvals = seq_along(categories))
-  ) %>% plotly::layout(xaxis = list(title = ""))
-
-}
-
-

### 2017-07-03 R/visdat-package.r

@@ -10,9 +10,6 @@
 #' \itemize{
 #'   \item [vis_dat()]
 #'   \item [vis_miss()]
-#'   \item [vis_guess()]
-#'   \item [vis_compare()]
-#'   \item [vis_miss_ly()]
 #' }
 #'
 #' Learn more about visdat at \url{www.njtierney.com/visdat/articles/using_visdat.html}

### 2017-07-03 tests/testthat/test-guess-type.R

@@ -1,15 +1,0 @@
-context("Guess type")
-
-
-test_that("guess_element correctly identifies individual elements", {
-  expect_equal(visdat:::guess_type(TRUE), "logical")
-  expect_equal(visdat:::guess_type(T), "logical")
-  expect_equal(visdat:::guess_type("TRUE"), "logical")
-  expect_equal(visdat:::guess_type("T"), "logical")
-  expect_equal(visdat:::guess_type("10"), "integer")
-  expect_equal(visdat:::guess_type(10), "integer")
-  expect_equal(visdat:::guess_type("10.1"), "double")
-  expect_equal(visdat:::guess_type(10.1), "double")
-  expect_equal(visdat:::guess_type("abc"), "character")
-  expect_equal(visdat:::guess_type("$%TG"), "character")
-})

### 2017-07-03 tests/testthat/test-vis-compare.R

@@ -1,23 +1,0 @@
-context("comparing vis")
-
-# make a new dataset of iris that contains some NA values
-
-iris_diff <- iris
-iris_diff[1:10, 1:2] <- NA
-
-vis_compare_plot <- vis_compare(iris, iris_diff)
-
-test_that("vis_compare works",{
-vdiffr::expect_doppelganger("vis_compare vanilla", vis_compare_plot)
-})
-
-iris_add <- iris
-iris_add$extra <- 1
-
-test_that("vis_compare will not accept two dataframes of differing dims",{
-  expect_error(
-    vis_compare(iris, iris_add))
-})
-
-  # stop("Dimensions of df1 and df2 are not the same. Unfortunately vis_compare
-  #      does not handles dataframes of the exact same dimension.")

### 2017-07-03 tests/testthat/test-vis-guess.R

@@ -1,18 +1,0 @@
-context("vis_guess")
-
-# try out all the options
-vis_guess_plot <- vis_guess(typical_data)
-vis_guess_plot_cb_safe <- vis_guess(typical_data, palette = "cb_safe")
-vis_guess_plot_qual <- vis_guess(typical_data, palette = "qual")
-vis_guess_plot_default <- vis_guess(typical_data, palette = "default")
-
-test_that("vis_guess creates the right plot",{
-  vdiffr::expect_doppelganger("vis_guess vanilla", vis_guess_plot)
-  vdiffr::expect_doppelganger("vis_guess cb safe", vis_guess_plot_cb_safe)
-  vdiffr::expect_doppelganger("vis_guess qual", vis_guess_plot_qual)
-  vdiffr::expect_doppelganger("vis_guess default", vis_guess_plot_default)
-})
-
-test_that("vis_guess fails when the wrong palette is provided",{
-  testthat::expect_error(vis_guess(typical_data, palette = "wat"))
-})

### 2017-07-04 README.Rmd
cutting down the package to minimal features for CRAN 0.1.0
move readme figures into man/

@@
 ---
-output:
-  md_document:
-    variant: markdown_github
+output: github_document
 ---
@@
   comment = "#>",
-  fig.path = "README-figs/README-"
+  fig.path = "man/figures/README-"
 )
@@
 <!-- Add an appVeyor badge -->
-[![Travis-CI Build Status](https://travis-ci.org/njtierney/visdat.svg?branch=master)](https://travis-ci.org/njtierney/visdat) [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/njtierney/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/njtierney/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/njtierney/visdat/master.svg)](https://codecov.io/github/njtierney/visdat?branch=master)
+
+[![Travis Build Status](https://travis-ci.org/njtierney/visdat.svg?branch=master)](https://travis-ci.org/njtierney/visdat)[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/njtierney/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/njtierney/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/njtierney/visdat/master.svg)](https://codecov.io/github/njtierney/visdat?branch=master)
 
@@
 You can read more about visdat in the vignette, ["using visdat""](http://www.njtierney.com/visdat/articles/using_visdat.html).
-
-There are also experimental functions, which include guessing values,  comparing two dataframes, and adding interactivity.
-These are described in detail in the vignette ["Experimental Features of visdat](http://www.njtierney.com/visdat/articles/experimental_features.html).
 
@@
 
-vis_dat(typical_larger_data)
+vis_dat(typical_data_large)
 

### 2017-07-24 README.Rmd
misc CRAN updates

@@
 
-```{r setup, echo = FALSE}
+```{r, echo = FALSE}
 
@@
 
-[![Travis Build Status](https://travis-ci.org/njtierney/visdat.svg?branch=master)](https://travis-ci.org/njtierney/visdat)[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/njtierney/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/njtierney/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/njtierney/visdat/master.svg)](https://codecov.io/github/njtierney/visdat?branch=master)
+[![Travis Build Status](https://travis-ci.org/njtierney/visdat.svg?branch=master)](https://travis-ci.org/njtierney/visdat)[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/njtierney/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/njtierney/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/njtierney/visdat/master.svg)](https://codecov.io/github/njtierney/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)
 
@@
 
-```{r installation, eval = FALSE}
+```{r eval = FALSE}
 # install.packages("devtools")
@@
 You can read more about visdat in the vignette, ["using visdat""](http://www.njtierney.com/visdat/articles/using_visdat.html).
+
+There are also experimental which are described in more detail in the vignette ["Experimental Features of visdat](http://www.njtierney.com/visdat/articles/experimental_features.html).
 

### 2017-07-24 R/internals.R

@@ -1,0 +1,175 @@
+#' (Internal) Gather rows into a format appropriate for grid visualisation
+#'
+#' @param x a dataframe
+#'
+#' @return data.frame gathered to have columns "variables", "valueType", and a
+#'   row id called "rows".
+#'
+vis_gather_ <- function(x){
+  x %>%
+  dplyr::mutate(rows = seq_len(nrow(x))) %>%
+    tidyr::gather_(key_col = "variable",
+                   value_col = "valueType",
+                   gather_cols = names(.)[-length(.)])
+}
+
+
+#' (Internal) Add values of each row as a column
+#'
+#' This adds information about each row, so that when called by plotly, the
+#'   values are made visible on hover. Warnings are suppressed because tidyr
+#'   gives a warning about type coercion, which is fine.
+#'
+#' @param x dataframe created from `vis_gather_`
+#'
+#' @return the x dataframe with the added column `value`.
+#'
+vis_extract_value_ <- function(x){
+
+  suppressWarnings(
+    tidyr::gather_(x,
+                   "variable",
+                   "value",
+                   names(x))$value
+  )
+
+}
+
+#' (Internal) Create a boilerplate for visualisations of the vis_ family
+#'
+#' @param x a dataframe in longformat as transformed by `vis_gather_` and
+#'   `vis_extract_value`.
+#'
+#' @return a ggplot object
+#'
+vis_create_ <- function(x){
+
+  ggplot2::ggplot(data = x,
+                  ggplot2::aes_string(x = "variable",
+                                      y = "rows",
+                                    # text assists with plotly mouseover
+                                    text = "value")) +
+  ggplot2::geom_raster(ggplot2::aes_string(fill = "valueType")) +
+  ggplot2::theme_minimal() +
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45,
+                                                     vjust = 1,
+                                                     hjust = 1)) +
+  ggplot2::labs(x = "",
+                y = "Observations") +
+    # flip the axes
+    ggplot2::scale_y_reverse() +
+    ggplot2::theme(axis.text.x = ggplot2::element_text(hjust = 0.5))
+
+}
+
+#' (Internal) Create labels for the columns containing the \% missing data
+#'
+#' @param x data.frame
+#' @param col_order_index the order of the columns
+#'
+#' @return data.frame containing the missingness percent down to 0.1 percent
+#'
+#' @export
[…]

### 2017-07-24 R/typical_data_large.R

@@ -54,13 +54,4 @@
 #'   \item{Year}{significant year to that individuals}
 #'   \item{Zip}{a zip code they have visited}
 #' }
-#'
-#' @note the following code was used to create this data
-#'
-#' library(wakefield)
-#'
-#' set.seed(1214)
-#' typical_larger_data <- r_data_theme(n = 300, data_theme = "the_works")
-#'
-#'
-"typical_larger_data"
+"typical_data_large"

### 2017-07-24 R/vis_dat.R

@@ -1,49 +1,59 @@
 #' Visualises a data.frame to tell you what it contains.
 #'
-#' `vis_dat` gives you an at-a-glance ggplot object of what is inside a dataframe. Cells are coloured according to what class they are and whether the values are missing. As `vis_dat` returns a ggplot object, it is very easy to customize and change labels, and customize the plot
+#' `vis_dat` gives you an at-a-glance ggplot object of what is inside a
+#'   dataframe. Cells are coloured according to what class they are and whether
+#'   the values are missing. As `vis_dat` returns a ggplot object, it is very
+#'   easy to customize and change labels, and customize the plot
 #'
 #' @param x a data.frame object
 #'
-#' @param sort_type logical TRUE/FALSE. When TRUE (default), it sorts by the type in the column to make it easier to see what is in the data
+#' @param sort_type logical TRUE/FALSE. When TRUE (default), it sorts by the
+#'   type in the column to make it easier to see what is in the data
 #'
-#' @param palette character "default", "qual" or "cb_safe". "default" (the default) provides the stock ggplot scale for separating the colours. "qual" uses an experimental qualitative colour scheme for providing distinct colours for each Type. "cb_safe" is a set of colours that are appropriate for those with colourblindness. "qual" and "cb_safe" are drawn from http://colorbrewer2.org/.
+#' @param palette character "default", "qual" or "cb_safe". "default" (the
+#'   default) provides the stock ggplot scale for separating the colours.
+#'   "qual" uses an experimental qualitative colour scheme for providing
+#'   distinct colours for each Type. "cb_safe" is a set of colours that are
+#'   appropriate for those with colourblindness. "qual" and "cb_safe" are drawn
+#'   from http://colorbrewer2.org/.
 #'
-#' @param flip logical TRUE, will flip the axis labels to be on top, resembling a dataframe
+#' @param warn_large_data logical default is TRUE
 #'
-#' @return `ggplot2` object displaying the type of values in the data frame and the position of any missing values.
+#' @param large_data_size integer default is 900000, this can be changed.
 #'
-#' @seealso [vis_miss()] [vis_miss_ly()] [vis_guess()] [vis_compare()]
+#' @return `ggplot2` object displaying the type of values in the data frame and
+#'   the position of any missing values.
+#'
+#' @seealso [vis_miss()]
 #'
 #' @examples
 #'
 #' vis_dat(airquality)
 #'
[…]

### 2017-07-24 R/vis_miss.R

@@ -1,20 +1,32 @@
 #' Visualise a data.frame to display missingness.
 #'
-#' `vis_miss` provides an at-a-glance ggplot of the missingness inside a dataframe, colouring cells according to missingness, where black indicates a missing cell and grey indicates a present cell. As it returns a ggplot object, it is very easy to customize and change labels, and so on.
+#' `vis_miss` provides an at-a-glance ggplot of the missingness inside a
+#'   dataframe, colouring cells according to missingness, where black indicates
+#'   a missing cell and grey indicates a present cell. As it returns a ggplot
+#'   object, it is very easy to customize and change labels.
 #'
 #' @param x a data.frame
 #'
-#' @param cluster logical TRUE/FALSE. TRUE specifies that you want to use hierarchical clustering (mcquitty method) to arrange rows according to missingness. FALSE specifies that you want to leave it as is
+#' @param cluster logical. TRUE specifies that you want to use hierarchical
+#'   clustering (mcquitty method) to arrange rows according to missingness.
+#'   FALSE specifies that you want to leave it as is.
 #'
-#' @param sort_miss logical TRUE/FALSE. TRUE arranges the columns in order of missingness
+#' @param sort_miss logical. TRUE arranges the columns in order of missingness
 #'
-#' @param show_perc logical TRUE/FALSE. TRUE now adds in the \% of missing/complete data in the whole dataset into the legend. Default value is TRUE
+#' @param show_perc logical. TRUE now adds in the \% of missing/complete data
+#'   in the whole dataset into the legend. Default value is TRUE.
 #'
-#' @param flip logical if TRUE, will flip the axis labels to be on top, resembling a dataframe
+#'@param show_perc_col logical. TRUE adds in the \% missing data in a given
+#'  column into the x axis. Can be disabled with FALSE
 #'
-#' @return `ggplot2` object displaying the position of missing values in the dataframe, and the percentage of values missing and present.
+#' @param warn_large_data logical default is TRUE
 #'
-#' @seealso [vis_miss_ly()] [vis_dat()] [vis_guess()] [vis_compare()]
+#' @param large_data_size integer default is 900000, this can be changed.
+#'
+#' @return `ggplot2` object displaying the position of missing values in the
+#'   dataframe, and the percentage of values missing and present.
+#'
+#' @seealso [vis_dat()]
 #'
 #' @examples
 #'
@@ -24,22 +36,27 @@
 #'
[…]

### 2017-07-24 data-raw/create-typical-data-large.R
change to typical_data_large from typical_larger_data

@@ -1,0 +1,10 @@
+# the following code was used to create this data
+library(wakefield)
+set.seed(1214)
+typical_data_large <- r_data_theme(n = 300, data_theme = "the_works")
+
+typical_data_large <- typical_data_large %>%
+  mutate(Income = as.factor(Income),
+         Age = as.character(Age))
+
+use_data(typical_data_large)

### 2017-07-24 tests/testthat/test-vis-compare.R
update pkg docs

@@ -8,6 +8,7 @@ iris_diff[1:10, 1:2] <- NA
 vis_compare_plot <- vis_compare(iris, iris_diff)
 
 test_that("vis_compare works",{
+  skip_on_cran()
 vdiffr::expect_doppelganger("vis_compare vanilla", vis_compare_plot)
 })
 

### 2017-07-24 tests/testthat/test-vis-dat.R

@@ -1,16 +1,28 @@
 context("vis_dat")
 
 # try out all the options
-vis_dat_plot <- vis_dat(airquality)
-vis_dat_plot_sort_type <- vis_dat(airquality, sort_type = FALSE)
-vis_dat_plot_pal_qual <- vis_dat(airquality, palette = "qual")
-vis_dat_plot_pal_qual <- vis_dat(airquality, palette = "cb_safe")
-vis_dat_plot_flip <- vis_dat(airquality, flip = TRUE)
+vis_dat_plot <- vis_dat(typical_data)
+vis_dat_plot_sort_type <- vis_dat(typical_data, sort_type = FALSE)
+vis_dat_plot_pal_qual <- vis_dat(typical_data, palette = "qual")
+vis_dat_plot_pal_cb <- vis_dat(typical_data, palette = "cb_safe")
 
 test_that("vis_dat creates the right plot",{
-  vdiffr::expect_doppelganger("vis_dat vanilla", vis_dat_plot)
-  vdiffr::expect_doppelganger("vis_dat sort_type", vis_dat_plot)
-  vdiffr::expect_doppelganger("vis_dat qualitative palette", vis_dat_plot)
-  vdiffr::expect_doppelganger("vis_dat colourblind safe palette", vis_dat_plot)
-  vdiffr::expect_doppelganger("vis_dat flip to dataframe", vis_dat_plot)
+  skip_on_cran()
+  ver <- as.character(gdtools::version_freetype())
+  cat(sprintf("FreeType version: %s\n", ver))
+  vdiffr::expect_doppelganger("vis_dat vanilla",
+                              vis_dat_plot)
+  vdiffr::expect_doppelganger("vis_dat sort_type",
+                              vis_dat_plot_sort_type)
+  vdiffr::expect_doppelganger("vis_dat qualitative palette",
+                              vis_dat_plot_pal_qual)
+  vdiffr::expect_doppelganger("vis_dat colourblind safe palette",
+                              vis_dat_plot_pal_cb)
+})
+
+
+test_that("vis_dat fails when the wrong palette is provided",{
+  ver <- as.character(gdtools::version_freetype())
+  cat(sprintf("FreeType version: %s\n", ver))
+  testthat::expect_error(vis_dat(typical_data, palette = "wat"))
 })

### 2017-07-24 tests/testthat/test-vis-guess.R

@@ -4,5 +4,6 @@ context("vis_guess")
 vis_guess_plot <- vis_guess(airquality)
 
 test_that("vis_miss_ly creates the right plot",{
+  skip_on_cran()
   vdiffr::expect_doppelganger("vis_guess vanilla", vis_guess_plot)
 })

### 2017-07-24 tests/testthat/test-vis-large.R

@@ -1,0 +1,12 @@
+context("Warn for large data")
+
+big_df <- matrix(rnorm(100000),
+                 nrow = 1000,
+                 ncol = 1000) %>%
+  as.data.frame()
+
+test_that("vis_dat and vis_miss throw warnings when the DF is above size",{
+  cat("nrows = ",nrow(big_df), "ncols = ", ncol(big_df))
+  expect_error(vis_dat(big_df))
+  expect_error(vis_miss(big_df))
+})

### 2017-07-24 tests/testthat/test-vis-miss-ly.R

@@ -3,6 +3,7 @@ context("vis_miss_ly")
 vis_miss_ly_plot <- vis_miss_ly(airquality)
 
 test_that("vis_miss_ly creates the right plot",{
+  skip_on_cran()
 vdiffr::expect_doppelganger("vis_miss_ly",
                             vis_miss_ly_plot)
 })

### 2017-07-24 tests/testthat/test-vis-miss.R

@@ -1,16 +1,17 @@
 context("vis_miss")
 
 # try out all the options
-vis_miss_plot <- vis_miss(airquality)
-vis_miss_plot_cluster <- vis_miss(airquality, cluster = TRUE)
-vis_miss_plot_sort_rows <- vis_miss(airquality, sort_miss = TRUE)
-vis_miss_plot_show_perc <- vis_miss(airquality, show_perc = FALSE)
-vis_miss_plot_show_flip <- vis_miss(airquality, flip = TRUE)
+vis_miss_plot <- vis_miss(typical_data)
+vis_miss_plot_cluster <- vis_miss(typical_data, cluster = TRUE)
+vis_miss_plot_sort_rows <- vis_miss(typical_data, sort_miss = TRUE)
+vis_miss_plot_show_perc <- vis_miss(typical_data, show_perc = FALSE)
 
 test_that("vis_miss creates the right plot",{
+  skip_on_cran()
+  ver <- as.character(gdtools::version_freetype())
+  cat(sprintf("FreeType version: %s\n", ver))
   vdiffr::expect_doppelganger("vis_miss vanilla", vis_miss_plot)
   vdiffr::expect_doppelganger("vis_miss cluster", vis_miss_plot_cluster)
   vdiffr::expect_doppelganger("vis_miss sort rows", vis_miss_plot_sort_rows)
   vdiffr::expect_doppelganger("vis_miss show percent", vis_miss_plot_show_perc)
-  vdiffr::expect_doppelganger("vis_miss show flip", vis_miss_plot_show_flip)
 })

### 2017-07-24 tests/testthat/test-visdat-internals.R

@@ -1,0 +1,39 @@
+context("Internals")
+
+test_vis_gather_ <- suppressWarnings(vis_gather_(typical_data))
+
+suppressWarnings(
+  test_old_gather <- typical_data %>%
+    dplyr::mutate(rows = seq_len(nrow(.))) %>%
+    tidyr::gather_(key_col = "variable",
+                   value_col = "valueType",
+                   gather_cols = names(.)[-length(.)])
+)
+
+test_that("vis_gather_ returns the same as previous",{
+
+  expect_equal(test_vis_gather_,
+               test_old_gather)
+
+})
+
+d_old <- typical_data %>%
+  purrr::map_df(fingerprint) %>%
+  vis_gather_()
+
+suppressWarnings(
+  d_old$value <-  tidyr::gather_(typical_data,
+                                 "variables",
+                                 "value", names(typical_data))$value
+)
+
+d_new <-
+  typical_data %>%
+  purrr::map_df(fingerprint) %>%
+  vis_gather_() %>%
+  dplyr::mutate(value = vis_extract_value_(typical_data))
+# get the values here so plotly can make them visible
+
+test_that("vis_extract_value performs the same as old method",{
+  expect_equal(d_old,d_new)
+})

### 2017-07-24 paper/response.Rmd

@@
 - I agree with you re the palette argument being provided for all plots, although I'm not sure how this could work for vis_miss.
-But perhaps I should allow for users to insert their own pallete for this.
+But perhaps I should allow for users to insert their own palette for this.
 Personally I think that there should be a standard palette where the common classes (character, factor, integer, numeric, datetime, etc) each have their own distinct colour, and other classes are given different factors.
@@
 
-- Thank you for picking up on this, I've updated the main descriotion file now
+- Thank you for picking up on this, I've updated the main description file now
 
@@
 - refactoring repeated code into internal functions to more clearly express what was happening, and avoiding repetition.
--
+This resulted in the creation of internal functions `vis_gather_`, `vis_extract_value`, and `vis_create_`
+- adding percentage missing to the column names in `vis_miss`
 
-Still remaining to do is:
+- For the moment I have decided to submit visdat to CRAN, with the main functions: 
+  - vis_dat
+  - vis_miss
 
-- Make a separate branch for the experimental features.
+Then, in a separate branch, there will be all of the features:
   - vis_compare()
@@
   - vis_*_ly() family
-- complete vis_guess() for release
-    - Adding the fast Rcpp code for guess_type
-- Add interactivity using pure plotly
-  - create `vis_dat_ly`, `vis_miss_ly`, and `vis_guess_ly` functions in plotly
-- Ensure that all of the `vis_*` family functions have the same internal arguments where appropriate:
 
-- For the moment I have decided to put visdat onto CRAN, with the main functions
-being: 
-  - vis_dat
-  - vis_miss
-- Then, in a subsequent release, I will add the functions:
-  - vis_guess (speed improvements required)
-  - vis_compare
-  - vis_dat_ly
-  - vis_miss_ly
-  - vis_guess_ly
-  - vis_compare_ly
+There is still more work to be done on these functions. Specifically:
 
-> example(vis_guess) / example(vis_compare) : gives warning message
+- vis_guess. This needs to be made much faster, and will incorporate the C code from @seaaan to do so.
+- vis_*_ly() family will be written in pure plotly, for speed.
 
-- These now gives the message:
+I now summarise some of the changes that I wanted to confirm with the reviewers, @seaaan and @batpigandme:
+
+1. Is the warning message for vis_guess/vis_compare code OK? `example(vis_guess) / example(vis_compare)`
 
@@
 ```
-Do you think this is OK?
 
-- note, find all "?" "..." and "^", and paste them down here.
+2.
+The default sort option.
+I think that the default sort_type option should be to put all of the similar columns together - to me this makes it easier to identify what is similar, and what is different; when I start looking at a dataset I usually want to know what variables are what class.
+I am willing to change this, however.
+
+3.
+Maximum row size issue.
+This is an interesting one!
+It was brought to my attention [here](https://github.com/njtierney/visdat/issues/32), by @MilesMcBain, but I couldn't replicate this issue.
+Similarly, I can actually display the nycflights data.
+So it seems that this is to do with the fact that computations power has a bit to do with it.
+I think that option 1 is the best case here, but option 2 actually gets a little bit difficult, because this leads to another issue of effectively downsampling a dataset whilst maintaining data structure.
+This could be a problem, say, if I did perform down-sampling and then the user looked at the data with visdat and plotly, they might get a mis-represented sample.
+Likewise, if there is a very large proportion of missing data, this might be difficult to preserve exactly.
+Ways around this might include some very explicit warning messages, or perhaps even a note directly on the plot that the data have been downsampled.
+^At the moment one solution forward is to have a check first, something like:
+
+```r
+vis_dat(really_large_data)
+
+Warning: This dataset contains over 10,000 rows, which may take a while to render. Are you sure that you want to continue?
+
+1: Yes
+2. No (You should use a sample of your data)
+```
+
+4.
+Speed for vis_guess.
+@seaaan wrote some nice code to make this work much much faster, and I think that this is a fantastic idea!
+But I'm not sure on the best way to move this into production of `visdat`, since it sounds like we would need to borrow a chunk of code from readr and then modify it.
+I wonder if perhaps the solution here is to do a pull request for this specific feature in `readr`?
+
+5. Can the two issues below be replicated?
+
+> I don't know why, but in the plotly section, a vis_dat() (non-interactive) plot of df_test appears between the first two interactive plots.
+I can't explain it, hopefully it's just a weird quirk on my computer.
+
+> I'm getting an error 83 because pandoc-citeproc is missing. I believe this is something I'm missing on my machine, and, thus you don't need to worry about. But, just in case, I'm letting you know!

### 2017-07-24 R/visdat-package.r
rerun build_site and set eval = FALSE for experimental features vignette

@@ -22,4 +22,4 @@
 NULL
 
 if(getRversion() >= "2.15.1")  utils::globalVariables(c("."))
-globalVariables(c("valueGuess"))
+globalVariables(c("valueGuess", "valueType"))

### 2017-07-28 README.Rmd
update badges

@@
 
-[![Travis Build Status](https://travis-ci.org/njtierney/visdat.svg?branch=master)](https://travis-ci.org/njtierney/visdat)[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/njtierney/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/njtierney/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/njtierney/visdat/master.svg)](https://codecov.io/github/njtierney/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)
+[![Travis Build Status](https://travis-ci.org/njtierney/visdat.svg?branch=master)](https://travis-ci.org/njtierney/visdat)[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/njtierney/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/njtierney/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/njtierney/visdat/master.svg)](https://codecov.io/github/njtierney/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](http://cran.rstudio.com/web/packages/visdat/index.html)
 

### 2017-07-28 README.Rmd
Fix links to vignettes
Note only fixed in Rmd, not compiled

@@
 
-You can read more about visdat in the vignette, ["using visdat""](http://www.njtierney.com/visdat/articles/using_visdat.html).
+You can read more about visdat in the vignette, ["using visdat"](http://visdat.njtierney.com/articles/using_visdat.html).
 
 There are also experimental functions, which include guessing values,  comparing two dataframes, and adding interactivity.
-These are described in detail in the vignette ["Experimental Features of visdat](http://www.njtierney.com/visdat/articles/experimental_features.html).
+These are described in detail in the vignette ["Experimental Features of visdat"](http://visdat.njtierney.com/articles/experimental_features.html).
 

### 2017-07-28 README.Rmd
add rOpenSci badge!

@@
 
-[![Travis Build Status](https://travis-ci.org/njtierney/visdat.svg?branch=master)](https://travis-ci.org/njtierney/visdat)[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/njtierney/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/njtierney/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/njtierney/visdat/master.svg)](https://codecov.io/github/njtierney/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](http://cran.rstudio.com/web/packages/visdat/index.html)
+[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/onboarding/issues/87)[![Travis Build Status](https://travis-ci.org/njtierney/visdat.svg?branch=master)](https://travis-ci.org/njtierney/visdat)[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/njtierney/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/njtierney/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/njtierney/visdat/master.svg)](https://codecov.io/github/njtierney/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](http://cran.rstudio.com/web/packages/visdat/index.html)
 

### 2017-07-28 R/internals.R
remove label_col_missing_pct as described in https://github.com/ropensci/onboarding/issues/87

@@ -69,7 +69,6 @@ vis_create_ <- function(x){
 #'
 #' @return data.frame containing the missingness percent down to 0.1 percent
 #'
-#' @export
 label_col_missing_pct <- function(x,
                                   col_order_index){
 

### 2017-07-28 README.Rmd
added acknowledgement of Noam Ross, Sean Huges, and Mara Averick (https://github.com/ropensci/onboarding/issues/87)

@@
 
-Thanks also to Carson Sievert for writing the code that combined `plotly` with `visdat`, and for Noam Ross for suggesting this in the first place.
+Thanks to Carson Sievert for writing the code that combined `plotly` with `visdat`, and for Noam Ross for suggesting this in the first place.
+
+Finally thank you to [rOpenSci](https://github.com/ropensci) and it's amazing [onboarding process](https://github.com/ropensci/onboarding), this process has made visdat a much better package, thanks to the editor Noam Ross (@noamross), and the reviewers Sean Hughes (@seaaan) and Mara Averick (@batpigandme).

### 2017-08-01 README.Rmd
add note that visdat is on CRAN, add the rOpenSci footer :tada:

@@
 
+visdat is available on CRAN
+
+```{r install-cran, eval = FALSE}
+
+install.packages("visdat")
+
+```
+
+If you would like to use the development version, install from github with:
+
 ```{r installation, eval = FALSE}
+
 # install.packages("devtools")
-
-devtools::install_github("njtierney/visdat")
+devtools::install_github("ropensci/visdat")
 ```
@@
 Finally thank you to [rOpenSci](https://github.com/ropensci) and it's amazing [onboarding process](https://github.com/ropensci/onboarding), this process has made visdat a much better package, thanks to the editor Noam Ross (@noamross), and the reviewers Sean Hughes (@seaaan) and Mara Averick (@batpigandme).
+
+[![ropensci_footer](https://ropensci.org/public_images/ropensci_footer.png)](https://ropensci.org)

### 2017-08-01 README.Rmd
change the badges to rOpenSci not njtierney

@@
 
-<!-- add a TravisCI badge -->
-<!-- Add an appVeyor badge -->
-
-[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/onboarding/issues/87)[![Travis Build Status](https://travis-ci.org/njtierney/visdat.svg?branch=master)](https://travis-ci.org/njtierney/visdat)[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/njtierney/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/njtierney/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/njtierney/visdat/master.svg)](https://codecov.io/github/njtierney/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](http://cran.rstudio.com/web/packages/visdat/index.html)
+[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/onboarding/issues/87)[![Travis Build Status](https://travis-ci.org/ropensci/visdat.svg?branch=master)](https://travis-ci.org/ropensci/visdat)[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ropensci/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/ropensci/visdat/master.svg)](https://codecov.io/github/ropensci/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](http://cran.rstudio.com/web/packages/visdat/index.html)
 

### 2017-08-01 paper/paper.Rmd
minor changes to paper - add a sentence to vis_dat and vis_miss, and change style slightly to be more appropriate for a journal.

@@
 affiliations:
- - name: Queensland University of Technology
+ - name: Monash University
    index: 1
-date: 09 January 2017
+date: 01 August 2017
 bibliography: paper.bib
@@
 
-When you get a new data set, you need to look at the data to get a sense of what it contains and potential problems with it.
-You read in your data, but you then need to look at it to understand what you need to do to make it ready for analysis.
-A phrase that often arises in this process is "looking at the data".
-It is helpful to look more closely at what this means.
-You can look at the first six rows of data, the head of the data:
+Reading in a new dataset means looking at the data to get a sense of what it contains, and potential problems and challenges with the data to get it analysis ready.
+"Looking at the data" can mean different things.
+Often times you look at the first six rows of data, the head of the data:
 
@@
 
-Another alterantive is `glimpse`, from the `dplyr` package [@dplyr]
+Or you can `glimpse` the data, using the `dplyr` package [@dplyr]
 
@@
 
-This has a better printing method that gives us more information about the data, displaying doubles and a factor.
-However, we don't always have data like the canonical iris dataset.
+This shows that there are 150 observations, and 5 variables, which are doubles and a factor, along with some of the values in the data.
+However, we usually don't have data like the canonical iris dataset.
 Let's take a look at some data that might be a bit more typical of "messy" data.
@@
 
-Looking at this, you might then ask:
+Looking at this, you might then ask: "Isn't it odd that Income is a factor?
+And Age is a character?".
+And you might start to wonder what else is different, what else changed?
+It might be a bit unclear where to go from there.
+Do you do some exploratory plots of the data?
+What if the plot looks weird because the data has strange types?
+What if the data has other strange features?
 
-> Isn't it odd that Income is a factor? And Age is a character?
-
-And you might start to wonder what else is different, what else changed?
-
-And it might be a bit unclear where to go from there.
-Do you plot the data?
-Why does my plot look wierd?
-What are these other strange features in the data?
-The `visdat` package provides visualisations of an entire dataframe at once.
+There is a need for a tool to do preliminary data visualisation, to identify these problems early.
+The `visdat` package provides this, creating visualisations of an entire dataframe at once.
 Initially inspired by [`csv-fingerprint`](https://github.com/setosa/csv-fingerprint), `visdat` provides tools to create heatmap-like visualisations of an entire dataframe.
@@
 
-`vis_dat()` helps explore the data class structure and missingness:
+`vis_dat()` helps explore the data class structure and missingness, by displaying the class for each variable and the missing data, and presenting the plot in an intuitive way - it reads top down as the data would.
+The columns are grouped by similar class as well, to put similarly classified data types together.
 
@@
 
-And the `vis_miss` function provides a custom plot for missing data.
+`vis_miss()` focuses on just displaying the present and missing data, again reading top down, but also displaying the percent of missing data in each column, and overall.
 
-```{r}
+```{r vis-miss}
 
@@
 
-These functions provide useful tools to help "get a look at the data", using principled visualisation techniques.
+These functions provide useful tools to help "get a look at the data", using preliminary visualisation techniques.
 The plots are built using ggplot2 [@ggplot2], which provides a consistent and powerful framework for visualisations.

### 2017-08-01 README.Rmd
Add DOI badge :badger emoji:

@@
 
-[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/onboarding/issues/87)[![Travis Build Status](https://travis-ci.org/ropensci/visdat.svg?branch=master)](https://travis-ci.org/ropensci/visdat)[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ropensci/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/ropensci/visdat/master.svg)](https://codecov.io/github/ropensci/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](http://cran.rstudio.com/web/packages/visdat/index.html)
+[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/onboarding/issues/87)[![Travis Build Status](https://travis-ci.org/ropensci/visdat.svg?branch=master)](https://travis-ci.org/ropensci/visdat)[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ropensci/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/ropensci/visdat/master.svg)](https://codecov.io/github/ropensci/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](http://cran.rstudio.com/web/packages/visdat/index.html)[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.837274.svg)](https://doi.org/10.5281/zenodo.837274)
 

### 2017-08-11 README.Rmd
update pkgdown site

--- 2017-08-01 README.Rmd
@@
 
-[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/onboarding/issues/87)[![Travis Build Status](https://travis-ci.org/ropensci/visdat.svg?branch=master)](https://travis-ci.org/ropensci/visdat)[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ropensci/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/ropensci/visdat/master.svg)](https://codecov.io/github/ropensci/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](http://cran.rstudio.com/web/packages/visdat/index.html)
+[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/onboarding/issues/87)[![Travis Build Status](https://travis-ci.org/ropensci/visdat.svg?branch=master)](https://travis-ci.org/ropensci/visdat)[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ropensci/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/ropensci/visdat/master.svg)](https://codecov.io/github/ropensci/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/naniar)](https://cran.r-project.org/package=naniar)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](http://cran.rstudio.com/web/packages/visdat/index.html)[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.837274.svg)](https://doi.org/10.5281/zenodo.837274)
 

### 2017-08-11 paper/paper.Rmd
update pkgdown site

@@
 ---
-title: 'visdat: Visualising whole data frames'
+title: 'visdat: Visualising Whole Data Frames'
 tags:
@@
 
+# Acknowlegements
+
+I would like to thank the two reviewers, Mara Averick and Sean Hughes, for their helpful suggestions that resulted in a much better package.
+
 # References

### 2017-08-16 README.Rmd
update README

--- 2017-08-01 README.Rmd
@@
 
-[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/onboarding/issues/87)[![Travis Build Status](https://travis-ci.org/ropensci/visdat.svg?branch=master)](https://travis-ci.org/ropensci/visdat)[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ropensci/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/ropensci/visdat/master.svg)](https://codecov.io/github/ropensci/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](http://cran.rstudio.com/web/packages/visdat/index.html)
+[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/onboarding/issues/87)[![Travis Build Status](https://travis-ci.org/ropensci/visdat.svg?branch=master)](https://travis-ci.org/ropensci/visdat)[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ropensci/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/ropensci/visdat/master.svg)](https://codecov.io/github/ropensci/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/naniar)](https://cran.r-project.org/package=naniar)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](http://cran.rstudio.com/web/packages/visdat/index.html)[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.837274.svg)](https://doi.org/10.5281/zenodo.837274)[![JOSS status](http://joss.theoj.org/papers/c85f57adbc565b064fb4bfc9b59a1b2a/status.svg)](http://joss.theoj.org/papers/c85f57adbc565b064fb4bfc9b59a1b2a)
 

### 2017-08-16 paper/paper.Rmd
update JOSS paper and adjust figures

--- 2017-05-10 paper/paper.Rmd
@@
 ---
-title: 'visdat: Visualising whole data frames'
+title: 'visdat: Visualising Whole Data Frames'
+authors:
+- affiliation: 1
+  name: Nicholas Tierney
+  orcid: 0000-0003-1460-8722
+date: "01 August 2017"
+output:
+  pdf_document: default
+  html_document:
+    keep_md: yes
+bibliography: paper.bib
 tags:
@@
 - exporatory data analysis
-authors:
- - name: Nicholas Tierney
-   orcid: 0000-0003-1460-8722
-   affiliation: 1
 affiliations:
- - name: Queensland University of Technology
-   index: 1
-date: 09 January 2017
-bibliography: paper.bib
-output: 
-  html_document: 
-    keep_md: yes
+- index: 1
+  name: Monash University
 ---
+
+JOSS requirements:
+
+- A summary describing the high-level functionality and purpose of the software for a diverse, non-specialist audience
+- A clear statement of need that illustrates the purpose of the software
+- A list of key references including a link to the software archive
+- Mentions (if applicable) of any ongoing research projects using the software or recent scholarly publications enabled by it
 
@@
 
-When you get a new data set, you need to look at the data to get a sense of what it contains and potential problems with it.
-You read in your data, but you then need to look at it to understand what you need to do to make it ready for analysis.
-A phrase that often arises in this process is "looking at the data".
-It is helpful to look more closely at what this means.
-You can look at the first six rows of data, the head of the data:
+When you receive a new dataset you need to look at the data to get a sense of what is in it and understand potential problems and challenges to get it analysis-ready.
+"Taking a look at the data" can mean different things.
+For example, examining statistical summaries (minimum, maximum, mean, interquartile range), finding missing values, checking data formatting, creating graphical summaries such as histograms, scatterplots, boxplots, and more.
 
-```{r head-iris}
+When handling typical "messy" data, these preliminary exploratory steps can become difficult to perform when you encounter data that is not what you expect.
+For example, income might be a factor instead of numeric, date could be a number not a character string or a date class, or values could be missing when they shouldn't be.
+Often times you realise that you have expectations of the data, but it can be hard to know what they are exactly until they are a problem.
+This is sort of like how you might not think to buy more lightbulbs until one goes out - when you go to use the data in an exploratory scatter plot, or in a preliminary model, you don't always realise your data is in the wrong format until that moment.
+What you need is a birds eye view of the data that tells you what classes are in the dataframe, and where the missing data are.
 
-head(iris)
+`visdat` is an R [@Rcore] package that provides a tool to "get a look at the data", creating heatmap-like visualisations of an entire dataframe - classes in the data, missing values, and comparisons between two datasets.
+`visdat` takes inspiration from [`csv-fingerprint`](https://github.com/setosa/csv-fingerprint), and is powered by ggplot2 [@ggplot2], which provides a consistent and powerful framework for visualisations that can be extended if needed.
 
-```
-
-Another alterantive is `glimpse`, from the `dplyr` package [@dplyr]
-
-```{r glimpse}
-
-dplyr::glimpse(iris)
-
-```
-
-This has a better printing method that gives us more information about the data, displaying doubles and a factor.
-However, we don't always have data like the canonical iris dataset.
-Let's take a look at some data that might be a bit more typical of "messy" data.
-
-```{r visdat-glimpse}
-
-library(visdat)
-dplyr::glimpse(typical_data)
-
-```
-
-Looking at this, you might then ask:
-
-> Isn't it odd that Income is a factor? And Age is a character?
-
-And you might start to wonder what else is different, what else changed?
-
-And it might be a bit unclear where to go from there.
-Do you plot the data?
-Why does my plot look wierd?
-What are these other strange features in the data?
-The `visdat` package provides visualisations of an entire dataframe at once.
-Initially inspired by [`csv-fingerprint`](https://github.com/setosa/csv-fingerprint), `visdat` provides tools to create heatmap-like visualisations of an entire dataframe.
-`visdat` is an R [@Rcore] package provides 2 main functions `vis_dat` and `vis_miss`.
-
-`vis_dat()` helps explore the data class structure and missingness:
+Plots are presented in an intuitive way, reading top down, just like your data. Below is a plot of some example typical data containing missing values and data of a variety of classes.
 
 ```{r load-data}
-
+library(visdat)
 vis_dat(typical_data)
@@
 
-And the `vis_miss` function provides a custom plot for missing data.
+`visdat` will continue to be improved over time, to improve speed in computation and also add interactive plots.
 
-```{r}
+# Acknowlegements
 
-vis_miss(typical_data)
-
-```
-
-These functions provide useful tools to help "get a look at the data", using principled visualisation techniques.
-The plots are built using ggplot2 [@ggplot2], which provides a consistent and powerful framework for visualisations.
-This means that users can customise and extend graphics from visdat very easily.
+I would like to thank the two reviewers, Mara Averick and Sean Hughes, for their helpful suggestions that resulted in a much better package, and rOpenSci for providing the support of the onboarding package review that facilitated these improvements.
 

### 2017-08-16 paper/paper.Rmd
Minor wording changes in JOSS paper

@@
 output:
-  pdf_document: default
   html_document:
     keep_md: yes
+  pdf_document: default
 bibliography: paper.bib
@@
 "Taking a look at the data" can mean different things.
-For example, examining statistical summaries (minimum, maximum, mean, interquartile range), finding missing values, checking data formatting, creating graphical summaries such as histograms, scatterplots, boxplots, and more.
+For example, examining statistical summaries (minimum, maximum, mean, inter-quartile range), finding missing values, checking data formatting, creating graphical summaries such as histograms, scatter plots, box plots, and more.
 
-When handling typical "messy" data, these preliminary exploratory steps can become difficult to perform when you encounter data that is not what you expect.
+When handling typical real-world data, these preliminary exploratory steps can become difficult to perform when values are not what you expect.
 For example, income might be a factor instead of numeric, date could be a number not a character string or a date class, or values could be missing when they shouldn't be.
-Often times you realise that you have expectations of the data, but it can be hard to know what they are exactly until they are a problem.
-This is sort of like how you might not think to buy more lightbulbs until one goes out - when you go to use the data in an exploratory scatter plot, or in a preliminary model, you don't always realise your data is in the wrong format until that moment.
-What you need is a birds eye view of the data that tells you what classes are in the dataframe, and where the missing data are.
+Often times you realise that you have expectations of the data, which are hard to realise until they are a problem.
+This is similar to how you might not think to buy more light bulbs until one goes out: when you use data in an exploratory scatter plot, or a preliminary model, you often don't realise your data is in the wrong format until that moment.
+What is needed is a birds eye view of the data, which tells you what classes are in the dataframe, and where the missing data are.
 
-`visdat` is an R [@Rcore] package that provides a tool to "get a look at the data", creating heatmap-like visualisations of an entire dataframe - classes in the data, missing values, and comparisons between two datasets.
+`visdat` is an R [@Rcore] package that provides a tool to "get a look at the data", creating heatmap-like visualisations of an entire dataframe providing information on classes in the data, missing values, and also comparisons between two datasets.
 `visdat` takes inspiration from [`csv-fingerprint`](https://github.com/setosa/csv-fingerprint), and is powered by ggplot2 [@ggplot2], which provides a consistent and powerful framework for visualisations that can be extended if needed.
 
-Plots are presented in an intuitive way, reading top down, just like your data. Below is a plot of some example typical data containing missing values and data of a variety of classes.
+Plots are presented in an intuitive way, reading top down, just like your data.
+Below is a plot using `vis_dat()` of some example typical data containing missing values and data of a variety of classes.
 
@@
 
-`visdat` will continue to be improved over time, to improve speed in computation and also add interactive plots.
+`visdat` will continue to be improved over time, to improve speed in computation and improve interactive plotting.
 

### 2017-08-17 paper/paper.Rmd

@@
 
-When you receive a new dataset you need to look at the data to get a sense of what is in it and understand potential problems and challenges to get it analysis-ready.
+When you receive a new dataset you need to look at the data to get a sense of what is in it, and understand potential problems and challenges to get it analysis-ready.
 "Taking a look at the data" can mean different things.
-For example, examining statistical summaries (minimum, maximum, mean, inter-quartile range), finding missing values, checking data formatting, creating graphical summaries such as histograms, scatter plots, box plots, and more.
+For example: examining statistical summaries (minimum, maximum, mean, inter-quartile range), finding missing values, checking data formatting, creating graphical summaries such as histograms, scatter plots, box plots, and more.
 
@@
 For example, income might be a factor instead of numeric, date could be a number not a character string or a date class, or values could be missing when they shouldn't be.
-Often times you realise that you have expectations of the data, which are hard to realise until they are a problem.
-This is similar to how you might not think to buy more light bulbs until one goes out: when you use data in an exploratory scatter plot, or a preliminary model, you often don't realise your data is in the wrong format until that moment.
+Often times, you discover that you had expectations of the data, which are hard to realise until they are a problem.
+This is similar to how one might not think to buy more light bulbs until one goes out: when you use data in an exploratory scatter plot, or a preliminary model, you often don't realise your data is in the wrong format until that moment.
 What is needed is a birds eye view of the data, which tells you what classes are in the dataframe, and where the missing data are.
 
-`visdat` is an R [@Rcore] package that provides a tool to "get a look at the data", creating heatmap-like visualisations of an entire dataframe providing information on classes in the data, missing values, and also comparisons between two datasets.
-`visdat` takes inspiration from [`csv-fingerprint`](https://github.com/setosa/csv-fingerprint), and is powered by ggplot2 [@ggplot2], which provides a consistent and powerful framework for visualisations that can be extended if needed.
+`visdat` is an R [@Rcore] package that provides a tool to "get a look at the data" by creating heatmap-like visualisations of an entire dataframe, which provides information on: classes in the data, missing values, and also comparisons between two datasets.
+`visdat` takes inspiration from [`csv-fingerprint`](https://github.com/setosa/csv-fingerprint), and is powered by ggplot2 [@ggplot2], which provides a consistent, powerful framework for visualisations that can be extended if needed.
 
-Plots are presented in an intuitive way, reading top down, just like your data.
-Below is a plot using `vis_dat()` of some example typical data containing missing values and data of a variety of classes.
+Plots are presented in an intuitive way, reading top down, just like your data. Below is a plot using `vis_dat()` of some typical data containing missing values and data of a variety of classes.
 

### 2017-08-17 paper/paper.Rmd
update paper for JOSS submission

@@
 ---
-
-JOSS requirements:
-
-- A summary describing the high-level functionality and purpose of the software for a diverse, non-specialist audience
-- A clear statement of need that illustrates the purpose of the software
-- A list of key references including a link to the software archive
-- Mentions (if applicable) of any ongoing research projects using the software or recent scholarly publications enabled by it
 

### 2017-08-21 README.Rmd
update DOI badges in README and other docs

--- 2017-08-11 README.Rmd
@@
 
-[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/onboarding/issues/87)[![Travis Build Status](https://travis-ci.org/ropensci/visdat.svg?branch=master)](https://travis-ci.org/ropensci/visdat)[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ropensci/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/ropensci/visdat/master.svg)](https://codecov.io/github/ropensci/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/naniar)](https://cran.r-project.org/package=naniar)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](http://cran.rstudio.com/web/packages/visdat/index.html)[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.837274.svg)](https://doi.org/10.5281/zenodo.837274)
+[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/onboarding/issues/87)[![JOSS status](http://joss.theoj.org/papers/c85f57adbc565b064fb4bfc9b59a1b2a/status.svg)](http://joss.theoj.org/papers/c85f57adbc565b064fb4bfc9b59a1b2a)[![DOI](https://zenodo.org/badge/50553382.svg)](https://zenodo.org/badge/latestdoi/50553382)
+
+[![Travis Build Status](https://travis-ci.org/ropensci/visdat.svg?branch=master)](https://travis-ci.org/ropensci/visdat)[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ropensci/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/ropensci/visdat/master.svg)](https://codecov.io/github/ropensci/visdat?branch=master)
+
+[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/naniar)](https://cran.r-project.org/package=naniar)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](http://cran.rstudio.com/web/packages/visdat/index.html)
 

### 2018-06-04 README.Rmd
update site and add logo
add lifecycle badge
update README with updated badge, fixes #80

@@
 
-# visdat
+# visdat <img src="man/figures/visdat-logo.png" align="right" />
 
@@
 
-[![Travis Build Status](https://travis-ci.org/ropensci/visdat.svg?branch=master)](https://travis-ci.org/ropensci/visdat)[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ropensci/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/ropensci/visdat/master.svg)](https://codecov.io/github/ropensci/visdat?branch=master)
+[![Travis Build Status](https://travis-ci.org/ropensci/visdat.svg?branch=master)](https://travis-ci.org/ropensci/visdat)[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/njtierney/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/njtierney/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/ropensci/visdat/master.svg)](https://codecov.io/github/ropensci/visdat?branch=master)
 
 [![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/naniar)](https://cran.r-project.org/package=naniar)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](http://cran.rstudio.com/web/packages/visdat/index.html)
+
+[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
 

### 2018-06-05 README.Rmd
fixed appveyor badge - #80

@@
 
-[![Travis Build Status](https://travis-ci.org/ropensci/visdat.svg?branch=master)](https://travis-ci.org/ropensci/visdat)[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/njtierney/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/njtierney/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/ropensci/visdat/master.svg)](https://codecov.io/github/ropensci/visdat?branch=master)
+[![Travis Build Status](https://travis-ci.org/ropensci/visdat.svg?branch=master)](https://travis-ci.org/ropensci/visdat)[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/ropensci/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/ropensci/visdat/master.svg)](https://codecov.io/github/ropensci/visdat?branch=master)
 

### 2018-06-06 README.Rmd
use project status badge instead of lifecycle

@@
 
-[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
+[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
 

### 2018-07-02 R/vis_compare.R
remove interactive components, which are now in the `interactive` branch - as per #88

@@ -11,7 +11,7 @@
 #' @return `ggplot2` object displaying which values in each data frame are
 #'   present in each other, and which are not.
 #'
-#' @seealso [vis_miss()] [vis_dat()] [vis_miss_ly()] [vis_compare()]
+#' @seealso [vis_miss()] [vis_dat()] [vis_compare()] [vis_expect()]
 #'
 #' @examples
 #'

### 2018-07-02 R/vis_compare_ly.R

@@ -1,31 +1,0 @@
-#' Interactive vis_compare compare two dataframes to find differences.s
-#'
-#' `vis_compare_ly`, like the other `vis_*_ly` families, gives an at-a-glance
-#'    interactive plotly plot of a dataset, using the `*` method. In thi case,
-#'    an interactive version of `vis_compare`, which is used to visualise and
-#'    compare **two** different dataframes of the same dimension, and so takes
-#'    two dataframes as arguments.
-#'
-#' @param df1 The first dataframe to compare
-#'
-#' @param df2 The second dataframe to compare to the first.
-#'
-#' @return `plotly` interactive plot, similar in appearance to `vis_compare`
-#'
-#' @seealso [vis_dat()] [vis_miss()] [vis_miss_ly()] [vis_dat_ly()]
-#'   [vis_compare()] [vis_guess()]
-#'
-#' @examples
-#'
-#' # make a new dataset of iris that contains some NA values
-#' iris_diff <- iris
-#' iris_diff[1:10, 1:2] <- NA
-#'
-#' vis_compare_ly(iris, iris_diff)
-#' @export
-vis_compare_ly <- function(df1,
-                           df2){
-
-  plotly::ggplotly(vis_compare(df1, df2))
-
-}

### 2018-07-02 R/vis_dat_ly.R

@@ -1,36 +1,0 @@
-#' Interactive plot of data in plotly
-#'
-#' This makes `vis_dat` interactive using plotly. At the moment this is an
-#'   initial implementation that just wraps `plotly::ggplotly()` around
-#'   `vis_dat`.
-#'
-#' @param x a data.frame object
-#'
-#' @param sort_type logical TRUE/FALSE. When TRUE (default), it sorts by the
-#'   type in the column to make it easier to see what is in the data
-#'
-#' @param palette character "default", "qual" or "cb_safe". "default" (the
-#'   default) provides the stock ggplot scale for separating the colours.
-#'   "qual" uses an experimental qualitative colour scheme for providing
-#'   distinct colours for each Type. "cb_safe" is a set of colours that are
-#'   appropriate for those with colourblindness. "qual" and "cb_safe" are drawn
-#'   from http://colorbrewer2.org/.
-#'
-#' @return `plotly` interactive plot, similar in appearance to `vis_dat`
-#'
-#' @seealso [vis_dat()] [vis_miss()] [vis_miss_ly()] [vis_guess()]
-#'   [vis_compare()]
-#'
-#' @export
-#'
-#' @examples
-#'
-#' vis_dat_ly(airquality)
-#'
-vis_dat_ly <- function(x,
-                       sort_type = TRUE,
-                       palette = "default"){
-
-  plotly::ggplotly(vis_dat(x, sort_type, palette))
-
-}

### 2018-07-02 R/vis_guess.R

@@ -17,7 +17,7 @@
 #' @return `ggplot2` object displaying the guess of the type of values in the
 #'   data frame and the position of any missing values.
 #'
-#' @seealso [vis_miss()] [vis_dat()] [vis_miss_ly()] [vis_compare()]
+#' @seealso [vis_miss()] [vis_dat()] [vis_compare()] [vis_expect()]
 #'
 #' @examples
 #'

### 2018-07-02 R/vis_miss_ly.R

@@ -1,61 +1,0 @@
-#' Interactive plot of missingness in plotly
-#'
-#' This is a basic implementation of vis_miss in plotly, some work is still
-#'   needed to fix the legend, but it is several (read, about 3000x) faster
-#'   than doing `ggplotly(vis_miss(data))`.
-#'
-#' @param x a dataframe
-#'
-#' @return `plotly` interactive plot, similar in appearance to `vis_miss`
-#'
-#' @seealso [vis_miss()] [vis_dat()] [vis_guess()] [vis_compare()]
-#'
-#' @export
-#'
-#' @examples
-#'
-#' vis_miss_ly(airquality)
-#' plotly::ggplotly(vis_dat(airquality))
-#'
-vis_miss_ly <- function(x){
-
-  x.na <- is.na(x)
-  d <- x.na
-  n <- nrow(x)
-  # printing for row numbers and column names
-  rows <- rep(1:n,ncol(x))
-  vars <- rep(colnames(x), each = n)
-
-  categories <- sort(unique(c(d)))
-  m <- matrix(match(d, categories), nrow = n)
-  txt <- matrix(paste(sprintf("value = %s", as.matrix(x)),
-                      sprintf("variables = %s", vars),
-                      sprintf("rows = %s", rows),
-                      sep = "<br />"),
-                nrow = n)
-
-  plotly::plot_ly(z = m,
-                  x = names(x),
-                  text = txt,
-                  hoverinfo = "text",
-                  type = "heatmap",
-                  colorscale = discretize_colorscale(c("#cccccc","#333333"))) %>%
-    plotly::colorbar(tickmode = "array",
-                     ticktext = c("Present","Missing"),
-                     tickvals = 1:2,
-                     len = 0) %>%
-    plotly::layout(
-      xaxis = list(side = "top"),
-      yaxis = list(autorange = "reversed"),
-      legend = list(orientation = 'h'))
-
-}
-
-discretize_colorscale <- function(palette, granularity = 100) {
-  n <- length(palette)
-  colorscale <- data.frame(range = seq(0, 1, length.out = n*granularity),
-                           color = rep(palette, each = granularity))
-
-  setNames(colorscale, NULL)
-}
-

### 2018-07-02 R/visdat-package.r

@@ -12,7 +12,7 @@
 #'   \item [vis_miss()]
 #'   \item [vis_guess()]
 #'   \item [vis_compare()]
-#'   \item [vis_miss_ly()]
+#'   \item [vis_expect()]
 #' }
 #'
 #' Learn more about visdat at \url{www.njtierney.com/visdat/articles/using_visdat.html}

### 2018-07-02 tests/testthat/test-vis-ly.R

@@ -1,28 +1,0 @@
-context("vis_ly")
-
-vis_miss_ly_plot <- vis_miss_ly(typical_data)
-
-test_that("vis_miss_ly creates the right plot",{
-  expect_is(vis_miss_ly_plot,"plotly")
-})
-
-# vis_dat_ly_plot <- vis_dat_ly(typical_data)
-#
-# test_that("vis_dat_ly creates the right plot",{
-#   expect_is(vis_dat_ly_plot,"plotly")
-# })
-#
-# vis_guess_ly_plot <- vis_guess_ly(typical_data)
-#
-# test_that("vis_guess_ly creates the right plot",{
-#   expect_is(vis_guess_ly_plot,"plotly")
-# })
-#
-# iris_diff <- iris
-# iris_diff[1:10, 1:2] <- NA
-#
-# vis_compare_ly_plot <- vis_compare_ly(iris, iris_diff)
-#
-# test_that("vis_compare_ly creates the right plot",{
-#   expect_is(vis_compare_ly_plot,"plotly")
-# })

### 2018-07-02 tests/testthat/test-vis-miss-ly.R

@@ -1,9 +1,0 @@
-context("vis_miss_ly")
-
-vis_miss_ly_plot <- vis_miss_ly(airquality)
-
-test_that("vis_miss_ly creates the right plot",{
-  skip_on_cran()
-vdiffr::expect_doppelganger("vis_miss_ly",
-                            vis_miss_ly_plot)
-})

### 2018-07-02 README.Rmd
updated and read over README and vignette

@@
 
-Initially inspired by [`csv-fingerprint`](https://github.com/setosa/csv-fingerprint), `vis_dat` helps you visualise a dataframe and "get a look at the data" by displaying the variable classes in a dataframe as a plot with `vis_dat`, and getting a brief look into missing data patterns using `vis_miss`.
+Initially inspired by
+[`csv-fingerprint`](https://github.com/setosa/csv-fingerprint), `vis_dat` helps
+you visualise a dataframe and "get a look at the data" by displaying the
+variable classes in a dataframe as a plot with `vis_dat`, and getting a brief
+look into missing data patterns using `vis_miss`.
 
-The name `visdat` was chosen as I think in the future it could be integrated with [`testdat`](https://github.com/ropensci/testdat).
-The idea being that first you visualise your data (`visdat`), then you run tests from `testdat` to fix them.
+`visdat` has 6 functions:
 
-There are two main commands in the `visdat` package:
+- `vis_dat()` visualises a dataframe showing you what the classes of the columns
+are, and also displaying the missing data.
 
-- `vis_dat()` visualises a dataframe showing you what the classes of the columns are, and also displaying the missing data.
+- `vis_miss()` visualises just the missing data, and allows for missingness to
+be clustered and columns rearranged. `vis_miss()` is similar to
+`missing.pattern.plot` from the
+[`mi`](https://cran.r-project.org/web/packages/mi/index.html) package.
+Unfortunately `missing.pattern.plot` is no longer in the `mi` package (as of
+14/02/2016).
 
-- `vis_miss()` visualises just the missing data, and allows for missingness to be clustered and columns rearranged.
-`vis_miss()` is similar to `missing.pattern.plot` from the [`mi`](https://cran.r-project.org/web/packages/mi/index.html) package.
-Unfortunately `missing.pattern.plot` is no longer in the `mi` package (as of 14/02/2016).
+- `vis_compare()` visualise differences between two dataframes of the same 
+  dimensions
+
+- `vis_expect()` visualise where certain conditions hold true in your data
+
+- `vis_cor()` visualise the correlation of variables in a nice heatmap
+
+- `vis_guess()` visualise the individual class of earch value in your data
 
@@
 
-There are also experimental functions, which include guessing values,  comparing two dataframes, and adding interactivity.
-These are described in detail in the vignette ["Experimental Features of visdat"](http://visdat.njtierney.com/articles/experimental_features.html).
-
-Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.
+Please note that this project is released with a 
+[Contributor Code of Conduct](CONDUCT.md). By participating in this project you 
+agree to abide by its terms.
 
@@
 
-Let's see what's inside the `airquality` dataset from base R, which contains information about daily air quality measurements in New York from May to September 1973.
-More information about the dataset can be found with `?airquality`.
+Let's see what's inside the `airquality` dataset from base R, which contains
+information about daily air quality measurements in New York from May to
+September 1973. More information about the dataset can be found with
+`?airquality`.
 
@@
 
-The plot above tells us that R reads this dataset as having numeric and integer values, with some missing data in `Ozone` and `Solar.R`.
-The classes are represented on the legend, and missing data represented by grey.
-The column/variable names are listed on the x axis.
-
-By default, `vis_dat` sorts the columns according to the type of the data in the vectors. You can turn this off by setting `sort_type = FALSE`.
-
-```{r vis-dat-aq-sort-type}
-
-vis_dat(airquality, 
-        sort_type = FALSE)
-
-```
-
-### With many kinds of data
-
-To demonstrate what visdat looks like when you have different kinds of data, we can look at the dataset `typical_data`, provided within `visdat`, and created with the excellent [`wakefield` package](https://github.com/trinker/wakefield).
-
-```{r vis-dat-typical-data}
-
-vis_dat(typical_data)
-
-```
-
-We can also look into using even wider data, looking at `typical_larger_data`
-
-```{r vis-dat-typical-larger-data}
-
-vis_dat(typical_data_large)
-
-```
+The plot above tells us that R reads this dataset as having numeric and integer
+values, with some missing data in `Ozone` and `Solar.R`. The classes are
+represented on the legend, and missing data represented by grey. The
+column/variable names are listed on the x axis.
 
@@
 
-We can explore the missing data further using `vis_miss()`.
+We can explore the missing data further using `vis_miss()`:
 
@@
 
-The percentages of missing/complete in `vis_miss` are accurate to 1 decimal place.
+Percentages of missing/complete in `vis_miss` are accurate to 1 decimal place.
 
-You can cluster the missingness by setting `cluster = TRUE`.
+You can cluster the missingness by setting `cluster = TRUE`:
 
@@
 
-The columns can also just be arranged by columns with most missingness, by setting `sort_miss = TRUE`.
+Columns can also be arranged by columns with most missingness, by setting
+`sort_miss = TRUE`:
 
@@
 
-`vis_miss` indicates when there is a very small amount of missing data at <0.1% missingness.
+`vis_miss` indicates when there is a very small amount of missing data at <0.1%
+missingness:
 
@@
 
-`vis_miss` will also indicate when there is no missing data at all.
+`vis_miss` will also indicate when there is no missing data at all:
 
@@
 
+To further explore the missingness structure in a dataset, I recommend the
+[`naniar`](https://github.com/njtierney/naniar) package, which provides more
+general tools for graphical and numerical exploration of missing values.
+
+## Using `vis_compare()`
+
+Sometimes you want to see what has changed in your data. `vis_compare()` displays the differences in two dataframes of the same size. Let's look at an example.
+
+Let's make some changes to the `chickwts`, and compare this new dataset:
+
+```{r vis-compare-iris}
+chickwts_diff <- chickwts
+chickwts_diff[sample(1:nrow(chickwts), 30),sample(1:ncol(chickwts), 2)] <- NA
+
+vis_compare(chickwts_diff, chickwts)
+
+```
+
+Here the differences are marked in blue.
+
+If you try and compare differences when the dimensions are different, you get 
+an ugly error:
+
+```{r vis-compare-error, eval = FALSE}
+
+chickwts_diff_2 <- chickwts
+chickwts_diff_2$new_col <- chickwts_diff_2$weight*2
+
+vis_compare(chickwts, chickwts_diff_2)
+# Error in vis_compare(chickwts, chickwts_diff_2) : 
+#   Dimensions of df1 and df2 are not the same. vis_compare requires dataframes of identical dimensions.
+```
+
+## Using `vis_expect()`
+
+`vis_expect` visualises certain conditions or values in your data. For example,
+If you are not sure whether to expect values greater than 25 in your data
+(airquality), you could write: `vis_expect(airquality, ~.x >= 25)`, and you can
+see if there are times where the  values in your data are greater than or equal
+to 25:
+
+```{r vis-expect}
+
+vis_expect(airquality, ~.x >= 25)
+
+```
+
+This shows the proportion of times that there are values greater than 25, as
+well as the missings.
+
+## Using `vis_cor()`
+
+To make it easy to plot correlations of your data, use `vis_cor`:
+
+```{r vis-cor}
+
+vis_cor(airquality)
+
+```
+
+## Using `vis_guess()`
+
+`vis_guess()` takes a guess at what each cell is. It's best illustrated using
+some messy data, which we'll make here:
+
+```{r create-messy-vec}
+
+messy_vector <- c(TRUE,
+                  T,
+                  "TRUE",
+                  "T",
+                  "01/01/01",
+                  "01/01/2001",
+                  NA,
+                  NaN,
+                  "NA",
+                  "Na",
+                  "na",
+                  "10",
+                  10,
+                  "10.1",
+                  10.1,
+                  "abc",
+                  "$%TG")
+
+set.seed(1114)
+messy_df <- data.frame(var1 = messy_vector,
+                       var2 = sample(messy_vector),
+                       var3 = sample(messy_vector))
+
+```
+
+```{r vis-guess-messy-df, fig.show='hold', out.width='50%'}
+
+vis_guess(messy_df)
+vis_dat(messy_df)
+
+```
+
+So here we see that there are many different kinds of data in your dataframe. As
+an analyst this might be a depressing finding. We can see this comparison above.
+
 # Thank yous
 
-Thank you to Ivan Hanigan who [first commented](http://www.njtierney.com/r/missing%20data/rbloggers/2015/12/01/ggplot-missing-data/#comment-2388263747) this suggestion after I made a blog post about an initial prototype `ggplot_missing`, and Jenny Bryan, whose [tweet](https://twitter.com/JennyBryan/status/679011378414268416) got me thinking about `vis_dat`, and for her code contributions that removed a lot of errors.
+Thank you to Ivan Hanigan who [first
+commented](http://www.njtierney.com/r/missing%20data/rbloggers/2015/12/01/ggplot-missing-data/#comment-2388263747)
+this suggestion after I made a blog post about an initial prototype
+`ggplot_missing`, and Jenny Bryan, whose
+[tweet](https://twitter.com/JennyBryan/status/679011378414268416) got me
+thinking about `vis_dat`, and for her code contributions that removed a lot of
+errors.
 
-Thank you to Hadley Wickham for suggesting the use of the internals of `readr` to make `vis_guess` work.
+Thank you to Hadley Wickham for suggesting the use of the internals of `readr`
+to make `vis_guess` work. Thank you to Miles McBain for his suggestions on how
+to improve `vis_guess`. This resulted in making it at least 2-3 times faster.
+Thanks to Carson Sievert for writing the code that combined `plotly` with
+`visdat`, and for Noam Ross for suggesting this in the first place. Thank you
+also to Earo Wang and Stuart Lee for their help in getting capturing expressions
+in `vis_expect`.
 
-Thank you to Miles McBain for his suggestions on how to improve `vis_guess`. This resulted in making it at least 2-3 times faster.
-
-Thanks to Carson Sievert for writing the code that combined `plotly` with `visdat`, and for Noam Ross for suggesting this in the first place.
-
-Finally thank you to [rOpenSci](https://github.com/ropensci) and it's amazing [onboarding process](https://github.com/ropensci/onboarding), this process has made visdat a much better package, thanks to the editor Noam Ross (@noamross), and the reviewers Sean Hughes (@seaaan) and Mara Averick (@batpigandme).
+Finally thank you to [rOpenSci](https://github.com/ropensci) and it's amazing
+[onboarding process](https://github.com/ropensci/onboarding), this process has
+made visdat a much better package, thanks to the editor Noam Ross (@noamross),
+and the reviewers Sean Hughes (@seaaan) and Mara Averick (@batpigandme).
 

### 2018-07-02 R/vis-compare.R

@@ -18,7 +18,6 @@
 #' # make a new dataset of iris that contains some NA values
 #' aq_diff <- airquality
 #' aq_diff[1:10, 1:2] <- NA
-#'
 #' vis_compare(airquality, aq_diff)
 #' @export
 vis_compare <- function(df1,

### 2018-07-02 R/vis-dat.R

@@ -39,11 +39,11 @@
 #'
 #' vis_dat(airquality)
 #'
+#' \dontrun{
 #' # experimental colourblind safe palette
 #' vis_dat(airquality, palette = "cb_safe")
 #' vis_dat(airquality, palette = "qual")
 #'
-#' \dontrun{
 #' # if you have a large dataset, you might want to try downsampling:
 #' library(nycflight13)
 #' library(dplyr)

### 2018-07-02 R/vis-expect.R

@@ -32,12 +32,10 @@
 #'             NA, NA
 #'             )
 #'
-#' vis_expect(dat_test,
-#'            ~.x == -1)
+#' vis_expect(dat_test, ~.x == -1)
 #'
 #' \dontrun{
-#' vis_expect(airquality,
-#'            ~.x == 5.1)
+#' vis_expect(airquality, ~.x == 5.1)
 #'
 #' # explore some common NA strings
 #'

### 2018-07-02 R/vis-cor.R
fix up package doco and YAML, closes #86

@@ -1,6 +1,6 @@
-#' Create a correlation heatmap ggplot of all of the data
+#' Visualise correlations amongst variables in your data as a heatmap
 #'
-#' Visualises correlations amongst variables using ggplot
+#' Visualise correlations amongst variables in your data as a heatmap
 #'
 #' @param data data.frame
 #' @param cor_method correlation method to use, from `cor`: "a character

### 2018-07-04 R/vis-compare.R
cleaned up some internal function doco

@@ -23,15 +23,15 @@
 vis_compare <- function(df1,
                         df2){
 
-  # could add a parameter, "sort_match", to help with
+  # could add a parameter, sort_match, to help with
   # sort_match logical TRUE/FALSE.
   # TRUE arranges the columns in order of most matches.
 
   # make a TRUE/FALSE matrix of the data.
   # Tells us whether it is the same (true) as the other dataset, or not (false)
 
-  if (!identical(dim(df1), dim(df2))){
-    stop("Dimensions of df1 and df2 are not the same. vis_compare requires dataframes of identical dimensions.")
+  if (!identical(dim(df1), dim(df2))) {
+    stop("vis_compare requires identical dimensions of df1 and df2")
   }
 
   v_identical <- Vectorize(identical)
@@ -39,16 +39,12 @@ vis_compare <- function(df1,
   df_diff <- purrr::map2(df1, df2, v_identical) %>%
     dplyr::as_data_frame()
 
-  d <-
-  df_diff %>%
+  d <- df_diff %>%
     as.data.frame() %>%
     purrr::map_df(compare_print) %>%
     vis_gather_() %>%
     dplyr::mutate(value_df1 = vis_extract_value_(df1),
                   value_df2 = vis_extract_value_(df2))
-
-  # d$value_df1 <- tidyr::gather_(df1, "variables", "value", names(df1))$value
-  # d$value_df2 <- tidyr::gather_(df2, "variables", "value", names(df2))$value
 
   # then we plot it
   ggplot2::ggplot(data = d,
@@ -57,16 +53,11 @@ vis_compare <- function(df1,
                     y = "rows")) +
                     # text assists with plotly mouseover
                     # text = c("value_df1", "value_df2"))) +
-    # this "test code has been removed as ggplot2 version 3.0.0
+    # this test code has been removed as ggplot2 version 3.0.0
     # breaks.
     # Logged in issue https://github.com/ropensci/visdat/issues/89
 
     ggplot2::geom_raster(ggplot2::aes_string(fill = "valueType")) +
-    # change the colour, so that missing is grey, present is black
-    # scale_fill_discrete(name = "",
-    #                     labels = c("Different",
-    #                                "Missing",
-    #                                "Same")) +
     ggplot2::theme_minimal() +
     ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45,
                                      vjust = 1,
@@ -75,7 +66,6 @@ vis_compare <- function(df1,
          y = "Observations",
[…]

### 2018-07-04 README.Rmd
update the README CRAN Badge

@@
 
-[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/naniar)](https://cran.r-project.org/package=naniar)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](http://cran.rstudio.com/web/packages/visdat/index.html)
+[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](http://cran.rstudio.com/web/packages/visdat/index.html)
 

### 2019-03-28 README.Rmd
remove empty lines between package badges, close #117

@@
 
-[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/onboarding/issues/87)[![JOSS status](http://joss.theoj.org/papers/c85f57adbc565b064fb4bfc9b59a1b2a/status.svg)](http://joss.theoj.org/papers/c85f57adbc565b064fb4bfc9b59a1b2a)[![DOI](https://zenodo.org/badge/50553382.svg)](https://zenodo.org/badge/latestdoi/50553382)
-
-[![Travis Build Status](https://travis-ci.org/ropensci/visdat.svg?branch=master)](https://travis-ci.org/ropensci/visdat)[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/ropensci/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/ropensci/visdat/master.svg)](https://codecov.io/github/ropensci/visdat?branch=master)
-
-[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](http://cran.rstudio.com/web/packages/visdat/index.html)
-
-[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
+[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/onboarding/issues/87)[![JOSS status](http://joss.theoj.org/papers/c85f57adbc565b064fb4bfc9b59a1b2a/status.svg)](http://joss.theoj.org/papers/c85f57adbc565b064fb4bfc9b59a1b2a)[![DOI](https://zenodo.org/badge/50553382.svg)](https://zenodo.org/badge/latestdoi/50553382)[![Travis Build Status](https://travis-ci.org/ropensci/visdat.svg?branch=master)](https://travis-ci.org/ropensci/visdat)[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/ropensci/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/ropensci/visdat/master.svg)](https://codecov.io/github/ropensci/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](http://cran.rstudio.com/web/packages/visdat/index.html)[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
 

### 2019-04-03 README.Rmd
use set.seed to make plots not change

@@
 ```{r vis-compare-iris}
+set.seed(2019-04-03-1105)
 chickwts_diff <- chickwts
@@
 
-set.seed(1114)
+set.seed(2019-04-03-1106)
 messy_df <- data.frame(var1 = messy_vector,

### 2019-04-03 tests/testthat/test-vis-large.R
add set.seed to ensure plot stays the same

@@ -1,5 +1,6 @@
 context("Warn for large data")
 
+set.seed(2019-04-03-1104)
 big_df <- matrix(rnorm(100000),
                  nrow = 1000,
                  ncol = 1000) %>%

### 2019-05-19 README.Rmd
Adds vis_value to visualise all values in dataset
* Updates error message when functions need all numeric data. This is for vis_cor and vis_value.
* use dplyr::n()
* use `guess_integer = TRUE` for `guess_type`.
* Flesh out documentation for vis_value

--- 2018-07-04 README.Rmd
@@
 
+## Using `vis_value`
+
+`vis_value()` visualises the values of your data on a 0 to 1 scale.
+
+```{r vis-value}
+vis_value(airquality)
+```
+
+It only works on numeric data, so you might get strange results if you are using   factors:
+
+```{r iris-error, eval = FALSE}
+library(ggplot2)
+vis_value(iris)
+```
+
+```
+data input can only contain numeric values, please subset the data to the numeric values you would like. dplyr::select_if(data, is.numeric) can be helpful here!
+```
+
+So you might need to subset the data beforehand like so:
+
+```{r iris-error-fix}
+library(dplyr)
+
+iris %>%
+  select_if(is.numeric) %>%
+  vis_value()
+```
+
 ## Using `vis_guess()`

### 2019-05-19 R/internals.R

@@ -265,3 +265,13 @@ all_numeric <- function(x, ...){
   all(as.logical(lapply(x, is.numeric)))
 }
 # Can I capture moving from a value to NA, or, from NA to another value?
+
+
+#' Scale a vector between 0 and one.
+#'
+#' @param x numeric vector
+#'
+#' @return numeric vector between 0 and 1
+scale_01 <- function(x) {
+  (x - min(x, na.rm = TRUE)) / diff(range(x, na.rm = TRUE))
+}

### 2019-05-19 R/vis-cor.R

@@ -29,7 +29,7 @@ vis_cor <- function(data,
                     ...){
 
   if (!all_numeric(data)) {
-    stop("data input can only contain numeric values, please subset the data to the numeric values you would like.")
+    stop("data input can only contain numeric values, please subset the data to the numeric values you would like. dplyr::select_if(data, is.numeric) can be helpful here!")
   } else {
 
       gather_cor(data,

### 2019-05-19 R/vis-expect.R

@@ -84,7 +84,7 @@ vis_expect <- function(data, expectation, show_perc = TRUE){
 
   vis_expect_plot <- data_expect %>%
     # expect_frame(expectation) %>%
-    dplyr::mutate(rows = 1:n()) %>%
+    dplyr::mutate(rows = 1:dplyr::n()) %>%
     tidyr::gather_(key_col = "variable",
                    value_col = "valueType",
                    gather_cols = names(.)[-length(.)]) %>%

### 2019-05-19 R/vis-guess.R

@@ -57,7 +57,8 @@ vis_guess <- function(x, palette = "default"){
       ggplot2::guides(fill = ggplot2::guide_legend(title = "Type")) +
       # flip the axes, add info for axes
       ggplot2::scale_x_discrete(position = "top",
-                                limits = names(x))
+                                limits = names(x)) +
+    ggplot2::theme(axis.text.x = ggplot2::element_text(hjust = 0))
 
   # specify a palette ----------------------------------------------------------
   add_vis_dat_pal(vis_plot, palette)
@@ -100,7 +101,8 @@ guess_type <- function(x){
 
   output[!nas] <- vapply(FUN = readr::guess_parser,
                          X = x[!nas],
-                         FUN.VALUE = character(1))
+                         FUN.VALUE = character(1),
+                         guess_integer = TRUE)
   output[nas] <- NA
   output
 

### 2019-05-19 R/vis-value.R

@@ -1,9 +1,15 @@
 #' Visualise the value of data values
 #'
-#' Visualise all of the values in the data on a 0 to 1 scale.
+#' Visualise all of the values in the data on a 0 to 1 scale. Only works on
+#'   numeric data - see examples for how to subset to only numeric data.
 #'
 #' @param data a data.frame
-#' @param na_colour a character vector of length one describing what colour you want the NA values to be. Default is "grey90"
+#' @param na_colour a character vector of length one describing what colour
+#'   you want the NA values to be. Default is "grey90"
+#' @param viridis_option A character string indicating the colormap option to
+#'   use. Four options are available: "magma" (or "A"), "inferno" (or "B"),
+#'   "plasma" (or "C"), "viridis" (or "D", the default option) and "cividis"
+#'   (or "E").
 #'
 #' @return a ggplot plot of the values
 #' @export
@@ -11,11 +17,25 @@
 #' @examples
 #'
 #' vis_value(airquality)
+#' vis_value(airquality, viridis_option = "A")
+#' vis_value(airquality, viridis_option = "B")
+#' vis_value(airquality, viridis_option = "C")
+#' vis_value(airquality, viridis_option = "E")
 #'
-vis_value <- function(data, na_colour = "grey90") {
-  scale_01 <- function(x) {
-    (x - min(x, na.rm = TRUE)) / diff(range(x, na.rm = TRUE))
-  }
+#' \dontrun{
+#' library(dplyr)
+#' diamonds %>%
+#'   select_if(is.numeric) %>%
+#'   vis_value()
+#' }
+#'
+vis_value <- function(data,
+                      na_colour = "grey90",
+                      viridis_option = "D") {
+
+  if (!all_numeric(data)) {
+    stop("data input can only contain numeric values, please subset the data to the numeric values you would like. dplyr::select_if(data, is.numeric) can be helpful here!")
+  } else {
 
   purrr::map_dfr(data, scale_01) %>%
     vis_gather_() %>%
@@ -26,7 +46,9 @@ vis_value <- function(data, na_colour = "grey90") {
     # add info about the axes
     ggplot2::scale_x_discrete(position = "top") +
     ggplot2::theme(axis.text.x = ggplot2::element_text(hjust = 0)) +
-    ggplot2::scale_fill_viridis_c(option = "D",
+    ggplot2::scale_fill_viridis_c(option = viridis_option,
                                   na.value = na_colour)
+
+  }
 
 }

### 2019-05-19 tests/testthat/test-vis-guess.R

@@ -1,5 +1,7 @@
 context("vis_guess")
 
+readr::guess_parser(typical_data$Sex[1])
+
 # try out all the options
 vis_guess_plot <- vis_guess(typical_data)
 vis_guess_plot_cb_safe <- vis_guess(typical_data, palette = "cb_safe")

### 2019-10-28 README.Rmd
Update README.Rmd
Removing the spaces in " >= " results in proper in-line display of the R code under "Using `vis_expect()`". It appears that markdown parses the character combination " >" as a line break even when surrounded by back ticks. (Could be manually fixed in README.md by removing line break)

@@
 If you are not sure whether to expect values greater than 25 in your data
-(airquality), you could write: `vis_expect(airquality, ~.x >= 25)`, and you can
+(airquality), you could write: `vis_expect(airquality, ~.x>=25)`, and you can
 see if there are times where the  values in your data are greater than or equal

### 2019-12-15 README.Rmd
use updated code of conduct

@@
 
-Please note that this project is released with a 
-[Contributor Code of Conduct](CONDUCT.md). By participating in this project you 
-agree to abide by its terms.
+## Code of Conduct
+
+Please note that the visdat project is released with a [Contributor Code of Conduct](http://visdat.njtierney.com/CODE_OF_CONDUCT.html).
+By contributing to this project, you agree to abide by its terms.
 

### 2019-12-15 R/internals.R
fix documentation errors

@@ -169,11 +169,11 @@ if (palette == "default"){
 
 } # close the function
 
-#' (Internal) Create labels for the columns containing the \% missing data
+#' Create labels for the columns containing the percent missing data
 #'
 #' @param x data.frame
 #' @param col_order_index the order of the columns
-#'
+#' @note internal
 #' @return data.frame containing the missingness percent down to 0.1 percent
 #'
 label_col_missing_pct <- function(x,

### 2019-12-15 R/vis-binary.R

@@ -7,10 +7,10 @@
 #' @param col_na colour for NA, default is "grey90"
 #'
 #' @return a ggplot plot of the binary values
-#' @export
 #'
 #' @examples
 #' vis_binary(dat_bin)
+#' @export
 vis_binary <- function(data,
                        col_zero = "salmon",
                        col_one = "steelblue2",

### 2019-12-19 README.Rmd
fix documentation errors
updates README to include vis_binary

@@
 
-- `vis_guess()` visualise the individual class of earch value in your data
+- `vis_guess()` visualise the individual class of each value in your data
+- `vis_value()` visualise the value class of each cell in your data
+- `vis_binary()` visualise the occurrence of binary values in your data
 
@@
 
+## Using `vis_binary()`
+
+`vis_binary()` visualises binary values. See below for use with example data, `dat_bin`
+
+```{r vis-bin}
+vis_binary(dat_bin)
+```
+
+If you don't have only binary values a warning will be shown.
+
+```{r vis-bin-airq, eval = FALSE}
+vis_binary(airquality)
+```
+
+```
+Error in test_if_all_binary(data) : 
+  data input can only contain binary values - this means either 0 or 1, or NA. Please subset the data to be binary values, or see ?vis_value.
+```
+
 ## Using `vis_guess()`

### 2021-07-05 README.Rmd
use GH actions

@@
 
-[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/onboarding/issues/87)[![JOSS status](http://joss.theoj.org/papers/c85f57adbc565b064fb4bfc9b59a1b2a/status.svg)](http://joss.theoj.org/papers/c85f57adbc565b064fb4bfc9b59a1b2a)[![DOI](https://zenodo.org/badge/50553382.svg)](https://zenodo.org/badge/latestdoi/50553382)[![Travis Build Status](https://travis-ci.org/ropensci/visdat.svg?branch=master)](https://travis-ci.org/ropensci/visdat)[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/ropensci/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/ropensci/visdat/master.svg)](https://codecov.io/github/ropensci/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](http://cran.rstudio.com/web/packages/visdat/index.html)[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
+<!-- badges: start -->
+[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/onboarding/issues/87)[![JOSS status](http://joss.theoj.org/papers/c85f57adbc565b064fb4bfc9b59a1b2a/status.svg)](http://joss.theoj.org/papers/c85f57adbc565b064fb4bfc9b59a1b2a)[![DOI](https://zenodo.org/badge/50553382.svg)](https://zenodo.org/badge/latestdoi/50553382)[![R-CMD-check](https://github.com/ropensci/visdat/workflows/R-CMD-check/badge.svg)](https://github.com/ropensci/visdat/actions)[![Codecov test coverage](https://codecov.io/gh/ropensci/visdat/branch/master/graph/badge.svg)](https://codecov.io/gh/ropensci/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](http://cran.rstudio.com/web/packages/visdat/index.html)[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
+<!-- badges: end -->
 

### 2021-07-05 README.Rmd

@@
 <!-- badges: start -->
-[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/onboarding/issues/87)[![JOSS status](http://joss.theoj.org/papers/c85f57adbc565b064fb4bfc9b59a1b2a/status.svg)](http://joss.theoj.org/papers/c85f57adbc565b064fb4bfc9b59a1b2a)[![DOI](https://zenodo.org/badge/50553382.svg)](https://zenodo.org/badge/latestdoi/50553382)[![R-CMD-check](https://github.com/ropensci/visdat/workflows/R-CMD-check/badge.svg)](https://github.com/ropensci/visdat/actions)[![Codecov test coverage](https://codecov.io/gh/ropensci/visdat/branch/master/graph/badge.svg)](https://codecov.io/gh/ropensci/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](http://cran.rstudio.com/web/packages/visdat/index.html)[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
+[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/software-review/issues/87)[![JOSS status](https://joss.theoj.org/papers/10.21105/joss.00355/status.svg)](https://joss.theoj.org/papers/10.21105/joss.00355)[![DOI](https://zenodo.org/badge/50553382.svg)](https://zenodo.org/badge/latestdoi/50553382)[![R-CMD-check](https://github.com/ropensci/visdat/workflows/R-CMD-check/badge.svg)](https://github.com/ropensci/visdat/actions)[![Codecov test coverage](https://codecov.io/gh/ropensci/visdat/branch/master/graph/badge.svg)](https://codecov.io/gh/ropensci/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](https://CRAN.R-project.org/package=visdat)[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/)
 <!-- badges: end -->
@@
 `missing.pattern.plot` from the
-[`mi`](https://cran.r-project.org/web/packages/mi/index.html) package.
+[`mi`](https://CRAN.R-project.org/package=mi) package.
 Unfortunately `missing.pattern.plot` is no longer in the `mi` package (as of
@@
 
-Please note that the visdat project is released with a [Contributor Code of Conduct](http://visdat.njtierney.com/CODE_OF_CONDUCT.html).
+Please note that the visdat project is released with a [Contributor Code of Conduct](https://github.com/ropensci/visdat/blob/master/CODE_OF_CONDUCT.md).
 By contributing to this project, you agree to abide by its terms.
@@
 Thank you to Ivan Hanigan who [first
-commented](http://www.njtierney.com/r/missing%20data/rbloggers/2015/12/01/ggplot-missing-data/#comment-2388263747)
+commented](https://www.njtierney.com/post/2015/11/12/ggplot-missing-data/)
 this suggestion after I made a blog post about an initial prototype
@@
 Finally thank you to [rOpenSci](https://github.com/ropensci) and it's amazing
-[onboarding process](https://github.com/ropensci/onboarding), this process has
+[onboarding process](https://github.com/ropensci/software-review), this process has
 made visdat a much better package, thanks to the editor Noam Ross (@noamross),

### 2023-02-01 README.Rmd
update the README and outputs

@@
 <!-- badges: start -->
-[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/software-review/issues/87)[![JOSS status](https://joss.theoj.org/papers/10.21105/joss.00355/status.svg)](https://joss.theoj.org/papers/10.21105/joss.00355)[![DOI](https://zenodo.org/badge/50553382.svg)](https://zenodo.org/badge/latestdoi/50553382)[![R-CMD-check](https://github.com/ropensci/visdat/workflows/R-CMD-check/badge.svg)](https://github.com/ropensci/visdat/actions)[![Codecov test coverage](https://codecov.io/gh/ropensci/visdat/branch/master/graph/badge.svg)](https://codecov.io/gh/ropensci/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](https://CRAN.R-project.org/package=visdat)[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/)
+[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/software-review/issues/87)[![JOSS status](https://joss.theoj.org/papers/10.21105/joss.00355/status.svg)](https://joss.theoj.org/papers/10.21105/joss.00355)[![DOI](https://zenodo.org/badge/50553382.svg)](https://zenodo.org/badge/latestdoi/50553382)[![R-CMD-check](https://github.com/ropensci/visdat/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ropensci/visdat/actions/workflows/R-CMD-check.yaml)[![Codecov test coverage](https://codecov.io/gh/ropensci/visdat/branch/master/graph/badge.svg)](https://codecov.io/gh/ropensci/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](https://CRAN.R-project.org/package=visdat)[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/)
 <!-- badges: end -->
@@
 
+The `vis_dat()` function also has a `facet` argument, so you can create small multiples of a similar plot for a level of a variable, e.g., Month:
+
+```{r vis-dat-month}
+vis_dat(airquality, facet = Month)
+```
+
+These currently also exist for `vis_miss()`, and the `vis_cor()` functions.
+
 ## Using `vis_miss()`
@@
 
-Percentages of missing/complete in `vis_miss` are accurate to 1 decimal place.
+Percentages of missing/complete in `vis_miss` are accurate to the integer (whole number).
+To get more accurate and thorough exploratory summaries of missingness, I would recommend the [`naniar`](https://github.com/njtierney/naniar) R package
 

### 2023-02-01 README.Rmd
check URLs

@@
 <!-- badges: start -->
-[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/software-review/issues/87)[![JOSS status](https://joss.theoj.org/papers/10.21105/joss.00355/status.svg)](https://joss.theoj.org/papers/10.21105/joss.00355)[![DOI](https://zenodo.org/badge/50553382.svg)](https://zenodo.org/badge/latestdoi/50553382)[![R-CMD-check](https://github.com/ropensci/visdat/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ropensci/visdat/actions/workflows/R-CMD-check.yaml)[![Codecov test coverage](https://codecov.io/gh/ropensci/visdat/branch/master/graph/badge.svg)](https://codecov.io/gh/ropensci/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](https://CRAN.R-project.org/package=visdat)[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/)
+[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/software-review/issues/87)[![JOSS status](https://joss.theoj.org/papers/10.21105/joss.00355/status.svg)](https://joss.theoj.org/papers/10.21105/joss.00355)[![DOI](https://zenodo.org/badge/50553382.svg)](https://zenodo.org/badge/latestdoi/50553382)[![R-CMD-check](https://github.com/ropensci/visdat/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ropensci/visdat/actions/workflows/R-CMD-check.yaml)[![Codecov test coverage](https://codecov.io/gh/ropensci/visdat/branch/master/graph/badge.svg)](https://app.codecov.io/gh/ropensci/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](https://CRAN.R-project.org/package=visdat)[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/)
 <!-- badges: end -->
@@
 
-You can read more about visdat in the vignette, ["using visdat"](http://visdat.njtierney.com/articles/using_visdat.html).
+You can read more about visdat in the vignette, ["using visdat"]https://docs.ropensci.org/visdat/articles/using_visdat.html).
 

### 2023-02-01 R/visdat-package.r

@@ -15,7 +15,7 @@
 #'   \item [vis_expect()]
 #' }
 #'
-#' Learn more about visdat at \url{http://visdat.njtierney.com/articles/using_visdat.html}
+#' Learn more about visdat at \url{https://docs.ropensci.org/visdat/}
 #' @name visdat
 #' @docType package
 #' @importFrom magrittr %>%
